### 2022-05-08 paper.tex
initial

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{hyperref}

\graphicspath{ {./images/} }
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,
    citecolor=blue,
    urlcolor=blue,
    linktocpage=true
}
\setcounter{tocdepth}{2} %show more in the toc

\newcommand{\kw}[1]{\texttt{#1}}
\renewcommand{\contentsname}{table of content}
\usepackage{indentfirst}
\usepackage[english]{babel}

% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand\UrlFont{\color{blue}\rmfamily}

%

\title{Infra-estrutura de Testes para Implementações de Referência do Standard ECMAScript}
\subtitle{}

%
\titlerunning{Live Metadata for Test262}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Diogo Costa Reis\\ist187526\\
\email{diogo.costa.reis@tecnico.ulisboa.pt}}
%
\authorrunning{Diogo Costa Reis}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Instituto Superior Técnico\\
Av. Rovisco Pais, 1\\
1049-001 Lisboa\\
Tel: +351 218 417 000\\
\email{mail@tecnico.ulisboa.pt}}
%

\begin{document}

% a solution to remove title and author from appearing in the table of contents: https://tex.stackexchange.com/a/318220
{\def\addcontentsline#1#2#3{}\maketitle}

%
\begin{abstract}
% TODO

\keywords{ECMAScript \and Specification Language \and Reference Interpreters \and Test262}
\end{abstract}

\newpage

\tableofcontents

\newpage

\section{Introduction}
\label{sec:Introduction}

\section{Goals}
\label{sec:Goals}

\section{Background}
\label{sec:Background}
This chapter provides an overview on the ECMAScript standard, the Test262 that are used to test the correct implementation of the ECMAScript standard, and finally an outline of the new metadata generated.

\subsection{ECMAScript}
\label{subsec:ECMAScript}
% overview da linguagem JS

% paragrafo - porque e' que javascript e' relevante (uma das linguagens mais usadas no momento)
% Existem muitas implementacoes diferentes da linguagem: client-side (browsers), server-side (Node.js), embedded devices (Jerryscript) -> Estas implementacoes têm de estar de coincider no comportamento observavel -> é particularmente importante na Web -> senao temos sites que em ...
JavaScript (JS) is a programming language mainly used in the development of client side web applications, also being one of the most popular programming languages. According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as \colorbox{orange}{second most active} languages on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} and the second most active on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}

% overview do standard -> descrita num standard
% Porque é muito importante que as várias implementacoes da linguagem coincidam -> o JavaScript está especificado num documento semi-formal que ...
% Falar sobre o standard
%   - o standard está como um interpretador de JavaScript em
%     pseudo-codigo - descreve detalhadamente os passos que
%    um interpretador de JS tem de executar ao avaliar qualquer
%    statement da linguagem
% Falar sobre o comité -  quem controla a evolucao do ECMAScript 
ECMAScript standard is the official document in which the JavaScript language is defined. This document is in constant evolution, being updated by the ECMA Technical Committee 39 (TC39), which is responsible for maintaining the standard. The standard is currently in its twelfth version.
%
JavaScript is specified in the ECMAScript standard\cite{ECMAScriptStandard}, which is written in English, and subsequently implemented by various  compilers and interpreters. Some of the JavaScript compilers are the ~\cite{Hop} and the JSC~\cite{JSC} compilers.
Most popular interpreters being nodejs~\cite{nodejs} and spidermonkey~\cite{spidermonkey}.
%
The standard defines the types, values, objects, properties, syntax, and semantics of JavaScript that must be the same in every JavaScript compiler and interpreter, while allowing JavaScript implementations to define additional types, values, object, properties, and functions.

% Estrutura do standard
The JavaScript language can be divided into three major components, those being expressions and commands, built-in libraries, and finally internal functions.
%
\begin{itemize}
\item Expressions and commands describe the behavior of static constructions, detailing the semantics of the diverse expressions (e.g., assignment expressions, built-in operators, etc.), commands (e.g., loop commands, conditions command, etc.), and built-in types (Undefined, Null, Boolean, Number, String and Object).
%
\item The internal functions of the language are used to define the semantics for both expressions and commands, as well as the built-in libraries. Internal functions are not exposed beyond the internal context of the language. In other words, no JavaScript program uses internal functions directly.
%
\item Finally, built-in libraries encompass all the internal objects available when a JavaScript program is executed. Internal objects expose many functions implemented by the language itself, including functions to manipulate numbers, text, arrays, objects, amongst other things.
\end{itemize}

% MetaParagrafo tres tipos
% Expressions and statements
% Exemplo do standard  e explicacao (if)
%
\paragraph{Semantics of IF statement}
Figure \ref{fig:If-Else Statement} shows a snippet of the ECMAScript standard description of the \texttt{IF} command. In order to evaluate \texttt{IF} commands with the shape:

\begin{center}
\texttt{if (Expression) Statement1 else Statement2}
\end{center}

\noindent the language begins by evaluating the \texttt{Expression} storing the result in the variable \texttt{exprRef} (step 1). The previous step will be used as Boolean, therefore, the result of the previous step will be converted to a Boolean using the internat functions \texttt{ToBoolean} and \texttt{GetValue}, and having the result stored in the variable \texttt{exprValue} (step 2).A different \texttt{Statement} will be followed depending on \texttt{exprValue}. If \texttt{exprValue} has the value \texttt{true} the variable \texttt{stmtCompletion} will have the evaluation of the first \texttt{Statement} (step 3). Otherwise, the variable \texttt{stmtCompletion} will store the result of evaluating the second \texttt{Statement} (step 4). Finally, a \texttt{Completion} will be returned, if the \texttt{stmtCompletion} has non empty value then it will be returned, however, when the value is empty it will be replaced with undefined (step 5).

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\textwidth]{images/if_statement.png}
    \caption{ECMAScript definition of an if-else statement}
    \label{fig:If-Else Statement}
\end{figure}

% example of Array.pop
% arrays são objetos como os outros
% arrys tem propriedades especiais
\paragraph{Example of Array.pop}
Figure \ref{fig:Array_pop_example} shows a simplified version of before \texttt{(a)} and after \texttt{(b)} of an array performing the pop function.
Before preforming \texttt{pop} \texttt{(a)}, the array has three  properties \texttt{length}, \texttt{0}, and \texttt{1}. Property \texttt{length} represents the size of the array that has value \texttt{2}. While the properties \texttt{0} and \texttt{1} store the first and second elements of the array respectively.
After \texttt{pop} being preformed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}). The \texttt{length} property highlighted in green is also updated since the size of the array changes to one.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.4\textwidth]{images/array_pop_example.png}
    \caption{Example Array.pop}
    \label{fig:Array_pop_example}
\end{figure}

% Built-ins (Array.pop)
% printscreen do standard e explicacao
\paragraph{Semantics of the Pop function}
Figure \ref{fig:Array_pop} shows a snippet of the ECMAScript standard description of the pop function in the Array Built-in. To begin with, the array will be converted to and Object using the \texttt{ToObject} function, and stored in the \texttt{O} variable (step 1).
Afterwards, the array length of the previously calculated variable will be calculated with the \texttt{LengthOfArrayLike} internal function, and storing the result in the \texttt{len} variable (step 2).
At this point there are to ways to proceed depending on the value of \texttt{len}. If the value is zero, the Array is empty, then the property \texttt{length} of \texttt{O} is set to zero and \texttt{undefined} is returned (step 3).
Otherwise, when \texttt{len} is different from zero, meaning that the Array is not empty, the Array's last element will be removed (described in Figure \ref{fig:Array_pop_example}) and returned (step 4).
To begin with, the language will assert that \texttt{len} is positive (step 4.a).
Afterwards, the \texttt{newLen} variable will store the value of \texttt{len} decremented by 1 (step 4.b).
The variable \texttt{index} will store the variable calculated in the previous step represented as a String converted with the \texttt{toString} function (step 4.c).
Then, stores the value of the \texttt{O} variable at the property corresponding to \texttt{index} in the \texttt{element} variable using the \texttt{Get} function (step 4.d).
Subsequently, deletes the previously mentioned property of the \texttt{O} variable with the \texttt{DeletePropertyOrThrow} function (step 4.e).
In addition, sets the \texttt{length} property of the \texttt{O} variable  to the \texttt{newLen} using the \texttt{Set} function (step 4.f).
Finally, returning the value of the variable \texttt{element} (step 4.g).

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.6\textwidth]{images/array_pop.png}
    \caption{ECMAScript definition of Array.pop}
    \label{fig:Array_pop}
\end{figure}

% Internal Functions (LenghtOfArrayLike)
% printscreen do standard e explicacao
\paragraph*{LengthOfArrayLike internal function}
Figure \ref{fig:LengthOfArrayLike} shows a snippet of the ECMAScript standard description of the \texttt{LengthOfArrayLike} internal function, that evaluates the function:

\begin{center}
\texttt{LengthOfArrayLike (obj)}
\end{center}

\noindent The language starts by asserting that \texttt{obj} is an \texttt{Object} (step 1). Afterwards, getting the value of the property \texttt{length} from \texttt{obj} using the function \texttt{Get}. Converting the previously mentioned value to an Integer that represents the length with the \texttt{ToLength} function, and finally returning said Integer.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{images/length_array_like.png}
    \caption{ECMAScript definition of the LengthOfArrayLike}
    \label{fig:LengthOfArrayLike}
\end{figure}

\subsection{Test262}
\label{subsec:Test262}

% paragrafo - JS tem muitas particularidades que dificultam o desemvolvimento e testagem  -> É muito dificil desenvolver
% novas implemenentacoes da linguagem
% Existe uma bateria de testes que testa as implementacoes da
% linguagem contra o standard
% Esta bateria de testes é dificil de manter
% Porque? - muitos testes, muitas features, em geral ha retrocompatibilidade mas ha um pequeno numero de casos onde a retrocompatibilidade nao se verifica -> os testes de ser modificados
Implementing a JavaScript engine is particularly difficult since it involves dealing with the many corner cases \colorbox{orange}{explain meaning of corner case} that exist in the language. In addition, while testing JavaScript code, test batteries will also need to test all corner cases to guarantee that they are correctly implemented.

%
% Implementacoes parciais da linguagem
% Na acadamia é normal desenvolverem-se implementacoes parciais da linguagem: não suportam a ultima versao, não suportam todos os objectos built-in, não suportam todas
% Pergunta: Quais é que são os testes apropriados?
% Normal: Respostas ad-hoc -> sem justificação rigorosa -> basicamente cada paper selecciona os testes que lhe da jeito

%%
%% Formato dos testes
% frontmatter, code
% exemplo

Every test of test262 has 2 parts, those being the frontmatter with some metadata about the test and the second part is the code of the test. As in the example \ref{fig:Test262_example}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=1\textwidth]{images/test262_array_test.png}
    \caption{Test262 es5id: 15.4.5-1}
    \label{fig:Test262_example}
\end{figure}

%% metadados ja incluidos
%% referir outra vez o exemplo

%% metadados que achamos relevantes e nao estao incluidos
%% ...

- formato de um teste
  - as várias secções e o seu próposito

- metadados oficialmente incluídos nos testes
  - que metadados é que os testes contém actualmente

- quais são os metadados que nós achamos serem relevantes e que estão em falta

\subsection{An Infrastructure for testing reference implementations of the ECMAScript standard}
\label{subsec:An Infrastructure for testing reference implementations of the ECMAScript standard}

%% Metodologia
- metadados da Test262 -> estruturados

- quais são os novos metadados incluídos

- sumário das abordagens utilizadas para os calcular

- shortcomings...

\section{Related Work}
\label{sec:Related Work}

\section{Design and Methodology}
\label{sec:Design and Methodology}

\section{Evaluation and Planning}
\label{sec:Evaluation and Planning}

\section{Conclusion}
\label{sec:Conclusion}

%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\bibliographystyle{splncs04}
\bibliography{references}

\end{document}

### 2022-05-10 paper.tex
pop example  and first paragraph

@@
 % paragrafo - porque e' que javascript e' relevante (uma das linguagens mais usadas no momento)
-% Existem muitas implementacoes diferentes da linguagem: client-side (browsers), server-side (Node.js), embedded devices (Jerryscript) -> Estas implementacoes têm de estar de coincider no comportamento observavel -> é particularmente importante na Web -> senao temos sites que em ...
 JavaScript (JS) is a programming language mainly used in the development of client side web applications, also being one of the most popular programming languages.
-According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as \colorbox{orange}{second most active} languages on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} and the second most active on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}
+According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as second most active languages on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} as well as on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}
 
+% Existem muitas implementacoes diferentes da linguagem: client-side (browsers), server-side (Node.js), embedded devices (Jerryscript) -> Estas implementacoes têm de estar de acordo no comportamento observavel -> é particularmente importante na Web -> senao temos sites que em ...
+% -----------------------
 % overview do standard -> descrita num standard
@@
 
+% Good melhorar se houver tempo no final
 % Estrutura do standard
@@
 % MetaParagrafo tres tipos
+The remaining subsection provides a description of the three types of artifact described in the standard.
+
 % Expressions and statements
 % Exemplo do standard  e explicacao (if)
-%
 \paragraph{Semantics of IF statement}
@@
 
-% example of Array.pop
 % arrays são objetos como os outros
 % arrys tem propriedades especiais
-\paragraph{Example of Array.pop}
-Figure \ref{fig:Array_pop_example} shows a simplified version of before \texttt{(a)} and after \texttt{(b)} of an array performing the pop function.
+% example of Array.pop
+% Built-ins (Array.pop)
+% printscreen do standard e explicacao
+\paragraph{Semantics of the Pop function}
+The Array built-in is an object as any other in JavaScript.
+The main difference is in its properties.
+Array Objects have a property \texttt{length} that contains the size of the array, as well as a property for each element of the array (from zero to \texttt{length} minus 1).
+
+Figure \ref{fig:Array_pop_example} shows a simplified version of an array performing the pop function, where \texttt{(a)} and \texttt{(b)} are the before and after respectively.
 Before preforming \texttt{pop} \texttt{(a)}, the array has three  properties \texttt{length}, \texttt{0}, and \texttt{1}.
 Property \texttt{length} represents the size of the array that has value \texttt{2}.
-While the properties \texttt{0} and \texttt{1} store the first and second elements of the array respectively.
+While the properties \texttt{0} and \texttt{1} store the first (\texttt{banana}) and second (\texttt{kiwi}) elements of the array respectively.
 After \texttt{pop} being preformed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}).
@@
 
+% TODO change length with quotation marks
 \begin{figure}[ht]
@@
 \end{figure}
-
-% Built-ins (Array.pop)
-% printscreen do standard e explicacao
-\paragraph{Semantics of the Pop function}
 Figure \ref{fig:Array_pop} shows a snippet of the ECMAScript standard description of the pop function in the Array Built-in.
@@
     \centering
-    \includegraphics[width=1\textwidth]{images/test262_array_test.png}
+    \includegraphics[width=1.2\textwidth]{images/test262_array_test.png}
     \caption{Test262 es5id: 15.4.5-1}

### 2022-05-11 paper.tex
fix small stuff

@@
 \noindent The language starts by asserting that \texttt{obj} is an \texttt{Object} (step 1).
-Afterwards, getting the value of the property \texttt{length} from \texttt{obj} using the function \texttt{Get}.
-Converting the previously mentioned value to an Integer that represents the length with the \texttt{ToLength} function, and finally returning said Integer.
+Afterwards, gets the value of the property \texttt{length} from \texttt{obj} using the function \texttt{Get}.
+Then, converts the previously mentioned value to an Integer that represents the length with the \texttt{ToLength} function, and finally returns said Integer (step 2).
 

### 2022-05-11 paper.tex
escrever 2 paragrafo (ECMA standard many implementations so needs to have a "rigid" standard)

@@
 % Falar sobre o comité -  quem controla a evolucao do ECMAScript
-ECMAScript standard is the official document in which the JavaScript language is defined.
+ECMAScript standard\cite{ECMAScriptStandard} is the official document in which the JavaScript language is defined.
 This document is in constant evolution, being updated by the ECMA Technical Committee 39 (TC39), which is responsible for maintaining the standard.
@@
 %
-JavaScript is specified in the ECMAScript standard\cite{ECMAScriptStandard}, which is written in English, and subsequently implemented by various  compilers and interpreters.
+JavaScript is specified in the ECMAScript standard using the English language.
+Subsequently, the standard is implemented by various  compilers and interpreters.
 Some of the JavaScript compilers are the ~\cite{Hop} and the JSC~\cite{JSC} compilers.
 Most popular interpreters being nodejs~\cite{nodejs} and spidermonkey~\cite{spidermonkey}.
+These are only four implementations amongst many others, which come along with the many use cases that JavaScript has.
+JavaScript is used mostly used in the web context, both client-side within browsers and server-side \texttt{(Node.js)}, but also in embedded devices.
 %
-The standard defines the types, values, objects, properties, syntax, and semantics of JavaScript that must be the same in every JavaScript compiler and interpreter, while allowing JavaScript implementations to define additional types, values, object, properties, and functions.
+Since JavaScript is used in so many scenarios and across so many different contexts, there is a great need for the ECMAScript standard to define in grate detail to ensure consistency.
+Browsers, for example, need to run JavaScript implementations that coincide so that websites are correctly rendered.
+In order to achieve coherent implementations, the standard defines the types, values, objects, properties, syntax, and semantics of JavaScript that must be the same in every JavaScript compiler and interpreter, while allowing JavaScript implementations to define additional types, values, object, properties, and functions.
 

### 2022-05-12 paper.tex
subsection test262 first version

--- 2022-05-11 paper.tex
@@
 % Falar sobre o comité -  quem controla a evolucao do ECMAScript
-ECMAScript standard is the official document in which the JavaScript language is defined.
+ECMAScript standard\cite{ECMAScriptStandard} is the official document, written in the English language, in which the JavaScript language is defined.
 This document is in constant evolution, being updated by the ECMA Technical Committee 39 (TC39), which is responsible for maintaining the standard.
@@
 %
-JavaScript is specified in the ECMAScript standard\cite{ECMAScriptStandard}, which is written in English, and subsequently implemented by various  compilers and interpreters.
-Some of the JavaScript compilers are the ~\cite{Hop} and the JSC~\cite{JSC} compilers.
-Most popular interpreters being nodejs~\cite{nodejs} and spidermonkey~\cite{spidermonkey}.
+The standard specifies the \texttt{JavaScript} language, to ensure it's multiple compilers and interpreters implementions are coherent.
+Some of the \texttt{JavaScript} compilers are the ~\cite{Hop} and the JSC~\cite{JSC} compilers, the most popular interpreters are nodejs~\cite{nodejs} and spidermonkey~\cite{spidermonkey}.
+These are only four implementations amongst many others, which come along with the many use cases that \texttt{JavaScript} has.
 %
-The standard defines the types, values, objects, properties, syntax, and semantics of JavaScript that must be the same in every JavaScript compiler and interpreter, while allowing JavaScript implementations to define additional types, values, object, properties, and functions.
+\texttt{JavaScript} is used mostly used in the web context, both client-side within browsers and server-side, but also in embedded devices.
+Since \texttt{JavaScript} is used in so many scenarios and across so many different contexts, there is highly important that ECMAScript standard is defined in great detail to ensure consistency.
+Browsers, for example, need to run \texttt{JavaScript} implementations that coincide so that websites are correctly rendered.
+In order to achieve coherent implementations, the standard defines the types, values, objects, properties, syntax, and semantics of \texttt{JavaScript} that must be the same in every \texttt{JavaScript} compiler and interpreter, while allowing \texttt{JavaScript} implementations to define additional types, values, object, properties, and functions.
 
-% Good melhorar se houver tempo no final
+% Good, more detail if there is time
 % Estrutura do standard
-The JavaScript language can be divided into three major components, those being expressions and commands, built-in libraries, and finally internal functions.
+The \texttt{JavaScript} language can be divided into three major components, those being expressions and commands, built-in libraries, and finally internal functions.
 %
@@
 % Porque? - muitos testes, muitas features, em geral ha retrocompatibilidade mas ha um pequeno numero de casos onde a retrocompatibilidade nao se verifica -> os testes de ser modificados
-Implementing a JavaScript engine is particularly difficult since it involves dealing with the many corner cases \colorbox{orange}{explain meaning of corner case} that exist in the language.
-In addition, while testing JavaScript code, test batteries will also need to test all corner cases to guarantee that they are correctly implemented.
+Implementing a \texttt{JavaScript} engine is particularly difficult since it involves dealing with the many corner cases that exist in the language.
+To test that corner cases are \colorbox{orange}{correctly dealt with there is \texttt{Test262}\cite{Test262}}, the ECMAScript standard test battery.
+Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity (large quantity of tests and features).\texttt{Test262} complexity grows with changes to the standard since in most cases retrocompatibility is maintained except for a few select cases.
+%
+% doesn't that imply wrongly that Test262 is comple???
+%
 
-%
 % Implementacoes parciais da linguagem
@@
 % Normal: Respostas ad-hoc -> sem justificação rigorosa -> basicamente cada paper selecciona os testes que lhe da jeito
+Due to ECMAScript standard being so extensive most implementations are only partial (namely implementations in academic context).
+Since the implementations are only partial, only tests for features that are implemented are relevant.
+Since selecting the relevant tests is not a trivial matter, each development team selects the appropiate tests.
+Each development selecting the tests raises the problem that there is not standard and precise way of picking the all the right tests.
 
@@
 % exemplo
-
-Every test of test262 has 2 parts, those being the frontmatter with some metadata about the test and the second part is the code of the test. As in the example \ref{fig:Test262_example}.
+Figure \ref{fig:Test262_example} shows a test from \texttt{Test262}.
+Every test of test262 has 3 parts: first is the copyright section represented with the comment \texttt{//} (lines 1 and 2), second is the \texttt{Frontmatter} section between \texttt{/*---} and \texttt{---*/} with some metadata about the test (lines 4 to 7), and finally is the \texttt{Body} section with the code of the test (lines 9 to 12).
+The copyright section has information about the owner and licence of the test.
+The \texttt{Frontmatter} section has the the test id (15.4.5-1) and a description of the test.
+Finally, the \texttt{Body}'s code tests the correct implementation of the standard.
 
@@
 %% referir outra vez o exemplo
+% - metadados oficialmente incluídos nos testes
+%   - que metadados é que os testes contém actualmente
+The \texttt{Frontmatter} has keywords to hold metadata of the test. These keywords are associated with some type of metadata about the test. Bellow is the list of possible keywords and its meaning:
+
+\begin{itemize}
+\item \texttt{description} - contains a short description about what will be tested;
+%
+\item \texttt{esid} - contains the hash identifier of the ECMAScript portion associated with the feature that will be tested (the identifier references the most recent version of ECMAScript when the test is created);
+%
+\item \texttt{info} - contains a deeper explanation of the test behavior, frequently includes a direct citation of the standard;
+%
+\item \texttt{negative} - indicates that the test throws an error; associated to the keyword will be the type of error the test is supposed to throw (e.g.
+\texttt{TypeError}, \texttt{ReferenceError}) as well as the phase in which the error is expected to be thrown (e.g.
+\texttt{parse} vs \texttt{resolution} vs \texttt{runtime});
+%
+\item \texttt{includes} - contains the list of harness files that should be included in the execution of the tests;
+%
+\item \texttt{author} - contains the identification of the author of the test;
+%
+\item \texttt{flags} - contains a list of booleans for each property.
+The properties being: 1 \texttt{onlyStrict}, the test is only executed in strict mode; 2 noStrict, the test will only be executed in mode "sloppy"; 3 \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; 4 \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; 5 \texttt{async}, only for asyncronus tests; 6 \texttt{generated}, flag that identifies files created by a test; 7 \texttt{CanBlockIsFalse} and 8 \texttt{CanBlockIsTrue}, indicatethe value of the property \texttt{CanBlock} ; 9 \colorbox{orange}{\texttt{non-deterministic}}, idicates that the semantics used on the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
+%
+% non-deterministic nao esta na tese do miguel trigo
+%
+%
+\item \texttt{features} - containds a list of features that are used in the test;
+%
+\item \texttt{es5id} and \texttt{es6id} - indicates that the feature being tested belongs to ECMAScript 5 and 6 respectively and contains the hash identifier of the section of the standard it bellongs to; these keywords have been deprecated and substituted by esid;
+\end{itemize}
 
@@
 %% ...
+The example in \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id.
+Besides the obvious upgrade from \texttt{es5id} to \texttt{esid} it would be usefull to have \texttt{includes} with the harness files needed to execute the test.
+The harness information is very usefull since it makes easly available what part of the harness is needed to run that test, opening the door for loading only part of the harness instead of the whole harness which is the current approach.
 
-- formato de um teste
-  - as várias secções e o seu próposito
+% - quais são os metadados que nós achamos serem relevantes e que estão em falta
+In order to have a more complete \texttt{Frontmatter} we suggest adding the following information:
 
-- metadados oficialmente incluídos nos testes
-  - que metadados é que os testes contém actualmente
+\begin{itemize}
+    \item \texttt{static construct} - list of static constructions used in the test
+    \item \texttt{version} - the ECMAScript version of the standard after which the test belongs to
+    \item \texttt{built-ins} - list of all the built-ins used in the test
+\end{itemize}
 
-- quais são os metadados que nós achamos serem relevantes e que estão em falta
+This metadata provides helpfull information to solve the problem mentioned before of selecting the relevant tests for partial implementations.
 
@@
 %% Metodologia
-- metadados da Test262 -> estruturados
+% - metadados da Test262 -> estruturados
 
-- quais são os novos metadados incluídos
+% - quais são os novos metadados incluídos
 
-- sumário das abordagens utilizadas para os calcular
+% - sumário das abordagens utilizadas para os calcular
 
-- shortcomings...
+% - shortcomings...
 

### 2022-05-16 paper.tex
resolver a maioria dos comentarios do prof

--- 2022-05-11 paper.tex
@@
 
+% Boa!
 % Existem muitas implementacoes diferentes da linguagem: client-side (browsers), server-side (Node.js), embedded devices (Jerryscript) -> Estas implementacoes têm de estar de acordo no comportamento observavel -> é particularmente importante na Web -> senao temos sites que em ...
@@
 %
-JavaScript is specified in the ECMAScript standard\cite{ECMAScriptStandard}, which is written in English, and subsequently implemented by various  compilers and interpreters.
-Some of the JavaScript compilers are the ~\cite{Hop} and the JSC~\cite{JSC} compilers.
-Most popular interpreters being nodejs~\cite{nodejs} and spidermonkey~\cite{spidermonkey}.
+The standard specifies the \texttt{JavaScript} language, to ensure its multiple compilers and interpreters implementations are coherent.
+Some of the \texttt{JavaScript} compilers are the Hop~\cite{Hop} and the JSC~\cite{JSC} compilers, the most popular interpreters are Node.js~\cite{Node.js} and SpiderMonkey~\cite{SpiderMonkey}.
+These are only four implementations among many others, which come along with the many use cases that \texttt{JavaScript} has.
 %
-The standard defines the types, values, objects, properties, syntax, and semantics of JavaScript that must be the same in every JavaScript compiler and interpreter, while allowing JavaScript implementations to define additional types, values, object, properties, and functions.
+\texttt{JavaScript} is mostly used in the web context, both client-side within browsers and server-side, but also in embedded devices.
+Since \texttt{JavaScript} is used in so many scenarios and across so many different contexts, it is highly important that ECMAScript standard is defined in great detail to ensure consistency.
+Browsers, for example, need to run \texttt{JavaScript} implementations that coincide so that websites are correctly rendered and exhibit the same behavior.
+In order to achieve coherent implementations, the standard defines the types, values, objects, properties, syntax, and semantics of \texttt{JavaScript} that must be the same in every \texttt{JavaScript} compiler and interpreter, while allowing \texttt{JavaScript} implementations to define additional types, values, object, properties, and functions.
 
-% Good melhorar se houver tempo no final
+% Good, more detail if there is time
 % Estrutura do standard
-The JavaScript language can be divided into three major components, those being expressions and commands, built-in libraries, and finally internal functions.
+The \texttt{JavaScript} language can be divided into three major components, those being expressions and commands, built-in libraries, and finally internal functions.
 %
@@
 \noindent the language begins by evaluating the \texttt{Expression} storing the result in the variable \texttt{exprRef} (step 1).
-The previous step will be used as Boolean, therefore, the result of the previous step will be converted to a Boolean using the internat functions \texttt{ToBoolean} and \texttt{GetValue}, and having the result stored in the variable \texttt{exprValue} (step 2).A different \texttt{Statement} will be followed depending on \texttt{exprValue}.
+The previous step will be used as Boolean, therefore, the result of the previous step will be converted to a Boolean using the internal functions \texttt{ToBoolean} and \texttt{GetValue}, and having the result stored in the variable \texttt{exprValue} (step 2).A different \texttt{Statement} will be followed depending on \texttt{exprValue}.
 If \texttt{exprValue} has the value \texttt{true} the variable \texttt{stmtCompletion} will have the evaluation of the first \texttt{Statement} (step 3).
@@
 Before preforming \texttt{pop} \texttt{(a)}, the array has three  properties \texttt{length}, \texttt{0}, and \texttt{1}.
-Property \texttt{length} represents the size of the array that has value \texttt{2}.
-While the properties \texttt{0} and \texttt{1} store the first (\texttt{banana}) and second (\texttt{kiwi}) elements of the array respectively.
-After \texttt{pop} being preformed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}).
-The \texttt{length} property highlighted in green is also updated since the size of the array changes to one.
+Property \texttt{length} represents the size of the array that has value \texttt{2}, while the properties \texttt{0} and \texttt{1} store the first (\texttt{banana}) and second (\texttt{kiwi}) elements of the array respectively.
+After \texttt{pop} is preformed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}) and the \texttt{length} property (highlighted in green) is decremented by one since the size of the array changes to one.
 
@@
 % Porque? - muitos testes, muitas features, em geral ha retrocompatibilidade mas ha um pequeno numero de casos onde a retrocompatibilidade nao se verifica -> os testes de ser modificados
-Implementing a JavaScript engine is particularly difficult since it involves dealing with the many corner cases \colorbox{orange}{explain meaning of corner case} that exist in the language.
-In addition, while testing JavaScript code, test batteries will also need to test all corner cases to guarantee that they are correctly implemented.
+Implementing a \texttt{JavaScript} engine is particularly difficult since it involves dealing with the many corner cases that exist in the language.
+To test that corner cases are correctly dealt with there is \texttt{Test262}\cite{Test262}, the ECMAScript standard test battery.
+Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity, the total number of tests is around 40000 divided into \colorbox{orange}{87} subfolders, each correspond to roughly one section of the standard.
+\texttt{Test262} complexity grows with changes to the standard since in most cases backward compatible is maintained except for a few select cases.
+% 87 subfolders on built-ins + language
+% 105 subfolders if add intl402 (which also has files that were ignore)
+%
 
-%
 % Implementacoes parciais da linguagem 
@@
 % Normal: Respostas ad-hoc -> sem justificação rigorosa -> basicamente cada paper selecciona os testes que lhe da jeito
+Due to the ECMAScript standard being so extensive most implementations are only partial, especially implementations and analysis developed in academic contexts.
+In order to test partial implementations, one must be able to obtain the applicable set of tests from all the tests contained in Test262.
+Selecting the applicable tests is not a trivial matter because there are too many tests and too many features.
+The current methodology is that each development team manually selects the tests that are applicable to their corresponding implementation.
+This raises the problem that there is not standard and precise way of picking the all the right tests.
+% TODO final deste paragrafo
+% How can one be sure to have all the correct applicable tests among 40K tests?
 
-%%
+% gosto
 %% Formato dos testes
@@
 % exemplo
-
-Every test of test262 has 2 parts, those being the frontmatter with some metadata about the test and the second part is the code of the test. As in the example \ref{fig:Test262_example}.
+Figure \ref{fig:Test262_example} shows a test from \texttt{Test262}.
+Every test of \texttt{Test262} has 3 parts: first is the copyright section represented with the comment \texttt{//} (lines 1 and 2), second is the \texttt{Frontmatter} section between \texttt{/*---} and \texttt{---*/} with some metadata about the test (lines 4 to 7), and finally is the \texttt{Body} section with the code of the test (lines 9 to 12).
+The copyright section has information about the owner and license of the test.
+The \texttt{Frontmatter} section has the test id (15.4.5-1) and a description of the test.
+Finally, the \texttt{Body}'s code tests the correct implementation of the standard.
 
@@
     \centering
-    \includegraphics[width=1.2\textwidth]{images/test262_array_test.png}
+    \includegraphics[width=1.0\textwidth]{images/test262_array_test.png}
     \caption{Test262 es5id: 15.4.5-1}
@@
 %% referir outra vez o exemplo
+% - metadados oficialmente incluídos nos testes
+%   - que metadados é que os testes contém actualmente
+The \texttt{Frontmatter} has keywords to hold metadata of the test.
+These keywords are associated with specific elements of metadata concerning the test.
+Bellow is the list of possible keywords and their meaning:
+
+\begin{itemize}
+\item \texttt{description} - contains a short description about what will be tested;
+%
+\item \texttt{esid} - contains the hash identifier of the ECMAScript portion associated with the feature that will be tested (the identifier references the most recent version of ECMAScript when the test is created);
+%
+\item \texttt{info} - contains a deeper explanation of the test behavior, frequently includes a direct citation of the standard;
+%
+\item \texttt{negative} - indicates that the test throws an error; associated to the keyword will be the type of error the test is supposed to be thrown (e.g.
+\texttt{TypeError}, \texttt{ReferenceError}) as well as the phase in which the error is expected to be thrown (e.g.
+\texttt{parse} vs \texttt{resolution} vs \texttt{runtime});
+%
+\item \texttt{includes} - contains the list of \texttt{harness} files that should be included in the execution of the tests (\texttt{Test262} makes use of a large number of auxiliary function defined in a dedicated library referred to as the \texttt{Test262} \texttt{harness} described later in this section);
+%
+\item \texttt{author} - contains the identification of the author of the test;
+%
+\item \texttt{flags} - contains a list of booleans for each test property, the properties being: (1) \texttt{onlyStrict}, the test is only executed in strict mode; (2) noStrict, the test will only be executed in mode \emph{sloppy}; (3) \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; (4) \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; (5) \texttt{async}, the test is contains asynchronous functions; (6) \texttt{generated}, the test generates the files specified by the property; (7) \texttt{CanBlockIsFalse} and (8) \texttt{CanBlockIsTrue}, the test will run if the property \texttt{CanBlock} of the \texttt{Agent Record} executing it is false and true respectively; (9) \texttt{non-deterministic}, indicates that the semantics used in the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
+% TODO clarificar se e' necessario citar o que e' 'Agent Record' uma vez que nao vou explicar
+%
+%
+\item \texttt{features} - contains a list of features that are used in the test;
+%
+\item \texttt{es5id} and \texttt{es6id} - indicates that the feature being tested belongs to ECMAScript 5 and 6 respectively and contains the hash identifier of the section of the standard it belongs to; these keywords have been deprecated and substituted by esid.
+\end{itemize}
+
+% Varias observacoes: 
+%   - A metada esta muitas vezes INCOMPLETA e algumas vezes incorrecta
+
+%  - é preciso calcular a metadata certa
+
+% As Example 5 illustrates, it is often the case that the metadata of a test is incomplete.
+Some tests also have the wrong metatada.
+As part of this thesis, we plan to process all the tests to check and correct their corresponding metadata.
+%TODO
 
@@
 %% ...
+The example in Figure \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id.
+Besides the obvious upgrade from \texttt{es5id} to \texttt{esid} it would be useful to have \texttt{includes} with the \texttt{harness} files needed to execute the test.
+The \texttt{harness} information is very useful since it makes it easy to identify the part of the \texttt{harness} needed to run that test, opening the door for loading only part of the \texttt{harness} instead of the whole \texttt{harness} which is the current approach.
 
-- formato de um teste
-  - as várias secções e o seu próposito
+% - quais são os metadados que nós achamos serem relevantes e que estão em falta
+\paragraph{New Metadata}
+In order to have a more complete \texttt{Frontmatter} we suggest adding the following information:
 
-- metadados oficialmente incluídos nos testes
-  - que metadados é que os testes contém actualmente
+\begin{itemize}
+    \item \texttt{static construct} - list of static constructions used in the test
+    \item \texttt{version} - the ECMAScript version of the standard after which the test belongs to
+    \item \texttt{built-ins} - list of all the built-ins used in the test
+\end{itemize}
 
-- quais são os metadados que nós achamos serem relevantes e que estão em falta
+This metadata provides helpful information to solve the problem mentioned before of selecting the relevant tests for partial implementations.
 
@@
 \label{subsec:An Infrastructure for testing reference implementations of the ECMAScript standard}
+%% Metodologia
+% - metadados da Test262 -> estruturados
 
-%% Metodologia
-- metadados da Test262 -> estruturados
+% - quais são os novos metadados incluídos
 
-- quais são os novos metadados incluídos
+% - sumário das abordagens utilizadas para os calcular
 
-- sumário das abordagens utilizadas para os calcular
+% - shortcomings...
+% ----------------------------------------------------
+% This thesis continues the thesis of Miguel Trigo entitled ..., in which the author processed all Test262 tests, checked their metadata and computed 
+% Resumo da tese do Miguel
 
-- shortcomings...
+% - construir uma base de dados com a metadata dos varios testes em formato JSON
+
+% Exemplo
+
+% - Corrigir a metadata dos testes incorrectos e calcular metadata adicional
+
+% - Metadata adicional calculada parra cada teste:
+
+% -
+
+% -
+
+% -
+
+% Para cada um dos tipos de metadata 
+% um paragrafo a explicar sucintamente como é que a metada foi calculada
+
+% - static
+
+% - version
+
+% - built-ins
 
@@
 \label{sec:Design and Methodology}
+%  grandes objectivos
+
+% - processamento completo da metadata oficial 
+% -> o processamento do miguel esta incompleto -> nao incluir informacao informacao sobre -> ... 
+% [pode criticar como entender]
+
+% - calculo mais preciso da nova metadata proposta
+
+% - o calculo da versao pode ser melhorado de varias maneiras
+
+% - Precisao
+
+% - usar mais engines
+
+% - Eficiencia
+
+% - paralelizar os engines
+
+% - Sistema de visualisacao da metadata
+
+% printscreens do site
 
@@
 \label{sec:Evaluation and Planning}
+% TODO GANT Diagram
 

### 2022-05-17 paper.tex
first draft of the subsection on the thises of Miguel Trigo

@@
 % MetaParagrafo tres tipos
-The remaining subsection provides a description of the three types of artifact described in the standard.
+\colorbox{orange}{The remaining} subsection provides a description of the three types of artifact described in the standard.
+% QUESTION ainda nao foi lido pelo prof
 
@@
 To test that corner cases are correctly dealt with there is \texttt{Test262}\cite{Test262}, the ECMAScript standard test battery.
-Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity, the total number of tests is around 40000 divided into \colorbox{orange}{87} subfolders, each correspond to roughly one section of the standard.
+Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity, the total number of tests is around \colorbox{orange}{39837} divided into \colorbox{orange}{87} subfolders, each correspond to roughly one section of the standard.
 \texttt{Test262} complexity grows with changes to the standard since in most cases backward compatible is maintained except for a few select cases.
 % 87 subfolders on built-ins + language
-% 105 subfolders if add intl402 (which also has files that were ignore)
-%
+% QUESTION round number? (40000)
+% QUESTION 105 subfolders if add intl402 (which also has files that were ignore)
 
@@
 The current methodology is that each development team manually selects the tests that are applicable to their corresponding implementation.
-This raises the problem that there is not standard and precise way of picking the all the right tests.
-% TODO final deste paragrafo
-% How can one be sure to have all the correct applicable tests among 40K tests?
+This raises the problem that there is not standard and precise way of picking the all the right tests from the almost 40000 tests in \texttt{Test262}, making the possibility of human error when selecting the applicable tests likely.
 
@@
 Figure \ref{fig:Test262_example} shows a test from \texttt{Test262}.
-Every test of \texttt{Test262} has 3 parts: first is the copyright section represented with the comment \texttt{//} (lines 1 and 2), second is the \texttt{Frontmatter} section between \texttt{/*---} and \texttt{---*/} with some metadata about the test (lines 4 to 7), and finally is the \texttt{Body} section with the code of the test (lines 9 to 12).
+Every test of \texttt{Test262} has 3 parts: first is the copyright section represented with the comment \texttt{//} (lines 1 and 2), second is the \texttt{Frontmatter} section between \texttt{/*---} and \texttt{---*/} with some metadata about the test (lines 4 to 7), and finally is the \texttt{Body} section with the code of the test (lines 9 to 13).
 The copyright section has information about the owner and license of the test.
@@
 %
-\item \texttt{flags} - contains a list of booleans for each test property, the properties being: (1) \texttt{onlyStrict}, the test is only executed in strict mode; (2) noStrict, the test will only be executed in mode \emph{sloppy}; (3) \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; (4) \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; (5) \texttt{async}, the test is contains asynchronous functions; (6) \texttt{generated}, the test generates the files specified by the property; (7) \texttt{CanBlockIsFalse} and (8) \texttt{CanBlockIsTrue}, the test will run if the property \texttt{CanBlock} of the \texttt{Agent Record} executing it is false and true respectively; (9) \texttt{non-deterministic}, indicates that the semantics used in the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
-% TODO clarificar se e' necessario citar o que e' 'Agent Record' uma vez que nao vou explicar
+\item \texttt{flags} - contains a list of booleans for each test property, the properties being: (1) \texttt{onlyStrict}, the test is only executed in strict mode; (2) noStrict, the test will only be executed in mode \emph{sloppy}; (3) \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; (4) \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; (5) \texttt{async}, the test is contains asynchronous functions; (6) \texttt{generated}, the test generates the files specified by the property; (7) \texttt{CanBlockIsFalse} and (8) \texttt{CanBlockIsTrue}, the test will run if the property \texttt{CanBlock} of the \colorbox{orange}{\texttt{Agent Record}} executing it is false and true respectively; (9) \texttt{non-deterministic}, indicates that the semantics used in the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
+% QUESTION clarificar se e' necessario citar o que e' 'Agent Record' uma vez que nao vou explicar
 %
@@
 %  - A metada esta muitas vezes INCOMPLETA e algumas vezes incorrecta 
+%  - é preciso calcular a metadata certa
+As Figure 5 illustrates, it is often the case that the metadata of a test is incomplete.
+Some tests also have the wrong metadata.
+As part of this thesis, we plan to process all the tests to check and correct their corresponding metadata as well as completing the metadata that is \colorbox{orange}{missing.}
 
-%  - é preciso calcular a metadata certa
-
-% As Example 5 illustrates, it is often the case that the metadata of a test is incomplete.
-Some tests also have the wrong metatada.
-As part of this thesis, we plan to process all the tests to check and correct their corresponding metadata.
-%TODO
+%% QUESTION juntar os paragrafos?
 
@@
 %% ...
-The example in Figure \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id.
+\colorbox{orange}{The example} in Figure \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id.
 Besides the obvious upgrade from \texttt{es5id} to \texttt{esid} it would be useful to have \texttt{includes} with the \texttt{harness} files needed to execute the test.
@@
 \begin{itemize}
-    \item \texttt{static construct} - list of static constructions used in the test
-    \item \texttt{version} - the ECMAScript version of the standard after which the test belongs to
-    \item \texttt{built-ins} - list of all the built-ins used in the test
+    \item \texttt{syntactic construct} - list of all syntactic constructions used in the test;
+    \item \texttt{version} - the ECMAScript version of the standard in which the feature being tested was introduced;
+    \item \texttt{built-ins} - list of all the built-ins used in the test.
+    \item \colorbox{orange}{\texttt{harness-functions}} - list of all the harness functions used to asses the test's results.
 \end{itemize}
+% QUESTION posso adicionar isto?
+acho que era suposto fazer parte da minha tese; se calhar nao faz sentido, mas sim ter a informacao do includes segundo o standar (com o nome dos ficheiros da harness necessarios para o test)
 
-This metadata provides helpful information to solve the problem mentioned before of selecting the relevant tests for partial implementations.
+This metadata provides helpful information to solve the problem mentioned before, selecting the applicable tests for partial implementations, by allowing developers to filter tests by \texttt{builtin}, \texttt{static construct} and \texttt{version} of the \texttt{ECMAScript} standard.
+This would provide consistency and standardization to the selection of applicable tests to a partial implementation of the standard.
+As for the \texttt{harness-functions}, it provides the information of about the functions of \texttt{harness} that are used in the test, that is relevant because only a small part of the \texttt{harness} is need in each test even though the \colorbox{orange}{whole library is loaded}.
+% QUESTION suponho que o maior problema seja que todos testes da harness sejam corridos e nao o facto de toda a biblioteca seja carregada
 
@@
 \label{subsec:An Infrastructure for testing reference implementations of the ECMAScript standard}
-%% Metodologia
-% - metadados da Test262 -> estruturados
 
-% - quais são os novos metadados incluídos
+% Resumo da tese do Miguel 
+This thesis continues the work done in the master thesis of Miguel Trigo titled \emph{Infra-estrutura de Testes para Implementações de Referência do Standard ECMAScript}.
+In Miguel Trigo's master thesis, he \colorbox{orange}{processed} all Test262 tests and checked their metadata,  correcting some errors that \colorbox{orange}{were} found and complementing some metadata that is missing in the \texttt{Frontmatter}.
+The author also added new metadata that out of the scope of the \texttt{Frontmatter}, added statistical data about the tests, and made all the previously mentioned data available in a \texttt{MongoDB} database.
+% QUESTION da' a sensacao que o miguel fez tudo a mao? devia mencionar que ele automou o processo de gerar metadata?
+% QUESTION past tense?
 
-% - sumário das abordagens utilizadas para os calcular
+% - Corrigir a metadata dos testes incorrectos e calcular metadata adicional ????
+% - construir uma base de dados com a metadata dos varios testes em formato JSON 
+Miguel Trigo processed the test in Figure \ref{fig:Test262_example} and generated the metadata in Figure \ref{fig:json_metadata}.
+The \texttt{Frontmatter} in the test has the \texttt{es5id} and \texttt{description}, in the generated metadata they are represented for the \texttt{name} and \texttt{value} pairs \texttt{esid} and \texttt{description} respectively \colorbox{orange}{(updating the deprecated es5id)}.
+% QUESTION conta como correcao? eu devo arranjar um exemplo de uma correcao que ele fez? ou so' mensionar?
+%
+%
+The following \texttt{names} were added by Miguel Trigo:
 
-% - shortcomings...
-% ----------------------------------------------------
-% This thesis continues the thesis of Miguel Trigo entitled ..., in which the author processed all Test262 tests, checked their metadata and computed 
-% Resumo da tese do Miguel
-
-% - construir uma base de dados com a metadata dos varios testes em formato JSON
+\begin{itemize}
+    \item \texttt{path} - stores the path to the test within \texttt{Test262};
+    \item \texttt{version} - stores the version of the \texttt{ECMAScript} standard that the test corresponds to;
+    \item \texttt{built-ins} - stores a subpart of the \texttt{path};
+    \item \texttt{Array} - stores the subpart of the \texttt{path} that follows \texttt{built-ins};
+    \item \texttt{syntactic\_construct} - stores an array with all the syntactic constructions used in the test;
+    \item \texttt{builtIns} - stores an array with all the builtins used in the test;
+    \item \texttt{asserts} - stores the number of times the \texttt{assert} library is used in the test;
+    \item \texttt{error} - \colorbox{orange}{TODO Nao sei o que representa};
+    \item \texttt{esprima} - stores whether or not the test is supported by \texttt{Esprima}\cite{Esprima} (\texttt{Esprima} is an standard-compliant \texttt{ECMAScript} parser that is also developed in \texttt{ECMAScript});
+    \item \texttt{lines} - stores the number of lines of code written in the test, ignoring the empty lines and comments (in this example the \texttt{value} is 3 because lines 12 and 13 in Figure \ref{fig:Test262_example} are one single line in the actual test in \texttt{Test262} that was split in the example to make the image more legible).
+\end{itemize}
 
 % Exemplo
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/json_metadata.png}
+    \caption{Metadata generated for test esid: 15.4.5-1.js}
+    \label{fig:json_metadata}
+\end{figure}
 
-% - Corrigir a metadata dos testes incorrectos e calcular metadata adicional
-
-% - Metadata adicional calculada parra cada teste:
-
-% -
-
-% -
-
-% -
+The metadata can be separated into four parts.
+Firstly, the \texttt{Frontmatter} part, \texttt{description} and \texttt{esid}, that stores information related to the \texttt{Frontmatter} of the test.
+Secondly, the path part comprised of \texttt{path}, \texttt{built-ins}, and \texttt{Array}, that contains the information related to the path of the test inside the \texttt{Test262}.
+Thirdly, the statistical part formed by \texttt{asserts}, \texttt{error}, \texttt{esprima}, and \texttt{lines}, that provides statistics on the code of the test and the metadata generation process.
+%
+% - Metadata adicional calculada para cada teste:
+Finally, the new metadata part which consists of \texttt{version}, \texttt{syntactic\_construct}, and \texttt{builtIns}, the new metadata is relevant information about the tests that should be added to the \texttt{Frontmatter}.
 
 % Para cada um dos tipos de metadata 
-% um paragrafo a explicar sucintamente como é que a metada foi calculada
+In the remaining of the subsection is concerned with how the new metadata, namely \texttt{syntactic\_construct}, \texttt{builtIn}, and \texttt{version}.
 
-% - static
+% - how syntactic cosntruct is calculated 
+\paragraph{Calculation of syntactic\_Construct metadata}
+The \texttt{syntactic\_construct} metadata is calculated with the help of \texttt{Esprima}, which allows the analysis of the \colorbox{orange}{syntactic tree} of the test.
+The tree analysis of the test is represented in a JSON format and the \texttt{syntactic\_constructs} are the \texttt{values} corresponding to the \texttt{names} \emph{type} of every element of the tree.
+This way, to get all the \texttt{syntactic\_construct} is only a matter of traversing the tree and adding any new one found.
+The main problem being that \texttt{Esprima} only supports till version 7 of the \texttt{ECMAScript} standard, and also the \texttt{negative} tests that \texttt{Esprima} does not support.
+% QUESTION imagem do JSON retornado pelo esprima?
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\textwidth]{images/test_syntactic_tree.png}
+    \caption{Example test}
+    \label{fig:test_syntactic_tree}
+\end{figure}
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=1\textwidth]{images/JSON_syntactic_tree.png}
+    \caption{syntactic tree of the example test in Figure \ref{fig:test_syntactic_tree}}
+    \label{fig:JSON_syntactic_tree}
+\end{figure}
 
-% - version
+% - how built-ins is calculated
+\paragraph{Calculation of built-ins metadata}
+The \texttt{built-ins} are calculated with two separate methods.
+The first method makes use of syntactic tree generated from \texttt{Esprima}, to find key \texttt{static\_constructs} to identify the \texttt{built-ins} being used.
+While the second method goes through the code of a test and finds all the functions and associates them with their \texttt{built-ins}, as well as the finding \texttt{built-ins} being used directly.
+In both methods, after all the \texttt{built-ins} of a test are identified, the \texttt{built-in} that corresponds to the most recent version of the \texttt{ECMAScript} standard identifies the version of that test.
 
-% - built-ins
+% - how version is calculated 
+\paragraph{Calculation of Version metadata}
+The \texttt{version} metadata is calculated by identifying the element, \texttt{built-in}, \texttt{syntactic\_construct} or change in behavior of functions that is part of the newest version of the \texttt{ECMAScript} standard.
+The version is calculated using 3 different approaches dynamic, static, and mixed.
+%
+The dynamic approach is based on the waterfall model, running the tests against an implementation of the standard.
+if the test passes then the test is associated with that version of the standard, otherwise the test will be run against the next version of the implementation of the standard.
+%
+The static approach uses the syntactic tree generated with \texttt{Esprima}.
+From the analysis of the syntactic tree all the objects, functions, properties, operators, variables, and syntactical constructions present in the test are found, then searches for the oldest version that contains all of them.
+%
+The mixed approach is calculated using the dynamic and static approaches.
+The mixed approach consists of running the results of the dynamic approach of each version and verifying that is does not use features introduced in a posterior version.
+The dynamic approach makes use of \texttt{SpyderMonkey} and \texttt{Node.js} implementations of the standard.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=1\textwidth]{images/waterfall_model.png}
+    \caption{waterfall model of the dynamic approach}
+    \label{fig:waterfall_model}
+\end{figure}
 
@@
 % TODO GANT Diagram
+\colorbox{orange}{pedir descricao da teses de uma das primeiras reunioes (iPad)}
 

### 2022-05-17 paper.tex
seccao 3.3 pre reuniao

@@
 \label{sec:Design and Methodology}
+
 %  grandes objectivos 
+This thesis is a continuation of Miguel Tringo's master thesis, aiming to complete the calculation of the \texttt{Frontmatter} metadata especially the \colorbox{orange}{metadata about the \texttt{harness}}, also aims to improve the generated metadata.
+Another aim of this thesis is to create a website for visualizing the metadata generated.
+\colorbox{orange}{Finally,} we aim to build a platform that allows users to submit the markdown generated from executing the \texttt{Test262} and compare different runs.
+% QUESTION campo features? (restantes talvez possam ser ignorados author, negative, flags, info and description)
+% QUESTION incluir?
 
@@
 % [pode criticar como entender]
+Miguel Trigo's thesis metadata is incomplete in various ways, tests with unknown \texttt{version} around 9000, tests without \texttt{built-ins} around 17000, and tests without \texttt{syntactic\_construct} around 13000 \colorbox{orange}{tests without esid around 600}.
+% QUESTION sao tests da versao seginte de ECMA que ainda nao tem seccao na versao actual?
+%
+The metadata from the thesis could be improved in way the data is arranged, for example, the subfolders of the path to the test are spread into the JSON Object of the test.
+The subfolders information being put into an array of ordered subfolders would increase the readability.
 
 % - calculo mais preciso da nova metadata proposta
-
 % - o calculo da versao pode ser melhorado de varias maneiras
-
+This thesis plans to improve the \texttt{version} metadata generation in two dimensions precision and efficiency.
 % - Precisao 
-
 % - usar mais engines
-
+For the precision, with more \texttt{JavaScript} engines being used, there would be more certainty when determining the version.
 % - Eficiencia 
-
 % - paralelizar os engines 
+As for the efficiency, it would be possible to parallelize the waterfall model of the dynamic approach running multiple tests at the same time.
+It is also possible to use \texttt{git diff} to identify the tests that were added or changed since the last time the dynamic approach was executed, \colorbox{orange}{ only needing to execute} the dynamic approach on the differences.
+% QUESTION e' problema se fizerem update noutro lado que afeta o teste indiretamente?
 
 % - Sistema de visualisacao da metadata 
+The website for visualizing the metadata aims to make access to the metadata and filtering it easily accessible. The website is planned to allow the search
 
 % printscreens do site 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=1\textwidth]{images/website.png}
+    \caption{\colorbox{orange}{Website} for searching the metadata in construction}
+    % QUESTION letra maiuscula na caption
+    \label{fig:website}
+\end{figure}
 

### 2022-05-21 paper.tex
rewritting MetaData v0, version before meeting with the teacher

@@
 \item Finally, built-in libraries encompass all the internal objects available when a JavaScript program is executed.
-Internal objects expose many functions implemented by the language itself, including functions to manipulate numbers, text, arrays, objects, amongst other things.
+Internal objects expose many functions implemented by the language itself, including functions to manipulate numbers, text, arrays, objects, among other things.
 \end{itemize}
@@
 \paragraph{New Metadata}
-In order to have a more complete \texttt{Frontmatter} we suggest adding the following information:
+Besides the metadata that Test262 tests currently include, it would be useful for them to have additional information regarding:
 
@@
     \item \texttt{built-ins} - list of all the built-ins used in the test.
-    \item \colorbox{orange}{\texttt{harness-functions}} - list of all the harness functions used to asses the test's results.
-\end{itemize}
-% QUESTION posso adicionar isto?
-acho que era suposto fazer parte da minha tese; se calhar nao faz sentido, mas sim ter a informacao do includes segundo o standar (com o nome dos ficheiros da harness necessarios para o test)
-
-This metadata provides helpful information to solve the problem mentioned before, selecting the applicable tests for partial implementations, by allowing developers to filter tests by \texttt{builtin}, \texttt{static construct} and \texttt{version} of the \texttt{ECMAScript} standard.
-This would provide consistency and standardization to the selection of applicable tests to a partial implementation of the standard.
-As for the \texttt{harness-functions}, it provides the information of about the functions of \texttt{harness} that are used in the test, that is relevant because only a small part of the \texttt{harness} is need in each test even though the \colorbox{orange}{whole library is loaded}.
-% QUESTION suponho que o maior problema seja que todos testes da harness sejam corridos e nao o facto de toda a biblioteca seja carregada
-
-\subsection{An Infrastructure for testing reference implementations of the ECMAScript standard}
-\label{subsec:An Infrastructure for testing reference implementations of the ECMAScript standard}
-
-% Resumo da tese do Miguel 
-This thesis continues the work done in the master thesis of Miguel Trigo titled \emph{Infra-estrutura de Testes para Implementações de Referência do Standard ECMAScript}.
-In Miguel Trigo's master thesis, he \colorbox{orange}{processed} all Test262 tests and checked their metadata,  correcting some errors that \colorbox{orange}{were} found and complementing some metadata that is missing in the \texttt{Frontmatter}.
-The author also added new metadata that out of the scope of the \texttt{Frontmatter}, added statistical data about the tests, and made all the previously mentioned data available in a \texttt{MongoDB} database.
-% QUESTION da' a sensacao que o miguel fez tudo a mao? devia mencionar que ele automou o processo de gerar metadata?
-% QUESTION past tense?
-
-% - Corrigir a metadata dos testes incorrectos e calcular metadata adicional ????
-% - construir uma base de dados com a metadata dos varios testes em formato JSON 
-Miguel Trigo processed the test in Figure \ref{fig:Test262_example} and generated the metadata in Figure \ref{fig:json_metadata}.
-The \texttt{Frontmatter} in the test has the \texttt{es5id} and \texttt{description}, in the generated metadata they are represented for the \texttt{name} and \texttt{value} pairs \texttt{esid} and \texttt{description} respectively \colorbox{orange}{(updating the deprecated es5id)}.
-% QUESTION conta como correcao? eu devo arranjar um exemplo de uma correcao que ele fez? ou so' mensionar?
-%
-%
-The following \texttt{names} were added by Miguel Trigo:
-
-\begin{itemize}
-    \item \texttt{path} - stores the path to the test within \texttt{Test262};
-    \item \texttt{version} - stores the version of the \texttt{ECMAScript} standard that the test corresponds to;
-    \item \texttt{built-ins} - stores a subpart of the \texttt{path};
-    \item \texttt{Array} - stores the subpart of the \texttt{path} that follows \texttt{built-ins};
-    \item \texttt{syntactic\_construct} - stores an array with all the syntactic constructions used in the test;
-    \item \texttt{builtIns} - stores an array with all the builtins used in the test;
-    \item \texttt{asserts} - stores the number of times the \texttt{assert} library is used in the test;
-    \item \texttt{error} - \colorbox{orange}{TODO Nao sei o que representa};
-    \item \texttt{esprima} - stores whether or not the test is supported by \texttt{Esprima}\cite{Esprima} (\texttt{Esprima} is an standard-compliant \texttt{ECMAScript} parser that is also developed in \texttt{ECMAScript});
-    \item \texttt{lines} - stores the number of lines of code written in the test, ignoring the empty lines and comments (in this example the \texttt{value} is 3 because lines 12 and 13 in Figure \ref{fig:Test262_example} are one single line in the actual test in \texttt{Test262} that was split in the example to make the image more legible).
+    \item \texttt{harness-functions} - list of all the harness functions required to run the test.
 \end{itemize}
 
-% Exemplo
+This above metadata critical for filtering the tests when considering partial implementations of the language.
+For instance, if a JS engine only supports the 5th edition of the standard, one must be able to obtain all the corresponding tests for that \texttt{version}.
+Analogously, if a JS engine only implements certain \texttt{built-in} objects, one has to be able to filter out the tests that make use of the \texttt{built-in} objects that it does not implement.
+This would provide consistency and standardization to the selection of applicable tests to a partial implementation of the standard.
+%
+As for the \texttt{harness-functions}, it provides important information about the functions of \texttt{harness} that are used in the test.
+That information is relevant because only a small part of the \texttt{harness} is need in each test even though the whole library is loaded and tested.
+The harness has a total of 7290 lines in 32 files and is tested by 96 tests.
+By only including the exact functions that a test requires, one can speedup the testing process significantly.
+% harness - has 32 files total 7290 lines (comments and blanck lines included)
+% harness tests - has 96 files total 2646 lines (comments and blanck lines included)
+
+\subsection{MetaData262 v0}
+\label{subsec:MetaData262 v0}
+
+% Para1 --> introduzir o trabalho do Miguel Trigo 
+This thesis continues the master thesis of Miguel Trigo\cite{Thesis_Miguel_Trigo}, in which he developed a preliminary version of \texttt{MetaData262}.
+More specifically, he built a MongoDB database storing the metadata of all Test262 tests, representing the metadata of each test as a JSON object.
+
+% Para2 --> estrutura da metadata
+In \texttt{MetaData262 v0}, each test is associated with a JSON object storing the various metadata properties of the test and their corresponding values.
+Those metadata properties being the keywords of the \texttt{Frontmatter} mentioned in the previous section.
+%
+Besides, the existing metadata properties, various new properties were added to \texttt{MetaData262 v0}.
+We can divide these new properties into 3 main groups: location, extra front matter, and statistics.
+Each of these groups will be discussed next.
+
+% Para 2.1 [path] (location)
+\paragraph{Location group}
+The location group contains information about the path to the test inside the \texttt{Test262}, having the property \texttt{path} associated to a string with the full path to the test.
+% (ignored) there are also some properties that are subparts of the path and store their child.
+
+% Para 2.2 [extra front matter: syntactic constructs, built-ins, version] 
+\paragraph{Extra front matter group}
+The extra front matter group contains new metadata generated for \texttt{MetaData262 v0}.
+The new metadata generated is associated with the properties: \texttt{syntactic\_construct}, \texttt{version}, and \texttt{builtIns} that were mentioned in the previous section as important metadata to add.
+\texttt{Version} stores the \texttt{ECMAScript} standard version the test belongs to.
+\texttt{Syntactic\_construct} stores an array with all the syntactic constructions that the test contains.
+As for \texttt{BuiltIns}, it stores all the built-ins that test interacts with and associates with each built-in its fields and functions that are used.
+
+% Para 2.3 [statistics]
+\paragraph{Statistics group}
+Finally, the statistics group contains some statistical data about the tests.
+The properties this group has associated are \texttt{asserts}, \texttt{error}, \texttt{esprima}, and \texttt{lines}.
+The \texttt{asserts} property holds the amount of \texttt{assert} statements in the test.
+The \texttt{error} property stores the amount of calls made to the \texttt{Test262} error functions.
+The \texttt{esprima} property stores \emph{supported} or \emph{not supported} depending on whether \texttt{Esprima}\cite{Esprima} supports the test.
+\texttt{Esprima} is a standard-compliant \texttt{ECMAScript} parser that is also developed in \texttt{ECMAScript}, \texttt{Esprima} fully supports up to version 7 of the standard.
+The property \texttt{lines} holds the number of lines of code the test has, comments as well as empty lines are ignored.
+
+% Para3 --> exemplo da MetaData262 v0 
+Figure \ref{fig:json_metadata} shows the JSON object storing the metadata generated from the Test262 test given in Figure \ref{fig:Test262_example}. 
+% Observacoes sobre o exemplo 
+The JSON object begins with the \texttt{path} property and the path to the test associated with it.
+After, the \texttt{version} property indicates that this test belongs to the fifth version of the standard, if the version is not able to be calculated the field will be omitted.
+Followed by properties in the test's \texttt{Frontmatter} (note that the deprecated \texttt{es5id} is replaced to the \texttt{esid} property).
+% ignorar as propriedades buit-ins e Array
+Next is the \texttt{static\_construct} property that holds an array with the syntactic constructors of the test that were generated.
+The \texttt{builtIns} property holds a JSON object, in which, each property corresponds to a built-in and associated to it are its functions and fields that are interacted with in the test.
+Finally, the statistics group of properties.
+The \texttt{asserts} property that indicates that one assert statement is used in this test, the \texttt{error} holding zero since no error functions of the \texttt{Test262} were used in this test, the \texttt{esprima} property that indicates this test is supported by \texttt{Esprima}, and the \texttt{lines} property indicates that this test has 3 lines of code (the example in Figure \ref{fig:Test262_example} splits the last line of the test in order to improve the readability of the figure hence only 3 lines are counted in the actual test).
+
 \begin{figure}[ht]
@@
 
-The metadata can be separated into four parts.
-Firstly, the \texttt{Frontmatter} part, \texttt{description} and \texttt{esid}, that stores information related to the \texttt{Frontmatter} of the test.
-Secondly, the path part comprised of \texttt{path}, \texttt{built-ins}, and \texttt{Array}, that contains the information related to the path of the test inside the \texttt{Test262}.
-Thirdly, the statistical part formed by \texttt{asserts}, \texttt{error}, \texttt{esprima}, and \texttt{lines}, that provides statistics on the code of the test and the metadata generation process.
+% Para4 --> a metada original em falta teve de ser calculada e os 
+% // bem como os novos campos da metadata -> 
+As part of the setting up of MetaData262 v0,  M.
+Trigo had to compute the missing \texttt{Frontmatter} properties as well as the values of the new properties that he introduced.
+In most cases, the computation was straightforward, only involving a simple syntactic analysis of each test.
+However, the computation of the version and the built-ins used in each test was more involved and we will go through it in more detail below.
+% [...]
+
+% Para5 --> Version Computation 
+\paragraph{Calculation of Version metadata}
+The \texttt{version} of the standard a test belongs to is calculated using 3 different approaches: dynamic, static, and hybrid.
 %
-% - Metadata adicional calculada para cada teste:
-Finally, the new metadata part which consists of \texttt{version}, \texttt{syntactic\_construct}, and \texttt{builtIns}, the new metadata is relevant information about the tests that should be added to the \texttt{Frontmatter}.
-
-% Para cada um dos tipos de metadata 
-In the remaining of the subsection is concerned with how the new metadata, namely \texttt{syntactic\_construct}, \texttt{builtIn}, and \texttt{version}.
-
-% - how syntactic cosntruct is calculated 
-\paragraph{Calculation of syntactic\_Construct metadata}
-The \texttt{syntactic\_construct} metadata is calculated with the help of \texttt{Esprima}, which allows the analysis of the \colorbox{orange}{syntactic tree} of the test.
-The tree analysis of the test is represented in a JSON format and the \texttt{syntactic\_constructs} are the \texttt{values} corresponding to the \texttt{names} \emph{type} of every element of the tree.
-This way, to get all the \texttt{syntactic\_construct} is only a matter of traversing the tree and adding any new one found.
-The main problem being that \texttt{Esprima} only supports till version 7 of the \texttt{ECMAScript} standard, and also the \texttt{negative} tests that \texttt{Esprima} does not support.
-% QUESTION imagem do JSON retornado pelo esprima?
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9\textwidth]{images/test_syntactic_tree.png}
-    \caption{Example test}
-    \label{fig:test_syntactic_tree}
-\end{figure}
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=1\textwidth]{images/JSON_syntactic_tree.png}
-    \caption{syntactic tree of the example test in Figure \ref{fig:test_syntactic_tree}}
-    \label{fig:JSON_syntactic_tree}
-\end{figure}
-
-% - how built-ins is calculated
-\paragraph{Calculation of built-ins metadata}
-The \texttt{built-ins} are calculated with two separate methods.
-The first method makes use of syntactic tree generated from \texttt{Esprima}, to find key \texttt{static\_constructs} to identify the \texttt{built-ins} being used.
-While the second method goes through the code of a test and finds all the functions and associates them with their \texttt{built-ins}, as well as the finding \texttt{built-ins} being used directly.
-In both methods, after all the \texttt{built-ins} of a test are identified, the \texttt{built-in} that corresponds to the most recent version of the \texttt{ECMAScript} standard identifies the version of that test.
-
-% - how version is calculated 
-\paragraph{Calculation of Version metadata}
-The \texttt{version} metadata is calculated by identifying the element, \texttt{built-in}, \texttt{syntactic\_construct} or change in behavior of functions that is part of the newest version of the \texttt{ECMAScript} standard.
-The version is calculated using 3 different approaches dynamic, static, and mixed.
+% dynamic approach
+The dynamic approach is based on the waterfall model, running the tests in various JS engines, each corresponding to a specific version of the standard.
+As Figure \ref{fig:waterfall_model} illustrates, each test starts by running in the JS engine corresponding to version 5 of the \texttt{ECMAScript} standard (represented as ES).
+If the test output is correct then the test belongs to version 5 of the standard, otherwise the test will be run in the engine implementing the next version of the standard.
+This process repeats until the last engine, associated with version 11 of the standard, if the test output is not the expected again then the test is not supported.
 %
-The dynamic approach is based on the waterfall model, running the tests against an implementation of the standard.
-if the test passes then the test is associated with that version of the standard, otherwise the test will be run against the next version of the implementation of the standard.
+% static approach
+The static approach evaluates the syntactic tree of the test generated by \texttt{Esprima}, a widely used JS parser, as a JSON object.
+Starts by finding all the \texttt{JavaScript} keywords in the AST.
+Then associates each JS keyword with the version of the standard it was introduced in.
+Finally, the keyword with the most recent version associated will dictate the test version.
+For example, if the test has only two keywords associated with version 6 and 8, then the test will belong to the version 8 of the standard.
 %
-The static approach uses the syntactic tree generated with \texttt{Esprima}.
-From the analysis of the syntactic tree all the objects, functions, properties, operators, variables, and syntactical constructions present in the test are found, then searches for the oldest version that contains all of them.
-%
-The mixed approach is calculated using the dynamic and static approaches.
-The mixed approach consists of running the results of the dynamic approach of each version and verifying that is does not use features introduced in a posterior version.
-The dynamic approach makes use of \texttt{SpyderMonkey} and \texttt{Node.js} implementations of the standard.
+% hybrid approach
+The hybrid approach is calculated using the results of the dynamic and static approaches.
+The hybrid approach merges the results by maintaining the higher version detected between each approach for each test.
+For instance, if the dynamic approach's result for a test is version 6 and the static's result for the same test is version 8, then, the test with belong to the version 8 of the standard according to the hybrid approach.
 
@@
     \includegraphics[width=1\textwidth]{images/waterfall_model.png}
-    \caption{waterfall model of the dynamic approach}
+    \caption{Waterfall model of the dynamic approach to calculate the ECMAScript version}
     \label{fig:waterfall_model}
 \end{figure}
+
+% Para6 --> Built-in Computation 
+\paragraph{Calculation of built-ins metadata}
+The \texttt{built-ins} used by each test are calculated using two separate approaches: dynamic and static.
+As in the calculation if the \texttt{ECMAScript} version both approaches are combined to improve the results.
+The static approach makes use of syntactic tree generated by \texttt{Esprima} as a JSON object from the test code.
+Next, traverses the AST to find expressions that interact with built-in objects.
+In addiction to the built-in, the fields and functions will also be outputted.
+As for the dynamic approach, it makes use of the built-ins functions in order to log whenever the function is called.
+In essence, a wrapper was created for each function that replaces it, the wrapper first logs that it was called, and then preforms the original method with the arguments that were supplied.
 
@@
 \label{sec:Design and Methodology}
+
+% Para1 --> problemas da metada existente na Metadata262 v0. 
+% --> as versoes estao mal calculadas porque nao existem JS engines 
+% que implementem exactamente uma versao e so essa versao 
+% --> versoes calculadas em instalacoes ad-hoc 
+% -- é preciso melhorar a analise dinamica considerando mais JS engines 
+% --> os built-ins tb estao mal calculados porque a analise dinamica 
+% dos built-ins ignora propriedades -> so considera chamadas a funcoes 
+% por isso, se um teste for apenas: Array.prototype.pop = 3; 
+% nao vai ser identificado como um teste que usa o array pela analise 
+% dinamica 
+% --> ha informacao em falta 
+% ficheiros da harness e funcoes da harness
+% Para 2 --> Problemas de acesso a metadata e obtencao de informacao estatistica 
+% sobre os testes 
+% --> neste momento quem quiser obter informacao estatisca sobre os 
+% testes tem de interagir com uma base de dados Mongo e depois processar a informacao 
+% como entender
+% O objectivo desta tese é resolver estes dois problemas. 
+% 5.1 - Precise Metadata for test262 tests 
+% --> calcular a metadata precisamente 
+% 5.1.1 versao -> 
+% Montar um sistema escalavel para o calculo das versoes 
+% associamos a cada versao um numero arbitrario de docker images cada uma 
+% com um JS engine supostamente dessa versao 
+% ---> parelelizacao etc
+% 5.1.2 built-ins -> melhorar a instrumentcao para o calculo dinamico dos 
+% built-ins --> em particular, apanhar property look-ups dinamicamente
+
+% 5.2 - MetaData Visualisation System 
+% portal para visualizacao da metadata dos testes + informacao estatisca sobre os testes 
+% --> gerar automaticamente 
+% --> print screens do que ja fez
+% --> discussao das bibliotecas que vai considera usar para criacao e display 
+% da informacao estatisca 
+% --> Exemplos de graficos a gerar 
+% -- ...
+
+% OLD DEPRECATED
+\paragraph{OLD DEPRECATED} \colorbox{orange}{Para rescrever tudo com a nova estrutura!}
 
@@
     \includegraphics[width=1\textwidth]{images/website.png}
-    \caption{\colorbox{orange}{Website} for searching the metadata in construction}
-    % QUESTION letra maiuscula na caption
+    \caption{Website for searching the metadata in construction}
     \label{fig:website}
@@
 \colorbox{orange}{pedir descricao da teses de uma das primeiras reunioes (iPad)}
+
+\colorbox{orange}{Qual e' o plano pois nao tenho a certeza do que e' suposto realizar}
 

### 2022-05-22 paper.tex
correct teacher's comments

@@
 % Para2 --> estrutura da metadata
-In \texttt{MetaData262 v0}, each test is associated with a JSON object storing the various metadata properties of the test and their corresponding values.
-Those metadata properties being the keywords of the \texttt{Frontmatter} mentioned in the previous section.
+In \texttt{MetaData262 v0}, each test is associated with a JSON object storing the various metadata properties of the test and their corresponding values, with those metadata properties being the keywords of the \texttt{Frontmatter} mentioned in the previous section.
 %
-Besides, the existing metadata properties, various new properties were added to \texttt{MetaData262 v0}.
-We can divide these new properties into 3 main groups: location, extra front matter, and statistics.
+Besides the existing metadata properties, various new properties were added to \texttt{MetaData262 v0}.
+We can divide these new properties into 3 main groups: location properties, extra front matter properties, and statistics properties.
 Each of these groups will be discussed next.
@@
 \paragraph{Location group}
-The location group contains information about the path to the test inside the \texttt{Test262}, having the property \texttt{path} associated to a string with the full path to the test.
-% (ignored) there are also some properties that are subparts of the path and store their child.
+The location group contains information about the path to the test inside the \texttt{Test262}.
+It consists of the properties \texttt{path}, storing the relative path to the test starting at the root of the \texttt{Test262} project, and splitPath, storing the array obtained by splitting the \texttt{path} string into its corresponding folders.
 
@@
 The extra front matter group contains new metadata generated for \texttt{MetaData262 v0}.
-The new metadata generated is associated with the properties: \texttt{syntactic\_construct}, \texttt{version}, and \texttt{builtIns} that were mentioned in the previous section as important metadata to add.
-\texttt{Version} stores the \texttt{ECMAScript} standard version the test belongs to.
-\texttt{Syntactic\_construct} stores an array with all the syntactic constructions that the test contains.
-As for \texttt{BuiltIns}, it stores all the built-ins that test interacts with and associates with each built-in its fields and functions that are used.
+The new metadata generated is associated with the properties: \texttt{syntactic\_constructs}, \texttt{version}, and \texttt{builtIns} that were mentioned in the previous section as important metadata to add.
+The property \texttt{version} stores the \texttt{ECMAScript} standard version the test belongs to.
+The property \texttt{syntactic\_construct} stores an array with all the syntactic constructions that the test contains.
+As for \texttt{builtIns}, it stores all the built-ins that the test interacts with, mapping each built-in to its fields and methods used by test.
 
@@
 Finally, the statistics group contains some statistical data about the tests.
-The properties this group has associated are \texttt{asserts}, \texttt{error}, \texttt{esprima}, and \texttt{lines}.
+It contains the properties: \texttt{asserts}, \texttt{error}, \texttt{esprima}, and \texttt{lines}.
 The \texttt{asserts} property holds the amount of \texttt{assert} statements in the test.
-The \texttt{error} property stores the amount of calls made to the \texttt{Test262} error functions.
-The \texttt{esprima} property stores \emph{supported} or \emph{not supported} depending on whether \texttt{Esprima}\cite{Esprima} supports the test.
-\texttt{Esprima} is a standard-compliant \texttt{ECMAScript} parser that is also developed in \texttt{ECMAScript}, \texttt{Esprima} fully supports up to version 7 of the standard.
-The property \texttt{lines} holds the number of lines of code the test has, comments as well as empty lines are ignored.
+The \texttt{error} property stores the amount of calls made to \texttt{Test262} error functions.
+The \texttt{esprima} property stores a boolean value indicating whether or not the Esprima\cite{Esprima} parser supports the test.
+\texttt{Esprima} is a standard-compliant \texttt{ECMAScript} parser fully developed in \texttt{ECMAScript}; \texttt{Esprima} fully supports up to version 7 of the standard.
+The property \texttt{lines} holds the number of lines of code the test has, excluding comments and empty lines.
+
+\subsubsection{Example}
 
@@
 % Observacoes sobre o exemplo 
-The JSON object begins with the \texttt{path} property and the path to the test associated with it.
-After, the \texttt{version} property indicates that this test belongs to the fifth version of the standard, if the version is not able to be calculated the field will be omitted.
-Followed by properties in the test's \texttt{Frontmatter} (note that the deprecated \texttt{es5id} is replaced to the \texttt{esid} property).
+The JSON object begins with the \texttt{path} property storing the path to the test at hand.
+After, the \texttt{version} property indicates that this test belongs to the fifth version of the standard.
+If \texttt{MetaData262 v0} was not able to compute the version of the test, this field is omitted.
+Next, we have the properties corresponding to the test's \texttt{Frontmatter} (note that the deprecated \texttt{es5id} is replaced to the \texttt{esid} property).
 % ignorar as propriedades buit-ins e Array
-Next is the \texttt{static\_construct} property that holds an array with the syntactic constructors of the test that were generated.
-The \texttt{builtIns} property holds a JSON object, in which, each property corresponds to a built-in and associated to it are its functions and fields that are interacted with in the test.
-Finally, the statistics group of properties.
-The \texttt{asserts} property that indicates that one assert statement is used in this test, the \texttt{error} holding zero since no error functions of the \texttt{Test262} were used in this test, the \texttt{esprima} property that indicates this test is supported by \texttt{Esprima}, and the \texttt{lines} property indicates that this test has 3 lines of code (the example in Figure \ref{fig:Test262_example} splits the last line of the test in order to improve the readability of the figure hence only 3 lines are counted in the actual test).
+Next we have the \texttt{static\_constructs} property that holds an array with the syntactic constructors of the test.
+The \texttt{builtIns} property holds a JSON object, in which, each property corresponds to a built-in mapped to its functions and fields that are interacted with in the test.
+Finally, we have the statistics group of properties.
+The \texttt{asserts} property indicates that this test uses one assert statement.
+The \texttt{error} holds the numeric value zero since this test uses no error functions of \texttt{Test262}.
+The \texttt{esprima} property indicates this test is supported by \texttt{Esprima}.
+Finally, the \texttt{lines} property indicates that this test has 3 lines of code.
+Note that the example in Figure \ref{fig:Test262_example} splits the last line of the test in order to improve the readability of the figure hence only 3 lines are counted in the actual test).
 
@@
 
+\subsubsection{Metadata Computation}
+
 % Para4 --> a metada original em falta teve de ser calculada e os 
@@
 In most cases, the computation was straightforward, only involving a simple syntactic analysis of each test.
-However, the computation of the version and the built-ins used in each test was more involved and we will go through it in more detail below.
-% [...]
+However, the computation of the version and the built-ins used in each test was more involved.
+We will go through this process in more detail below.
 
@@
 \paragraph{Calculation of Version metadata}
-The \texttt{version} of the standard a test belongs to is calculated using 3 different approaches: dynamic, static, and hybrid.
+The \texttt{version} of the standard a test belongs to is calculated using 3 different approaches: dynamic, static, and hybrid.Below we give a brief overview of each method:
+
+\begin{itemize}
+  \item \emph{Dynamic Approach:}
+The dynamic approach is based on the waterfall model, running the tests in various JS engines, each corresponding to a specific version of the standard.
+As Figure \ref{fig:waterfall_model} illustrates, \texttt{MetaData262 v0} starts by running each test in the JS engine corresponding to version 5 of the \texttt{ECMAScript} standard (represented as ES).
+If the test output is correct then the test is labelled as belonging to version 5 of the standard, otherwise the test will be run in the engine implementing the next version of the standard.
+This process repeats until the last engine, associated with version 11 of the standard.
+If the test output is not the expected again, then the test is not supported.
 %
-% dynamic approach
-The dynamic approach is based on the waterfall model, running the tests in various JS engines, each corresponding to a specific version of the standard.
-As Figure \ref{fig:waterfall_model} illustrates, each test starts by running in the JS engine corresponding to version 5 of the \texttt{ECMAScript} standard (represented as ES).
-If the test output is correct then the test belongs to version 5 of the standard, otherwise the test will be run in the engine implementing the next version of the standard.
-This process repeats until the last engine, associated with version 11 of the standard, if the test output is not the expected again then the test is not supported.
+  \item \emph{Static Approach:}
+The static approach determines the version of a test by analyzing the static keywords of the test.
+Essentially, a test is labeled with the version corresponding to the keyword with the latest associated version; formally:
+\[version(test) = max \{ version(keyword) | keyword \in test \} \]
+For instance, if a test contains two keywords introduced in versions 6 and 8, then the test will belong to the version 8 of the standard.
+\texttt{MetaData262 v0} implements the static approach in the following way.
+It maintains an internal map associating each keyword with the version in which it was introduced.
+Given a test to be labeled, \texttt{MetaData262 v0} first computes the AST of the test using the \texttt{Esprima} parser and then traverses the AST to obtain all the keywords used in the test.
+Finally, it labels the test with the version of the latest keyword (as discussed above).
 %
-% static approach
-The static approach evaluates the syntactic tree of the test generated by \texttt{Esprima}, a widely used JS parser, as a JSON object.
-Starts by finding all the \texttt{JavaScript} keywords in the AST.
-Then associates each JS keyword with the version of the standard it was introduced in.
-Finally, the keyword with the most recent version associated will dictate the test version.
-For example, if the test has only two keywords associated with version 6 and 8, then the test will belong to the version 8 of the standard.
-%
-% hybrid approach
+  \item \emph{Hybrid Approach:}
 The hybrid approach is calculated using the results of the dynamic and static approaches.
 The hybrid approach merges the results by maintaining the higher version detected between each approach for each test.
-For instance, if the dynamic approach's result for a test is version 6 and the static's result for the same test is version 8, then, the test with belong to the version 8 of the standard according to the hybrid approach.
+For instance, if the dynamic approach's result for a test is version 6 and the static's result for the same test is version 8, then, the test is labeled as belonging to version 8 of the standard according to the hybrid approach.
+\end{itemize}
 
@@
 The \texttt{built-ins} used by each test are calculated using two separate approaches: dynamic and static.
-As in the calculation if the \texttt{ECMAScript} version both approaches are combined to improve the results.
-The static approach makes use of syntactic tree generated by \texttt{Esprima} as a JSON object from the test code.
-Next, traverses the AST to find expressions that interact with built-in objects.
-In addiction to the built-in, the fields and functions will also be outputted.
-As for the dynamic approach, it makes use of the built-ins functions in order to log whenever the function is called.
-In essence, a wrapper was created for each function that replaces it, the wrapper first logs that it was called, and then preforms the original method with the arguments that were supplied.
+As in the calculation of the \texttt{ECMAScript} version both approaches are combined to improve the results.
+Below, we describe each approach separately:
+
+\begin{itemize}
+  \item \emph{Static Approach:}
+The static approach maintains an internal map associating each built-in method and property with the name of the corresponding  built-in object (e.g.
+the method hasOwnProperty is associated with the built-in object \texttt{Object}).
+The static approach first computes the AST of the test using \texttt{Esprima} and then traverses the AST checking for each method and property for the built-in object it is associated with.
+In the end, it outputs a JSON object mapping each built-in object to an array with the names of its methods and properties that the test uses.
+%
+  \item \emph{Dynamic Approach:}
+The dynamic approach works by running the actual code of the test and logging the interactions of the test with the JavaScript built-ins.
+To this end, \texttt{MetaData262 v0} wraps the methods of every built-in object in an auxiliary function that calls the corresponding methods and adds an entry to the log registering that call.
+Finally, \texttt{MetaData262 v0} analyses the log information to determine the built-in methods used by the test.
+\end{itemize}
 
@@
 % ficheiros da harness e funcoes da harness
+% data de criação do teste 
 % Para 2 --> Problemas de acesso a metadata e obtencao de informacao estatistica 
@@
 % 5.1 - Precise Metadata for test262 tests 
-% --> calcular a metadata precisamente 
+% --> calcular a metadata precisamente e de forma facilmente reproducible 
 % 5.1.1 versao -> 
@@
 \label{sec:Evaluation and Planning}
-% TODO GANT Diagram
-\colorbox{orange}{pedir descricao da teses de uma das primeiras reunioes (iPad)}
 
-\colorbox{orange}{Qual e' o plano pois nao tenho a certeza do que e' suposto realizar}
+% Evaluation 
+% - Discutimos avaliação das duas componentes do projecto 
+% Componente 1 -> sistema para calculo da metadata
+% É muito difícil de avaliar porque não há ground truth -> nós não sabemos qual é a versão dos testes
+% - vamos seleccionar um conjunto aleatório de testes e analisa-los à mão comparando a versão calculada com a versão que sabemos ser a versão do teste 
+% - vamos tb determinar a versão utilizando à data em que o teste foi inserido na Test262 -> pela data sabemos a versão corrente -> vamos olhar particularmente para os testes em que a versão calculada difere da versão da data dos testes
+% -> Atenção que a versão à data do teste não é necessariamente a versão do teste -> mas dá uma indicação
+% --> testes analisados manualmente são os testes para os quais há grandes discrepâncias entre a versão à data do teste e a versão calculada para o teste
+% Componente 2 -> sistema de visualisacao 
+% -- testes unitários para testar as funcionalidades base da aplicação 
+% -- end-to-end tests -> testes do front-end 
+% Planning -> usar texto equivalente ao Miguel Trigo 
+% - Melhoria do sistema de calculo da metadata - 2 meses
+% - Conclusão do sistema de visualização - 2 meses
+% - Avaliação - 1 mês 
+% - Thesis Writing - 2 meses 
+% Julho - (Agosto férias) - Novembro
 

### 2022-05-23 paper.tex
first draft of the Evaluation and planning

@@
 
-% printscreens do site 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=1\textwidth]{images/website.png}
-    \caption{Website for searching the metadata in construction}
-    \label{fig:website}
-\end{figure}
+% % printscreens do site 
+% \begin{figure}[ht]
+%     \centering
+%     \includegraphics[width=1\textwidth]{images/website.png}
+%     \caption{Website for searching the metadata in construction}
+%     \label{fig:website}
+% \end{figure}
 
@@
 % - Discutimos avaliação das duas componentes do projecto
+
 % Componente 1 -> sistema para calculo da metadata
@@
 % --> testes analisados manualmente são os testes para os quais há grandes discrepâncias entre a versão à data do teste e a versão calculada para o teste
+
 % Componente 2 -> sistema de visualisacao
@@
 % -- end-to-end tests -> testes do front-end
+
+% Para 1 --> components a serem testadas
+This thesis we will make improvements to the metadata calculation of \texttt{MetaData262 v0} and create a visualization system for it.
+The metadata improvements are divided into precision and performance.
+The evaluation of this thesis contributions is described in more detail bellow.
+
+% Para 2 --> Metadata calculation precision  
+\paragraph{Evaluation of metadata calculation precision}
+The precision of the metadata calculated is very hard to access since the is no ground truth available.
+For example the versions of the tests are not available.
+Thus, to evaluate the the metadata calculation we will select a small random group of tests to analyze manually and compare the the version calculated.
+A second evaluation will be made to the tests that have the biggest difference between the version calculated and the version that was the latest at the time the test was added to \texttt{Test262}.
+Firstly, we will find the date the test was added to \texttt{Test262} and associate it with the latest \texttt{ECMAScript} version at that time.
+After, we will select the tests with the biggest discrepancy between the versions.
+Finally we will preform a manual evaluation to ensure that the right version was calculated.
+Note that the version discrepancy is merely indicative as more the once tests were detected as missing and subsequently added to complement \texttt{Test262}.
+
+% Para 3 --> Metadata calculation performance
+\paragraph{Evaluation of metadata calculation performance}
+To evaluate performance improvements made to \texttt{MetaData262 v0} we will compare the time needed to preform the metadata calculation.
+This must be ran in the same system and with the least concurrent processes running in order to make sure we have better results.
+
+% Para 4 --> visualization system
+\paragraph{Evaluation of the visualization system}
+The visualization system evaluation will have three types of tests: unit tests, end-to-end tests and load tests.
+The unit tests will ensure the correctness and reliability of the system.
+The end-to-end tests will guarantee that the website interface works correctly.
+Finally, the load tests will allow us to estimate the amount of users that the systems can support.
+
+\subsection{Planning}
+\label{sub:planning}
+
+% subsection planning (end)
 % Planning -> usar texto equivalente ao Miguel Trigo 
@@
 % Julho - (Agosto férias) - Novembro
+
+The Plan for this thesis is based on the tasks that need to be preformed in order to accomplish the objectives defined in section \ref{sec:Goals}.
+Thus, this thesis has 4 main task to accomplish: improving the metadata calculation, concluding the visualization system, Evaluating the results, and writing the thesis.
+The planning for these tasks is represented in the Gantt diagram in Figure \ref{fig:gantt} and each task is described in more detail afterwards.
+Note that the month of August has a grey background due to it being a vacations.
+
+% gantt tese
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=1\textwidth]{images/gantt_thesis.png}
+    \caption{Gantt diagram with the thesis plan}
+    \label{fig:gantt}
+\end{figure}
+
+% gantt tese tempo dividido por mes
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=1\textwidth]{images/gantt_thesis_month.png}
+    \caption{Gantt diagram with the thesis plan}
+    \label{fig:gantt_month}
+\end{figure}
+
+\paragraph{Improving metadata calculation}
+This task consists of improving \texttt{MetaData262 v0} by improving its precision and performance as discussed in section \ref{sec:Design and Methodology}.
+The precision will be improved with more JS engines and a JS parser that supports more recent versions of the standard.
+As for the performance, we will parallelize the metadata calculation.
+
+\paragraph{Concluding visualization system}
+This task consists of concluding the \texttt{MetaData262 v0} visualization system by adding the possibility to view statistics about the queried tests and improving the system's reliability through preparing it to handle wrong queries.
+
+\paragraph{Evaluation}
+This task consists of evaluating the results of the metadata generated. The most prone to error metadata will be checked by hand in order to guarantee the correctness of the results presented.
+
+\paragraph{Writing the thesis}
+This last task will consist in a report of the execution of the previous tasks. In the report we will present the results obtained as well as the conclusions reached after finishing this project.
 

### 2022-05-24 deprecated_design_methodology.tex
first draft of design and methodology

--- 2022-05-21 paper.tex
@@
-% This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.20 of 2017/10/04
-%
-\documentclass[runningheads]{llncs}
-%
-\usepackage[utf8]{inputenc}
-\usepackage{listings}
-\usepackage{xcolor}
-\usepackage{graphicx}
-\usepackage{tabularx}
-\usepackage{hyperref}
-
-\graphicspath{ {./images/} }
-\hypersetup{
-    colorlinks=true,
-    linkcolor=blue,
-    filecolor=blue,
-    citecolor=blue,
-    urlcolor=blue,
-    linktocpage=true
-}
-\setcounter{tocdepth}{2} %show more in the toc
-
-\newcommand{\kw}[1]{\texttt{#1}}
-\renewcommand{\contentsname}{table of content}
-\usepackage{indentfirst}
-\usepackage[english]{babel}
-
-% If you use the hyperref package, please uncomment the following line
-% to display URLs in blue roman font according to Springer's eBook style:
-\renewcommand\UrlFont{\color{blue}\rmfamily}
-
-%
-
-\title{Infra-estrutura de Testes para Implementações de Referência do Standard ECMAScript}
-\subtitle{}
-
-%
-\titlerunning{Live Metadata for Test262}
-% If the paper title is too long for the running head, you can set
-% an abbreviated paper title here
-%
-\author{Diogo Costa Reis\\ist187526\\
-\email{diogo.costa.reis@tecnico.ulisboa.pt}}
-%
-\authorrunning{Diogo Costa Reis}
-% First names are abbreviated in the running head.
-% If there are more than two authors, 'et al.' is used.
-%
-\institute{Instituto Superior Técnico\\
-Av. Rovisco Pais, 1\\
-1049-001 Lisboa\\
-Tel: +351 218 417 000\\
-\email{mail@tecnico.ulisboa.pt}}
-%
-
-\begin{document}
-
-% a solution to remove title and author from appearing in the table of contents: https://tex.stackexchange.com/a/318220
-{\def\addcontentsline#1#2#3{}\maketitle}
-
-%
-\begin{abstract}
-% TODO
-
-\keywords{ECMAScript \and Specification Language \and Reference Interpreters \and Test262}
-\end{abstract}
-
-\newpage
-
-\tableofcontents
-
-\newpage
-
-\section{Introduction}
-\label{sec:Introduction}
-
-\section{Goals}
-\label{sec:Goals}
-
-\section{Background}
-\label{sec:Background}
-This chapter provides an overview on the ECMAScript standard, the Test262 that are used to test the correct implementation of the ECMAScript standard, and finally an outline of the new metadata generated.
-
-\subsection{ECMAScript}
-\label{subsec:ECMAScript}
-% overview da linguagem JS
-
-% paragrafo - porque e' que javascript e' relevante (uma das linguagens mais usadas no momento)
-JavaScript (JS) is a programming language mainly used in the development of client side web applications, also being one of the most popular programming languages.
-According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as second most active languages on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} as well as on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}
-
-% Boa!
-% Existem muitas implementacoes diferentes da linguagem: client-side (browsers), server-side (Node.js), embedded devices (Jerryscript) -> Estas implementacoes têm de estar de acordo no comportamento observavel -> é particularmente importante na Web -> senao temos sites que em ...
-% -----------------------
-% overview do standard -> descrita num standard
-% Porque é muito importante que as várias implementacoes da linguagem coincidam -> o JavaScript está especificado num documento semi-formal que ...
-% Falar sobre o standard
-%   - o standard está como um interpretador de JavaScript em
-%     pseudo-codigo - descreve detalhadamente os passos que
-%    um interpretador de JS tem de executar ao avaliar qualquer
-%    statement da linguagem
-% Falar sobre o comité -  quem controla a evolucao do ECMAScript
-ECMAScript standard\cite{ECMAScriptStandard} is the official document, written in the English language, in which the JavaScript language is defined.
-This document is in constant evolution, being updated by the ECMA Technical Committee 39 (TC39), which is responsible for maintaining the standard.
-The standard is currently in its twelfth version.
-%
-The standard specifies the \texttt{JavaScript} language, to ensure its multiple compilers and interpreters implementations are coherent.
-Some of the \texttt{JavaScript} compilers are the Hop~\cite{Hop} and the JSC~\cite{JSC} compilers, the most popular interpreters are Node.js~\cite{Node.js} and SpiderMonkey~\cite{SpiderMonkey}.
-These are only four implementations among many others, which come along with the many use cases that \texttt{JavaScript} has.
-%
-\texttt{JavaScript} is mostly used in the web context, both client-side within browsers and server-side, but also in embedded devices.
-Since \texttt{JavaScript} is used in so many scenarios and across so many different contexts, it is highly important that ECMAScript standard is defined in great detail to ensure consistency.
-Browsers, for example, need to run \texttt{JavaScript} implementations that coincide so that websites are correctly rendered and exhibit the same behavior.
-In order to achieve coherent implementations, the standard defines the types, values, objects, properties, syntax, and semantics of \texttt{JavaScript} that must be the same in every \texttt{JavaScript} compiler and interpreter, while allowing \texttt{JavaScript} implementations to define additional types, values, object, properties, and functions.
-
-% Good, more detail if there is time
-% Estrutura do standard
-The \texttt{JavaScript} language can be divided into three major components, those being expressions and commands, built-in libraries, and finally internal functions.
-%
-\begin{itemize}
-\item Expressions and commands describe the behavior of static constructions, detailing the semantics of the diverse expressions (e.g., assignment expressions, built-in operators, etc.), commands (e.g., loop commands, conditions command, etc.), and built-in types (Undefined, Null, Boolean, Number, String and Object).
-%
-\item The internal functions of the language are used to define the semantics for both expressions and commands, as well as the built-in libraries.
-Internal functions are not exposed beyond the internal context of the language.
-In other words, no JavaScript program uses internal functions directly.
-%
-\item Finally, built-in libraries encompass all the internal objects available when a JavaScript program is executed.
-Internal objects expose many functions implemented by the language itself, including functions to manipulate numbers, text, arrays, objects, among other things.
-\end{itemize}
-
-% MetaParagrafo tres tipos
-\colorbox{orange}{The remaining} subsection provides a description of the three types of artifact described in the standard.
-% QUESTION ainda nao foi lido pelo prof
-
-% Expressions and statements
-% Exemplo do standard  e explicacao (if)
-\paragraph{Semantics of IF statement}
-Figure \ref{fig:If-Else Statement} shows a snippet of the ECMAScript standard description of the \texttt{IF} command. In order to evaluate \texttt{IF} commands with the shape:
-
-\begin{center}
-\texttt{if (Expression) Statement1 else Statement2}
-\end{center}
-
-\noindent the language begins by evaluating the \texttt{Expression} storing the result in the variable \texttt{exprRef} (step 1).
-The previous step will be used as Boolean, therefore, the result of the previous step will be converted to a Boolean using the internal functions \texttt{ToBoolean} and \texttt{GetValue}, and having the result stored in the variable \texttt{exprValue} (step 2).A different \texttt{Statement} will be followed depending on \texttt{exprValue}.
-If \texttt{exprValue} has the value \texttt{true} the variable \texttt{stmtCompletion} will have the evaluation of the first \texttt{Statement} (step 3).
-Otherwise, the variable \texttt{stmtCompletion} will store the result of evaluating the second \texttt{Statement} (step 4).
-Finally, a \texttt{Completion} will be returned, if the \texttt{stmtCompletion} has non empty value then it will be returned, however, when the value is empty it will be replaced with undefined (step 5).
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.8\textwidth]{images/if_statement.png}
-    \caption{ECMAScript definition of an if-else statement}
-    \label{fig:If-Else Statement}
-\end{figure}
-
-% arrays são objetos como os outros
-% arrys tem propriedades especiais
-% example of Array.pop
-% Built-ins (Array.pop)
-% printscreen do standard e explicacao
-\paragraph{Semantics of the Pop function}
-The Array built-in is an object as any other in JavaScript.
-The main difference is in its properties.
-Array Objects have a property \texttt{length} that contains the size of the array, as well as a property for each element of the array (from zero to \texttt{length} minus 1).
-
-Figure \ref{fig:Array_pop_example} shows a simplified version of an array performing the pop function, where \texttt{(a)} and \texttt{(b)} are the before and after respectively.
-Before preforming \texttt{pop} \texttt{(a)}, the array has three  properties \texttt{length}, \texttt{0}, and \texttt{1}.
-Property \texttt{length} represents the size of the array that has value \texttt{2}, while the properties \texttt{0} and \texttt{1} store the first (\texttt{banana}) and second (\texttt{kiwi}) elements of the array respectively.
-After \texttt{pop} is preformed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}) and the \texttt{length} property (highlighted in green) is decremented by one since the size of the array changes to one.
-
-% TODO change length with quotation marks
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.4\textwidth]{images/array_pop_example.png}
-    \caption{Example Array.pop}
-    \label{fig:Array_pop_example}
-\end{figure}
-Figure \ref{fig:Array_pop} shows a snippet of the ECMAScript standard description of the pop function in the Array Built-in.
-To begin with, the array will be converted to and Object using the \texttt{ToObject} function, and stored in the \texttt{O} variable (step 1).
-Afterwards, the array length of the previously calculated variable will be calculated with the \texttt{LengthOfArrayLike} internal function, and storing the result in the \texttt{len} variable (step 2).
-At this point there are to ways to proceed depending on the value of \texttt{len}.
-If the value is zero, the Array is empty, then the property \texttt{length} of \texttt{O} is set to zero and \texttt{undefined} is returned (step 3).
-Otherwise, when \texttt{len} is different from zero, meaning that the Array is not empty, the Array's last element will be removed (described in Figure \ref{fig:Array_pop_example}) and returned (step 4).
-To begin with, the language will assert that \texttt{len} is positive (step 4.a).
-Afterwards, the \texttt{newLen} variable will store the value of \texttt{len} decremented by 1 (step 4.b).
-The variable \texttt{index} will store the variable calculated in the previous step represented as a String converted with the \texttt{toString} function (step 4.c).
-Then, stores the value of the \texttt{O} variable at the property corresponding to \texttt{index} in the \texttt{element} variable using the \texttt{Get} function (step 4.d).
-Subsequently, deletes the previously mentioned property of the \texttt{O} variable with the \texttt{DeletePropertyOrThrow} function (step 4.e).
-In addition, sets the \texttt{length} property of the \texttt{O} variable  to the \texttt{newLen} using the \texttt{Set} function (step 4.f).
-Finally, returning the value of the variable \texttt{element} (step 4.g).
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.6\textwidth]{images/array_pop.png}
-    \caption{ECMAScript definition of Array.pop}
-    \label{fig:Array_pop}
-\end{figure}
-
-% Internal Functions (LenghtOfArrayLike)
-% printscreen do standard e explicacao
-\paragraph*{LengthOfArrayLike internal function}
-Figure \ref{fig:LengthOfArrayLike} shows a snippet of the ECMAScript standard description of the \texttt{LengthOfArrayLike} internal function, that evaluates the function:
-
-\begin{center}
-\texttt{LengthOfArrayLike (obj)}
-\end{center}
-
-\noindent The language starts by asserting that \texttt{obj} is an \texttt{Object} (step 1).
-Afterwards, gets the value of the property \texttt{length} from \texttt{obj} using the function \texttt{Get}.
-Then, converts the previously mentioned value to an Integer that represents the length with the \texttt{ToLength} function, and finally returns said Integer (step 2).
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.5\textwidth]{images/length_array_like.png}
-    \caption{ECMAScript definition of the LengthOfArrayLike}
-    \label{fig:LengthOfArrayLike}
-\end{figure}
-
-\subsection{Test262}
-\label{subsec:Test262}
-
-% paragrafo - JS tem muitas particularidades que dificultam o desemvolvimento e testagem  -> É muito dificil desenvolver
-% novas implemenentacoes da linguagem
-% Existe uma bateria de testes que testa as implementacoes da
-% linguagem contra o standard
-% Esta bateria de testes é dificil de manter
-% Porque? - muitos testes, muitas features, em geral ha retrocompatibilidade mas ha um pequeno numero de casos onde a retrocompatibilidade nao se verifica -> os testes de ser modificados
-Implementing a \texttt{JavaScript} engine is particularly difficult since it involves dealing with the many corner cases that exist in the language.
-To test that corner cases are correctly dealt with there is \texttt{Test262}\cite{Test262}, the ECMAScript standard test battery.
-Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity, the total number of tests is around \colorbox{orange}{39837} divided into \colorbox{orange}{87} subfolders, each correspond to roughly one section of the standard.
-\texttt{Test262} complexity grows with changes to the standard since in most cases backward compatible is maintained except for a few select cases.
-% 87 subfolders on built-ins + language
-% QUESTION round number? (40000)
-% QUESTION 105 subfolders if add intl402 (which also has files that were ignore)
-
-% Implementacoes parciais da linguagem 
-% Na acadamia é normal desenvolverem-se implementacoes parciais da linguagem: não suportam a ultima versao, não suportam todos os objectos built-in, não suportam todas
-% Pergunta: Quais é que são os testes apropriados?
-% Normal: Respostas ad-hoc -> sem justificação rigorosa -> basicamente cada paper selecciona os testes que lhe da jeito
-Due to the ECMAScript standard being so extensive most implementations are only partial, especially implementations and analysis developed in academic contexts.
-In order to test partial implementations, one must be able to obtain the applicable set of tests from all the tests contained in Test262.
-Selecting the applicable tests is not a trivial matter because there are too many tests and too many features.
-The current methodology is that each development team manually selects the tests that are applicable to their corresponding implementation.
-This raises the problem that there is not standard and precise way of picking the all the right tests from the almost 40000 tests in \texttt{Test262}, making the possibility of human error when selecting the applicable tests likely.
-
-% gosto
-%% Formato dos testes
-% frontmatter, code
-% exemplo
-Figure \ref{fig:Test262_example} shows a test from \texttt{Test262}.
-Every test of \texttt{Test262} has 3 parts: first is the copyright section represented with the comment \texttt{//} (lines 1 and 2), second is the \texttt{Frontmatter} section between \texttt{/*---} and \texttt{---*/} with some metadata about the test (lines 4 to 7), and finally is the \texttt{Body} section with the code of the test (lines 9 to 13).
-The copyright section has information about the owner and license of the test.
-The \texttt{Frontmatter} section has the test id (15.4.5-1) and a description of the test.
-Finally, the \texttt{Body}'s code tests the correct implementation of the standard.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=1.0\textwidth]{images/test262_array_test.png}
-    \caption{Test262 es5id: 15.4.5-1}
-    \label{fig:Test262_example}
-\end{figure}
-
-%% metadados ja incluidos
-%% referir outra vez o exemplo
-% - metadados oficialmente incluídos nos testes
-%   - que metadados é que os testes contém actualmente
-The \texttt{Frontmatter} has keywords to hold metadata of the test.
-These keywords are associated with specific elements of metadata concerning the test.
-Bellow is the list of possible keywords and their meaning:
-
-\begin{itemize}
-\item \texttt{description} - contains a short description about what will be tested;
-%
-\item \texttt{esid} - contains the hash identifier of the ECMAScript portion associated with the feature that will be tested (the identifier references the most recent version of ECMAScript when the test is created);
-%
-\item \texttt{info} - contains a deeper explanation of the test behavior, frequently includes a direct citation of the standard;
-%
-\item \texttt{negative} - indicates that the test throws an error; associated to the keyword will be the type of error the test is supposed to be thrown (e.g.
-\texttt{TypeError}, \texttt{ReferenceError}) as well as the phase in which the error is expected to be thrown (e.g.
-\texttt{parse} vs \texttt{resolution} vs \texttt{runtime});
-%
-\item \texttt{includes} - contains the list of \texttt{harness} files that should be included in the execution of the tests (\texttt{Test262} makes use of a large number of auxiliary function defined in a dedicated library referred to as the \texttt{Test262} \texttt{harness} described later in this section);
-%
-\item \texttt{author} - contains the identification of the author of the test;
-%
-\item \texttt{flags} - contains a list of booleans for each test property, the properties being: (1) \texttt{onlyStrict}, the test is only executed in strict mode; (2) noStrict, the test will only be executed in mode \emph{sloppy}; (3) \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; (4) \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; (5) \texttt{async}, the test is contains asynchronous functions; (6) \texttt{generated}, the test generates the files specified by the property; (7) \texttt{CanBlockIsFalse} and (8) \texttt{CanBlockIsTrue}, the test will run if the property \texttt{CanBlock} of the \colorbox{orange}{\texttt{Agent Record}} executing it is false and true respectively; (9) \texttt{non-deterministic}, indicates that the semantics used in the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
-% QUESTION clarificar se e' necessario citar o que e' 'Agent Record' uma vez que nao vou explicar
-%
-%
-\item \texttt{features} - contains a list of features that are used in the test;
-%
-\item \texttt{es5id} and \texttt{es6id} - indicates that the feature being tested belongs to ECMAScript 5 and 6 respectively and contains the hash identifier of the section of the standard it belongs to; these keywords have been deprecated and substituted by esid.
-\end{itemize}
-
-% Varias observacoes: 
-%  - A metada esta muitas vezes INCOMPLETA e algumas vezes incorrecta 
-%  - é preciso calcular a metadata certa
-As Figure 5 illustrates, it is often the case that the metadata of a test is incomplete.
-Some tests also have the wrong metadata.
-As part of this thesis, we plan to process all the tests to check and correct their corresponding metadata as well as completing the metadata that is \colorbox{orange}{missing.}
-
-%% QUESTION juntar os paragrafos?
-
-%% metadados que achamos relevantes e nao estao incluidos
-%% ...
-\colorbox{orange}{The example} in Figure \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id.
-Besides the obvious upgrade from \texttt{es5id} to \texttt{esid} it would be useful to have \texttt{includes} with the \texttt{harness} files needed to execute the test.
-The \texttt{harness} information is very useful since it makes it easy to identify the part of the \texttt{harness} needed to run that test, opening the door for loading only part of the \texttt{harness} instead of the whole \texttt{harness} which is the current approach.
-
-% - quais são os metadados que nós achamos serem relevantes e que estão em falta
-\paragraph{New Metadata}
-Besides the metadata that Test262 tests currently include, it would be useful for them to have additional information regarding:
-
-\begin{itemize}
-    \item \texttt{syntactic construct} - list of all syntactic constructions used in the test;
-    \item \texttt{version} - the ECMAScript version of the standard in which the feature being tested was introduced;
-    \item \texttt{built-ins} - list of all the built-ins used in the test.
-    \item \texttt{harness-functions} - list of all the harness functions required to run the test.
-\end{itemize}
-
-This above metadata critical for filtering the tests when considering partial implementations of the language.
-For instance, if a JS engine only supports the 5th edition of the standard, one must be able to obtain all the corresponding tests for that \texttt{version}.
-Analogously, if a JS engine only implements certain \texttt{built-in} objects, one has to be able to filter out the tests that make use of the \texttt{built-in} objects that it does not implement.
-This would provide consistency and standardization to the selection of applicable tests to a partial implementation of the standard.
-%
-As for the \texttt{harness-functions}, it provides important information about the functions of \texttt{harness} that are used in the test.
-That information is relevant because only a small part of the \texttt{harness} is need in each test even though the whole library is loaded and tested.
-The harness has a total of 7290 lines in 32 files and is tested by 96 tests.
-By only including the exact functions that a test requires, one can speedup the testing process significantly.
-% harness - has 32 files total 7290 lines (comments and blanck lines included)
-% harness tests - has 96 files total 2646 lines (comments and blanck lines included)
-
-\subsection{MetaData262 v0}
-\label{subsec:MetaData262 v0}
-
-% Para1 --> introduzir o trabalho do Miguel Trigo 
-This thesis continues the master thesis of Miguel Trigo\cite{Thesis_Miguel_Trigo}, in which he developed a preliminary version of \texttt{MetaData262}.
-More specifically, he built a MongoDB database storing the metadata of all Test262 tests, representing the metadata of each test as a JSON object.
-
-% Para2 --> estrutura da metadata
-In \texttt{MetaData262 v0}, each test is associated with a JSON object storing the various metadata properties of the test and their corresponding values.
-Those metadata properties being the keywords of the \texttt{Frontmatter} mentioned in the previous section.
-%
-Besides, the existing metadata properties, various new properties were added to \texttt{MetaData262 v0}.
-We can divide these new properties into 3 main groups: location, extra front matter, and statistics.
-Each of these groups will be discussed next.
-
-% Para 2.1 [path] (location)
-\paragraph{Location group}
-The location group contains information about the path to the test inside the \texttt{Test262}, having the property \texttt{path} associated to a string with the full path to the test.
-% (ignored) there are also some properties that are subparts of the path and store their child.
-
-% Para 2.2 [extra front matter: syntactic constructs, built-ins, version] 
-\paragraph{Extra front matter group}
-The extra front matter group contains new metadata generated for \texttt{MetaData262 v0}.
-The new metadata generated is associated with the properties: \texttt{syntactic\_construct}, \texttt{version}, and \texttt{builtIns} that were mentioned in the previous section as important metadata to add.
-\texttt{Version} stores the \texttt{ECMAScript} standard version the test belongs to.
-\texttt{Syntactic\_construct} stores an array with all the syntactic constructions that the test contains.
-As for \texttt{BuiltIns}, it stores all the built-ins that test interacts with and associates with each built-in its fields and functions that are used.
-
-% Para 2.3 [statistics]
-\paragraph{Statistics group}
-Finally, the statistics group contains some statistical data about the tests.
-The properties this group has associated are \texttt{asserts}, \texttt{error}, \texttt{esprima}, and \texttt{lines}.
-The \texttt{asserts} property holds the amount of \texttt{assert} statements in the test.
-The \texttt{error} property stores the amount of calls made to the \texttt{Test262} error functions.
-The \texttt{esprima} property stores \emph{supported} or \emph{not supported} depending on whether \texttt{Esprima}\cite{Esprima} supports the test.
-\texttt{Esprima} is a standard-compliant \texttt{ECMAScript} parser that is also developed in \texttt{ECMAScript}, \texttt{Esprima} fully supports up to version 7 of the standard.
-The property \texttt{lines} holds the number of lines of code the test has, comments as well as empty lines are ignored.
-
-% Para3 --> exemplo da MetaData262 v0 
-Figure \ref{fig:json_metadata} shows the JSON object storing the metadata generated from the Test262 test given in Figure \ref{fig:Test262_example}. 
-% Observacoes sobre o exemplo 
-The JSON object begins with the \texttt{path} property and the path to the test associated with it.
-After, the \texttt{version} property indicates that this test belongs to the fifth version of the standard, if the version is not able to be calculated the field will be omitted.
-Followed by properties in the test's \texttt{Frontmatter} (note that the deprecated \texttt{es5id} is replaced to the \texttt{esid} property).
-% ignorar as propriedades buit-ins e Array
-Next is the \texttt{static\_construct} property that holds an array with the syntactic constructors of the test that were generated.
-The \texttt{builtIns} property holds a JSON object, in which, each property corresponds to a built-in and associated to it are its functions and fields that are interacted with in the test.
-Finally, the statistics group of properties.
-The \texttt{asserts} property that indicates that one assert statement is used in this test, the \texttt{error} holding zero since no error functions of the \texttt{Test262} were used in this test, the \texttt{esprima} property that indicates this test is supported by \texttt{Esprima}, and the \texttt{lines} property indicates that this test has 3 lines of code (the example in Figure \ref{fig:Test262_example} splits the last line of the test in order to improve the readability of the figure hence only 3 lines are counted in the actual test).
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.9\textwidth]{images/json_metadata.png}
-    \caption{Metadata generated for test esid: 15.4.5-1.js}
-    \label{fig:json_metadata}
-\end{figure}
-
-% Para4 --> a metada original em falta teve de ser calculada e os 
-% // bem como os novos campos da metadata -> 
-As part of the setting up of MetaData262 v0,  M.
-Trigo had to compute the missing \texttt{Frontmatter} properties as well as the values of the new properties that he introduced.
-In most cases, the computation was straightforward, only involving a simple syntactic analysis of each test.
-However, the computation of the version and the built-ins used in each test was more involved and we will go through it in more detail below.
-% [...]
-
-% Para5 --> Version Computation 
-\paragraph{Calculation of Version metadata}
-The \texttt{version} of the standard a test belongs to is calculated using 3 different approaches: dynamic, static, and hybrid.
-%
-% dynamic approach
-The dynamic approach is based on the waterfall model, running the tests in various JS engines, each corresponding to a specific version of the standard.
-As Figure \ref{fig:waterfall_model} illustrates, each test starts by running in the JS engine corresponding to version 5 of the \texttt{ECMAScript} standard (represented as ES).
-If the test output is correct then the test belongs to version 5 of the standard, otherwise the test will be run in the engine implementing the next version of the standard.
-This process repeats until the last engine, associated with version 11 of the standard, if the test output is not the expected again then the test is not supported.
-%
-% static approach
-The static approach evaluates the syntactic tree of the test generated by \texttt{Esprima}, a widely used JS parser, as a JSON object.
-Starts by finding all the \texttt{JavaScript} keywords in the AST.
-Then associates each JS keyword with the version of the standard it was introduced in.
-Finally, the keyword with the most recent version associated will dictate the test version.
-For example, if the test has only two keywords associated with version 6 and 8, then the test will belong to the version 8 of the standard.
-%
-% hybrid approach
-The hybrid approach is calculated using the results of the dynamic and static approaches.
-The hybrid approach merges the results by maintaining the higher version detected between each approach for each test.
-For instance, if the dynamic approach's result for a test is version 6 and the static's result for the same test is version 8, then, the test with belong to the version 8 of the standard according to the hybrid approach.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=1\textwidth]{images/waterfall_model.png}
-    \caption{Waterfall model of the dynamic approach to calculate the ECMAScript version}
-    \label{fig:waterfall_model}
-\end{figure}
-
-% Para6 --> Built-in Computation 
-\paragraph{Calculation of built-ins metadata}
-The \texttt{built-ins} used by each test are calculated using two separate approaches: dynamic and static.
-As in the calculation if the \texttt{ECMAScript} version both approaches are combined to improve the results.
-The static approach makes use of syntactic tree generated by \texttt{Esprima} as a JSON object from the test code.
-Next, traverses the AST to find expressions that interact with built-in objects.
-In addiction to the built-in, the fields and functions will also be outputted.
-As for the dynamic approach, it makes use of the built-ins functions in order to log whenever the function is called.
-In essence, a wrapper was created for each function that replaces it, the wrapper first logs that it was called, and then preforms the original method with the arguments that were supplied.
-
-\section{Related Work}
-\label{sec:Related Work}
-
-\section{Design and Methodology}
-\label{sec:Design and Methodology}
-
-% Para1 --> problemas da metada existente na Metadata262 v0. 
-% --> as versoes estao mal calculadas porque nao existem JS engines 
-% que implementem exactamente uma versao e so essa versao 
-% --> versoes calculadas em instalacoes ad-hoc 
-% -- é preciso melhorar a analise dinamica considerando mais JS engines 
-% --> os built-ins tb estao mal calculados porque a analise dinamica 
-% dos built-ins ignora propriedades -> so considera chamadas a funcoes 
-% por isso, se um teste for apenas: Array.prototype.pop = 3; 
-% nao vai ser identificado como um teste que usa o array pela analise 
-% dinamica 
-% --> ha informacao em falta 
-% ficheiros da harness e funcoes da harness
-% Para 2 --> Problemas de acesso a metadata e obtencao de informacao estatistica 
-% sobre os testes 
-% --> neste momento quem quiser obter informacao estatisca sobre os 
-% testes tem de interagir com uma base de dados Mongo e depois processar a informacao 
-% como entender
-% O objectivo desta tese é resolver estes dois problemas. 
-% 5.1 - Precise Metadata for test262 tests 
-% --> calcular a metadata precisamente 
-% 5.1.1 versao -> 
-% Montar um sistema escalavel para o calculo das versoes 
-% associamos a cada versao um numero arbitrario de docker images cada uma 
-% com um JS engine supostamente dessa versao 
-% ---> parelelizacao etc
-% 5.1.2 built-ins -> melhorar a instrumentcao para o calculo dinamico dos 
-% built-ins --> em particular, apanhar property look-ups dinamicamente
-
-% 5.2 - MetaData Visualisation System 
-% portal para visualizacao da metadata dos testes + informacao estatisca sobre os testes 
-% --> gerar automaticamente 
-% --> print screens do que ja fez
-% --> discussao das bibliotecas que vai considera usar para criacao e display 
-% da informacao estatisca 
-% --> Exemplos de graficos a gerar 
-% -- ...
-
 % OLD DEPRECATED
@@
 The website for visualizing the metadata aims to make access to the metadata and filtering it easily accessible. The website is planned to allow the search
-
-% printscreens do site 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=1\textwidth]{images/website.png}
-    \caption{Website for searching the metadata in construction}
-    \label{fig:website}
-\end{figure}
-
-\section{Evaluation and Planning}
-\label{sec:Evaluation and Planning}
-% TODO GANT Diagram
-\colorbox{orange}{pedir descricao da teses de uma das primeiras reunioes (iPad)}
-
-\colorbox{orange}{Qual e' o plano pois nao tenho a certeza do que e' suposto realizar}
-
-\section{Conclusion}
-\label{sec:Conclusion}
-
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-
-%
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
-\bibliographystyle{splncs04}
-\bibliography{references}
-
-\end{document}

### 2022-05-24 paper.tex
first draft of design and methodology

@@
 \usepackage{hyperref}
+\usepackage{subcaption}
 
@@
 
-% Para1 --> problemas da metada existente na Metadata262 v0. 
+In this section we will discuss the improvements this thesis will make to \texttt{MetaData262 v0} as well as the MetaData262 visualization system that will be created to improve ease of access to the MetaData262 information.
+
+\subsection{Improvements to MetaData262 v0}
+\label{sub:improvements_to_metadata262_v0}
+
+% Para1 --> problemas do calculo da metada existente na Metadata262 v0. 
 % --> as versoes estao mal calculadas porque nao existem JS engines 
@@
 % data de criação do teste 
-% Para 2 --> Problemas de acesso a metadata e obtencao de informacao estatistica 
-% sobre os testes 
-% --> neste momento quem quiser obter informacao estatisca sobre os 
-% testes tem de interagir com uma base de dados Mongo e depois processar a informacao 
-% como entender
-% O objectivo desta tese é resolver estes dois problemas. 
+The metadata calculated by \texttt{MetaData262 v0} is flawed several ways: in the version calculation, in the built-in calculation, and in the harness calculation.
+These flaws will described in more detail next.
+
+\begin{itemize}
+\item \emph{Version Calculation} - The version calculation is flawed in both the static and dynamic approach, meaning that the hybrid approach that uses the results of the previous two will also be compromised.
+The static approach is flawed since the \texttt{Esprima} parser only supports fully up until \texttt{ECMAScript} version 7.
+The solution to that is to find a JS parser that supports a later version of the standard.
+The dynamic approach's flaw lies in the way it works, since there is no JS engine that fully supports a version of the standard and only that one.
+To improve the dynamic approach we will add more JS engines.
+%
+\item \emph{Built-in Calculation} - The built-in calculation is also flawed in both the static and dynamic approach.
+The static approach has the same flaw since \texttt{Esprima} is the parser used.
+Therefore the same improvement can be made, finding a newer parser.
+As for the dynamic approach is is flawed due to the fact that only function calls are detected, if the properties of a built-in are change directly then it will not be detected.
+For example if the test preforms \texttt{Array.prototype.pop = 3;} then the property prototype of the Array built-in will be changed even though no function call was made and it will go undetected by the dynamic approach.
+%
+\item \emph{Harness Calculation} - The harness calculation very flawed containing only the little information available in the \texttt{Frontmatter} of a few tests.
+This calculation can be improved by completing \texttt{includes} metadata and adding the \texttt{harness-functions} metadata.
+Both are mentioned in earlier sections, the \texttt{includes} stores the files of the \texttt{harness} needed to run the test and the \texttt{harness-functions} will store the functions from those files that are actually used.
+\end{itemize}
+
+% Para 2 --> Scalability/Performance and Replicability of the calculation
 % 5.1 - Precise Metadata for test262 tests 
@@
 % built-ins --> em particular, apanhar property look-ups dinamicamente
+Improving the calculations of the \texttt{MetaData} will enhance the issue of performance, we will improve the performance with the use of \texttt{docker}\cite{docker} containers and this way we will also improve the ease to replicate the calculations.
+The duration to calculate the version with the dynamic approach scales linearly with the amount of engines per standard version being used.
+In \texttt{MetaData262 v0} only two JS engines are being used and the dynamic approach duration takes multiple hours to run.
+With the use of \texttt{docker} containers we will be able to make the parallelize the work of the dynamic approaches through multiple docker images.
+Containers running in parallel will not only improve the performance but also allow MetaData262 to be replicated in an easier way since the is less requirements and the replication will be standardized.
+
+\subsection{MetaData262 Visualization System}
+\label{sub:metadata262_visualization_system}
+
+% --> Problemas de acesso a metadata e obtencao de informacao estatistica 
+% sobre os testes 
+% --> neste momento quem quiser obter informacao estatisca sobre os 
+% testes tem de interagir com uma base de dados Mongo e depois processar a informacao 
+% como entender
+% O objectivo desta tese é resolver estes dois problemas.
+
+% Para 0 --> Motivar a criacao do system de visualizacao
+The \texttt{MetaData262 Visualization System} appears to ease the barrier to access the \texttt{MetaData262 v0} information and statistics.
+Right now the metadata is hosted in a MongoDB database leading to the need to have knowledge of MongoDB to get the information.
+After getting the information there would most likely be a need to process that data and maybe even generating charts to provide statistics of the data.
+This is visualization system will make this process much easier.
 
@@
 
-% OLD DEPRECATED
-\paragraph{OLD DEPRECATED} \colorbox{orange}{Para rescrever tudo com a nova estrutura!}
+% Para 1 --> O que o website vai mostrar e o que para isso e' necessario fazer
+The \texttt{MetaData262 Visualization System} is in its core a web application that allows users visualize the metadata calculated easily and updates it self whenever there are changes in \texttt{Test 262}.
+Users are will to filter the metadata for the built-ins and version of the test.
+Users will also be able visualize the queried information in different ways: \texttt{PATH}, \texttt{JSON}, and \texttt{STATS}.
+The \texttt{PATH} will only show the path to the tests in \texttt{Test262}.
+The \texttt{JSON} will allow the users to see the full metadata calculated for the tests.
+Finally, the \texttt{STATS} will show some statistics about the tests queried.
+To ensure that the information is up to date this system needs to update it self since this is a thesis project and there is no active developer team.
+In order to accomplish that, the system will need check for changes, be it in the form of new tests or changes to existing ones, and calculate the new metadata.
 
-%  grandes objectivos 
-This thesis is a continuation of Miguel Tringo's master thesis, aiming to complete the calculation of the \texttt{Frontmatter} metadata especially the \colorbox{orange}{metadata about the \texttt{harness}}, also aims to improve the generated metadata.
-Another aim of this thesis is to create a website for visualizing the metadata generated.
-\colorbox{orange}{Finally,} we aim to build a platform that allows users to submit the markdown generated from executing the \texttt{Test262} and compare different runs.
-% QUESTION campo features? (restantes talvez possam ser ignorados author, negative, flags, info and description)
-% QUESTION incluir?
+% Para 2 --> O que ja esta feito
+The development of the web application has already began, in \colorbox{orange}{Figure} \ref{fig:website} it is possible to see the progress already made.
+The image has 2 main parts: the filters part, and the information part.
+%
+The upper part is the filters part, where it is possible to select the filters for querying the metadata.
+The first filter, \texttt{BuiltIn Belongs} allows to select tests from the subfolders of the built-in folder in the \texttt{Test262}.
+The \texttt{BuiltIn Contained} enables the user to filter tests by the built-ins the tests can use.
+The \texttt{BuiltIn Belongs} and \texttt{BuiltIn Contained} filters are combined using \texttt{AND} or \texttt{OR} relations.
+The \texttt{AND} relation will only return tests that satisfy both filters.
+As for the \texttt{OR} relation it will return every test that satisfies either of the filters.
+After we have the \texttt{Version} filter that allows filtering the tests by the \texttt{ECMAScript} version.
+Next we have the \texttt{BACKEND SEARCH} and \texttt{LOCAL SEARCH} search buttons that will query the metadata database for the tests that satisfy the filters set.
+The \texttt{BACKEND SEARCH} filter the tests in the backend server while the \texttt{LOCAL SEARCH} it will filter the tests in the machine that is used to access the website.
+%
+In the example the filters selected are \texttt{Object}  and \texttt{Array} for the \texttt{BuiltIn Belong} and \texttt{BuiltIn Contained} respectively with an \texttt{AND} relation and the sixth version is also selected.
+%
+Finally, the information part of the image contains the result of the query.
+The information can already be displayed in two of the three ways mentioned above, in the image the \texttt{PATH} is selected.
+Afterwards the image shows it took 16 milliseconds to have the query results available to the user.
+Next the image displays that a total of 32 tests satisfy the filters selected.
+Finally, the image shows the path of the tests tests selected.
+Note that not all tests are shown at the same time, in order to improve performance only 10 tests are shown at a time, at the bottom of the image is the pagination menu.
+% QUESTION mudar o titulo do website para MetaData262 Visualization System ??
 
-% - processamento completo da metadata oficial 
-% -> o processamento do miguel esta incompleto -> nao incluir informacao informacao sobre -> ... 
-% [pode criticar como entender]
-Miguel Trigo's thesis metadata is incomplete in various ways, tests with unknown \texttt{version} around 9000, tests without \texttt{built-ins} around 17000, and tests without \texttt{syntactic\_construct} around 13000 \colorbox{orange}{tests without esid around 600}.
-% QUESTION sao tests da versao seginte de ECMA que ainda nao tem seccao na versao actual?
-%
-The metadata from the thesis could be improved in way the data is arranged, for example, the subfolders of the path to the test are spread into the JSON Object of the test.
-The subfolders information being put into an array of ordered subfolders would increase the readability.
+% printscreens do site 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=1\textwidth]{images/website.png}
+    \caption{Website for searching the metadata in construction}
+    \label{fig:website}
+\end{figure}
 
-% - calculo mais preciso da nova metadata proposta
-% - o calculo da versao pode ser melhorado de varias maneiras
-This thesis plans to improve the \texttt{version} metadata generation in two dimensions precision and efficiency.
-% - Precisao 
-% - usar mais engines
-For the precision, with more \texttt{JavaScript} engines being used, there would be more certainty when determining the version.
-% - Eficiencia 
-% - paralelizar os engines 
-As for the efficiency, it would be possible to parallelize the waterfall model of the dynamic approach running multiple tests at the same time.
-It is also possible to use \texttt{git diff} to identify the tests that were added or changed since the last time the dynamic approach was executed, \colorbox{orange}{ only needing to execute} the dynamic approach on the differences.
-% QUESTION e' problema se fizerem update noutro lado que afeta o teste indiretamente?
+% Para 3 --> Trade off the fazer a filtragem local or no servidor
+In Figure \ref{fig:website} there are two search buttons because there at the time trade offs being made with either way of searching, the following results were run in the laptop serving the web application.
+The backend search starts getting slow whenever the amount of tests being returned surpasses 2500 in the local machine, taking on roughly half a second to get the results.
+The local search has the problem that it takes around 4 seconds to get all the metadata from the server at page load.
+The local search has the advantage of filtering the tests in a few milliseconds.
+Caching the metadata in the users browser would allow the metadata to be only loaded one, making the usability of the web application much better.
 
-% - Sistema de visualisacao da metadata 
-The website for visualizing the metadata aims to make access to the metadata and filtering it easily accessible. The website is planned to allow the search
+% Para 4 --> Os graficos que queremos gerar --> bibliotecas
+This web application will allow the user to view charts about the metadata of the tests queried.
+M.
+Trigo's thesis he generated the charts shown in Figure \ref{fig:grafico_bigodes} and Figure \ref{fig:grafico_stackBar} as well as some other pie charts and bar charts.
+In the web application those charts will be generated in real time with the help of external libraries.
+Especially the charts below, the box and whiskers chart and the stacked bar chart, are not supported by many of the most popular libraries.
+Two of the libraries that are capable of generating all the charts needed are the \texttt{KendoReact}\cite{KendoReact} an UI library and \texttt{Victory}\cite{Victory} a chart library.
+\texttt{KendoReact} is a full UI library, charts being only part of its 126 components.
+Whereas, the \texttt{Victory} library has only 34 components and all are directly related with generating charts.
+The Victory library also seams simpler to use, the community around it seams more active which would be important in case of difficulty while implementing the charts needed.
 
-% % printscreens do site 
-% \begin{figure}[ht]
-%     \centering
-%     \includegraphics[width=1\textwidth]{images/website.png}
-%     \caption{Website for searching the metadata in construction}
-%     \label{fig:website}
-% \end{figure}
+\begin{figure}
+\centering
+\begin{minipage}{.5\textwidth}
+  \centering
+  \includegraphics[width=0.95\textwidth]{images/n_asserts.png}
+  \captionof{figure}{M. Trigo chart number of asserts by category}
+  \label{fig:grafico_bigodes}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+  \centering
+  \includegraphics[width=0.95\textwidth]{images/version.png}
+  \captionof{figure}{M. Trigo chart versions by category}
+  \label{fig:grafico_stackBar}
+\end{minipage}
+\end{figure}
+
+% Para 5 --> breve descricao do processo para manter a metadata atualizada
+% ignorado por estar com 3 paginas so' sobre o website
 
@@
 \paragraph{Concluding visualization system}
-This task consists of concluding the \texttt{MetaData262 v0} visualization system by adding the possibility to view statistics about the queried tests and improving the system's reliability through preparing it to handle wrong queries.
+This task consists of concluding \texttt{MetaData262 v0} visualization system by adding the possibility to view statistics about the queried tests and improving the system's reliability through preparing it to handle wrong queries.
 

### 2022-05-25 paper.tex
solve easy comments

@@
 
-In this section we will discuss the improvements this thesis will make to \texttt{MetaData262 v0} as well as the MetaData262 visualization system that will be created to improve ease of access to the MetaData262 information.
+We recall that this thesis has two goals.
+Firstly, we want to improve \texttt{MetaData262 v0} in two ways: add support for missing metadata elements, and improve the infrastructure for the metadata computation, making the obtained metadata more reliable and the process faster and reproducible.
+Secondly, we want to create a visualisation system for easy selection of tests and display of test statistics.
 
@@
 % data de criação do teste 
-The metadata calculated by \texttt{MetaData262 v0} is flawed several ways: in the version calculation, in the built-in calculation, and in the harness calculation.
-These flaws will described in more detail next.
+The metadata calculated by \texttt{MetaData262 v0} is flawed in several ways: in the version calculation, in the built-in calculation, and in the harness calculation.
+We describe these flaws and how to fix them in more detail next.
 
 \begin{itemize}
-\item \emph{Version Calculation} - The version calculation is flawed in both the static and dynamic approach, meaning that the hybrid approach that uses the results of the previous two will also be compromised.
-The static approach is flawed since the \texttt{Esprima} parser only supports fully up until \texttt{ECMAScript} version 7.
-The solution to that is to find a JS parser that supports a later version of the standard.
+\item \emph{Version Calculation} - The version calculation is flawed in both the static and dynamic approachs, meaning that the hybrid approach that uses the results of the previous two will also be compromised.
+The static approach is flawed since the \texttt{Esprima} parser only fully supports up to \texttt{ECMAScript} version 7.
+Improvements can be made by finding a \texttt{JavaScript} parser that supports a later version of the standard.
 The dynamic approach's flaw lies in the way it works, since there is no JS engine that fully supports a version of the standard and only that one.
 To improve the dynamic approach we will add more JS engines.
+By adding more data points to the calculation, we will increase the reliability of the obtained metadata.
 %
-\item \emph{Built-in Calculation} - The built-in calculation is also flawed in both the static and dynamic approach.
-The static approach has the same flaw since \texttt{Esprima} is the parser used.
-Therefore the same improvement can be made, finding a newer parser.
-As for the dynamic approach is is flawed due to the fact that only function calls are detected, if the properties of a built-in are change directly then it will not be detected.
-For example if the test preforms \texttt{Array.prototype.pop = 3;} then the property prototype of the Array built-in will be changed even though no function call was made and it will go undetected by the dynamic approach.
+\item \emph{Built-in Calculation} - The built-in calculation is also flawed in both the static and dynamic approachs.
+The static approach has the same flaw since it also uses \texttt{Esprima} as parser.
+Therefore the same improvement can be made by finding a newer parser.
+As for the dynamic approach, it is flawed due to the fact that only function calls are logged and therefore detected.
+If the properties of a built-in are changed directly then it will not be detected.
+For example if a test preforms \texttt{Array.prototype.pop = 3;} then the property \texttt{prototype} of the Array built-in will be changed even though no function call was made.
+This can not be detected by the current dynamic approach.
 %
-\item \emph{Harness Calculation} - The harness calculation very flawed containing only the little information available in the \texttt{Frontmatter} of a few tests.
-This calculation can be improved by completing \texttt{includes} metadata and adding the \texttt{harness-functions} metadata.
-Both are mentioned in earlier sections, the \texttt{includes} stores the files of the \texttt{harness} needed to run the test and the \texttt{harness-functions} will store the functions from those files that are actually used.
+\item \emph{Harness Calculation} - The \texttt{harness} calculation is not done, including only the little information available in the \texttt{Frontmatter} of a few tests.
+This calculation can be improved by completing \texttt{includes} metadata and adding an additional \texttt{harness-functions} metadata.
+Both \texttt{includes} and \texttt{harness-functions} are mentioned in earlier sections, the first stores the files of the \texttt{harness} needed to run the test and the second will store the functions from those files that are actually used in the test.
 \end{itemize}
@@
 % built-ins --> em particular, apanhar property look-ups dinamicamente
-Improving the calculations of the \texttt{MetaData} will enhance the issue of performance, we will improve the performance with the use of \texttt{docker}\cite{docker} containers and this way we will also improve the ease to replicate the calculations.
+The improvements the \texttt{MetaData} calculation precision, in particular the version dynamic approach, will enhance the issue of performance.
+This issue will be improved with the use of \texttt{docker}\cite{docker} containers and this way we will also improve the ease to replicate the calculations of the \texttt{MetaData}.
+%
 The duration to calculate the version with the dynamic approach scales linearly with the amount of engines per standard version being used.
-In \texttt{MetaData262 v0} only two JS engines are being used and the dynamic approach duration takes multiple hours to run.
-With the use of \texttt{docker} containers we will be able to make the parallelize the work of the dynamic approaches through multiple docker images.
-Containers running in parallel will not only improve the performance but also allow MetaData262 to be replicated in an easier way since the is less requirements and the replication will be standardized.
+In \texttt{MetaData262 v0} only two JS engines are being used and the dynamic approach already takes multiple hours to run.
+%
+With the use of \texttt{docker} containers we will be able to parallelize the work of the dynamic approaches by multiplying the docker images that run the JS engines.
+Containers running in parallel will not only improve the performance but also allow MetaData262 to be replicated in an easier way since it requires a smaller configuration effort to replicate the \texttt{MetaData} calculation.
+% TODO rescrever comentarios do prof
 
@@
 Right now the metadata is hosted in a MongoDB database leading to the need to have knowledge of MongoDB to get the information.
-After getting the information there would most likely be a need to process that data and maybe even generating charts to provide statistics of the data.
+After getting the information there would most likely be a need to process that metadata and maybe even generating charts to provide statistics.
 This is visualization system will make this process much easier.
@@
 % Para 1 --> O que o website vai mostrar e o que para isso e' necessario fazer
-The \texttt{MetaData262 Visualization System} is in its core a web application that allows users visualize the metadata calculated easily and updates it self whenever there are changes in \texttt{Test 262}.
-Users are will to filter the metadata for the built-ins and version of the test.
+The \texttt{MetaData262 Visualization System} is in its core a web application that allows users visualize and filter the \texttt{Test262} metadata easily.
+The system will also update itself whenever there are changes in \texttt{Test262}.
+Users are able to filter the metadata by the built-ins and version of the test.
 Users will also be able visualize the queried information in different ways: \texttt{PATH}, \texttt{JSON}, and \texttt{STATS}.
@@
 Finally, the \texttt{STATS} will show some statistics about the tests queried.
-To ensure that the information is up to date this system needs to update it self since this is a thesis project and there is no active developer team.
+To ensure that the information is up to date, this system needs to update itself since this is a thesis project and there is no active developer team.
 In order to accomplish that, the system will need check for changes, be it in the form of new tests or changes to existing ones, and calculate the new metadata.
@@
 % Para 2 --> O que ja esta feito
-The development of the web application has already began, in \colorbox{orange}{Figure} \ref{fig:website} it is possible to see the progress already made.
+The development of the web application has already began, \colorbox{orange}{Figure} \ref{fig:website} shows the progress already made.
 The image has 2 main parts: the filters part, and the information part.
@@
 The \texttt{BuiltIn Contained} enables the user to filter tests by the built-ins the tests can use.
+%
 The \texttt{BuiltIn Belongs} and \texttt{BuiltIn Contained} filters are combined using \texttt{AND} or \texttt{OR} relations.
 The \texttt{AND} relation will only return tests that satisfy both filters.
-As for the \texttt{OR} relation it will return every test that satisfies either of the filters.
-After we have the \texttt{Version} filter that allows filtering the tests by the \texttt{ECMAScript} version.
-Next we have the \texttt{BACKEND SEARCH} and \texttt{LOCAL SEARCH} search buttons that will query the metadata database for the tests that satisfy the filters set.
+As for the \texttt{OR} relation, it will return every test that satisfies either of the filters.
+%
+Then, we have the \texttt{Version} filter that allows filtering the tests by the \texttt{ECMAScript} version.
+Next, we have the \texttt{BACKEND SEARCH} and \texttt{LOCAL SEARCH} search buttons that will query the metadata database for the tests that satisfy the filters set.
 The \texttt{BACKEND SEARCH} filter the tests in the backend server while the \texttt{LOCAL SEARCH} it will filter the tests in the machine that is used to access the website.
 %
-In the example the filters selected are \texttt{Object}  and \texttt{Array} for the \texttt{BuiltIn Belong} and \texttt{BuiltIn Contained} respectively with an \texttt{AND} relation and the sixth version is also selected.
+In the example, the filters selected are \texttt{Object}  and \texttt{Array} for the \texttt{BuiltIn Belong} and \texttt{BuiltIn Contained} respectively with an \texttt{AND} relation and the sixth version is also selected.
 %
-Finally, the information part of the image contains the result of the query.
+The information part of the image contains the result of the query.
 The information can already be displayed in two of the three ways mentioned above, in the image the \texttt{PATH} is selected.
-Afterwards the image shows it took 16 milliseconds to have the query results available to the user.
-Next the image displays that a total of 32 tests satisfy the filters selected.
-Finally, the image shows the path of the tests tests selected.
+The image also shows it took 16 milliseconds to have the query results available to the user.
+The image displays that a total of 32 tests satisfy the filters selected.
+The image shows the path of the tests selected.
 Note that not all tests are shown at the same time, in order to improve performance only 10 tests are shown at a time, at the bottom of the image is the pagination menu.
@@
 
+% TODO rescrever tudo com comentarios do prof
 % Para 3 --> Trade off the fazer a filtragem local or no servidor
@@
 
+% TODO rescrever tudo com comentarios do prof
 % Para 4 --> Os graficos que queremos gerar --> bibliotecas

### 2022-05-25 paper.tex
improvements in the Evalution and planning section

@@
 
-% Para 1 --> components a serem testadas
-This thesis we will make improvements to the metadata calculation of \texttt{MetaData262 v0} and create a visualization system for it.
-The metadata improvements are divided into precision and performance.
-The evaluation of this thesis contributions is described in more detail bellow.
+% Para 1 --> components a serem testados
+The evaluation of this thesis will focus in its two main elemtents: the metadata calculation infrastructure and the \texttt{Metadata262 Visualization System}.
+The metadata calculation infrastructure will be evaluated for its precision and performance.
+As for the \texttt{Metadata262 Visualization System}, it will be evaluated for its reliability.
+Each createria of the evatuation will be described in more detail bellow.
 
@@
 The precision of the metadata calculated is very hard to access since the is no ground truth available.
-For example the versions of the tests are not available.
-Thus, to evaluate the the metadata calculation we will select a small random group of tests to analyze manually and compare the the version calculated.
+The version metadata calculated is being calculated for the first time.
+Thus, to evaluate the presition of the metadata calculation we will select a small random group of tests to analyze manually.
 A second evaluation will be made to the tests that have the biggest difference between the version calculated and the version that was the latest at the time the test was added to \texttt{Test262}.
-Firstly, we will find the date the test was added to \texttt{Test262} and associate it with the latest \texttt{ECMAScript} version at that time.
-After, we will select the tests with the biggest discrepancy between the versions.
+Firstly, we will find the date the tests were added to \texttt{Test262} and associate them with the latest \texttt{ECMAScript} version at that time.
+Afterwards, we will select the tests with the biggest discrepancy between the version assotiated and the version calculated.
 Finally we will preform a manual evaluation to ensure that the right version was calculated.
-Note that the version discrepancy is merely indicative as more the once tests were detected as missing and subsequently added to complement \texttt{Test262}.
+Note that the version discrepancy is merely indicative, more the once tests were detected missing and subsequently added to complement \texttt{Test262}.
+These tests were added when they were detected missing and not when the version was released.
 
@@
 \paragraph{Evaluation of metadata calculation performance}
-To evaluate performance improvements made to \texttt{MetaData262 v0} we will compare the time needed to preform the metadata calculation.
-This must be ran in the same system and with the least concurrent processes running in order to make sure we have better results.
+To evaluate performance improvements made to \texttt{MetaData262 v0} we will compare the time needed to preform the metadata calculation between the old and new infrastructures.
+This must be ran in the same system and with the least concurrent processes running in order to make sure we have the least interference, this way achieving the most trustworthy results.
 
 % Para 4 --> visualization system
-\paragraph{Evaluation of the visualization system}
+\paragraph{Evaluation of the visualization system reliability}
 The visualization system evaluation will have three types of tests: unit tests, end-to-end tests and load tests.
-The unit tests will ensure the correctness and reliability of the system.
+The unit tests will ensure the correctness of the system's results.
 The end-to-end tests will guarantee that the website interface works correctly.
-Finally, the load tests will allow us to estimate the amount of users that the systems can support.
+Finally, the load tests will allow us to estimate the amount of users that the systems can support, this way allowig us to estimate the resorces need.
 
@@
 The Plan for this thesis is based on the tasks that need to be preformed in order to accomplish the objectives defined in section \ref{sec:Goals}.
-Thus, this thesis has 4 main task to accomplish: improving the metadata calculation, concluding the visualization system, Evaluating the results, and writing the thesis.
-The planning for these tasks is represented in the Gantt diagram in Figure \ref{fig:gantt} and each task is described in more detail afterwards.
-Note that the month of August has a grey background due to it being a vacations.
+Thus, this thesis has 4 main task to accomplish: improving the metadata calculation, concluding the visualization system, evaluating the results, and writing the thesis.
+These tasks are represented in the Gantt diagram in Figure \ref{fig:gantt} and each task is described in more detail afterwards.
+Note that the month of August has a grey background due to it being a vacations period.
 
@@
 The precision will be improved with more JS engines and a JS parser that supports more recent versions of the standard.
-As for the performance, we will parallelize the metadata calculation.
+As for the performance, we will parallelize the metadata calculation using \texttt{docker} containers.
 
 \paragraph{Concluding visualization system}
-This task consists of concluding \texttt{MetaData262 v0} visualization system by adding the possibility to view statistics about the queried tests and improving the system's reliability through preparing it to handle wrong queries.
+This task consists of concluding \texttt{MetaData262 v0} visualization system by adding the possibility to view statistics about the queried tests and improving the system's reliability.
 
 \paragraph{Evaluation}
-This task consists of evaluating the results of the metadata generated. The most prone to error metadata will be checked by hand in order to guarantee the correctness of the results presented.
+This task consists of evaluating the results of the metadata generated.
+The most prone to error metadata calculation will be checked by hand in order to guarantee the correctness of the results presented.
+The development of the tests for the visualization system and the system itself will overlap durin the first two weeks of the evaluation.
 
 \paragraph{Writing the thesis}
-This last task will consist in a report of the execution of the previous tasks. In the report we will present the results obtained as well as the conclusions reached after finishing this project.
+This last task will consist of reporting the execution of the previous tasks, by presenting the results obtained as well as the conclusions reached after finishing this project.
 

### 2022-05-26 paper.tex
added the abstract, concution and fixed the missing bibliografy

@@
 \begin{abstract}
-% TODO
+The ECMAScript standard is the specification for JavaScript which is one of the most popular programming languages in the world.
+Implementing the standard is challenging in various ways, both due to its complexity and its scale.
+Test262 was created to alleviate the implementations difficulty by providing a conformance test suite that allows developers to test their implementations.
+There is the project MetaData262 v0 that calculates the metadata of the Test262 test suite.
+With this thesis, we will improve the MetaData262 v0 metadata calculations as well as build infrastructure to view the metadata through a web interface.
+The web interface will also allow to filter tests by some characteristics such as built-ins and version of the standard as well as display statistics of the metadata.
 
@@
 
-% Boa!
 % Existem muitas implementacoes diferentes da linguagem: client-side (browsers), server-side (Node.js), embedded devices (Jerryscript) -> Estas implementacoes têm de estar de acordo no comportamento observavel -> é particularmente importante na Web -> senao temos sites que em ...
@@
 % Falar sobre o comité -  quem controla a evolucao do ECMAScript
-ECMAScript standard\cite{ECMAScriptStandard} is the official document, written in the English language, in which the JavaScript language is defined.
+ECMAScript standard\cite{ECMAScript} is the official document, written in the English language, in which the JavaScript language is defined.
 This document is in constant evolution, being updated by the ECMA Technical Committee 39 (TC39), which is responsible for maintaining the standard.
@@
 
-% Good, more detail if there is time
 % Estrutura do standard
@@
 \paragraph{Semantics of IF statement}
-Figure \ref{fig:If-Else Statement} shows a snippet of the ECMAScript standard description of the \texttt{IF} command. In order to evaluate \texttt{IF} commands with the shape:
+Figure \ref{fig:If-Else_Statement} shows a snippet of the ECMAScript standard description of the \texttt{IF} command. In order to evaluate \texttt{IF} commands with the shape:
 
@@
     \caption{ECMAScript definition of an if-else statement}
-    \label{fig:If-Else Statement}
+    \label{fig:If-Else_Statement}
 \end{figure}
@@
 
+Estudos científicos sobre análise de programas e técnicas de instrumentação para o JavaScript são variados e cobrem um grande leque de especificações como: sistemas de tipos (\cite{Type_Analysis_for_JavaScript}, \cite{Towards_Type_Inference_for_JavaScript}, \cite{Understanding_TypeScript},  \cite{Fast_and_precise_hybrid_type_inference_for_JavaScript}, \cite{Depedent_Types_for_JavaScript}, \cite{Safe_&_Efficient_Gradual_Typing_for_TypeScript},  \cite{Type_Inference_for_Static_Compilation_of_JavaScript}), interpretadores abstratos (\cite{SAFE:Formal_Specification_and_Implementation_of_a_Scalable_Analysis_Framework_for_ECMAScript}, \cite{JSAI:a_static_analysis_platform_for_JavaScript}, \cite{Scalable_and_Precise_Static_Analysis_of_JavaScript_Applications_via_Loop-Sensitivity},  \cite{Combining_string_abstract_domains_for_JavaScript_analysis:an_evaluation)}), análises \emph{points-to} (\cite{Points-to_analysis_for_JavaScript}), lógicas de Hoare (\cite{Javert-2017}, \cite{Javert-2019}), semântica operacional (\cite{JSCert-2014}, \cite{KJS-2015}, \cite{JSExplain-2018}), representações intermédias e compiladores (\cite{Dynamic_property_caches:a_step_towards_faster_JavaScript_proxy_objects}).
+Em seguida, apresentamos os trabalhos relacionados mais relevantes para o projeto tese a ser desenvolvido, organizando-os numa linha temporal dividida nas Tabelas \ref{tab:related-work_timeline-1} e \ref{tab:related-work_timeline-2}.
+Esta linha temporal foca-se nos marcos mais importantes no trabalho de investigação focado na formalização da linguagem JavaScript desde a sua criação em 1995.
+
+% https://tex.stackexchange.com/a/196808
+\begin{table}[h!]
+\vspace{-2em}
+\renewcommand{\arraystretch}{1.8}
+
+\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
+  1995 & O JavaScript é lançado e implementado no Netscape Navigator 2.0 beta 3.\\
+  2005 & Thiemann apresenta uma primeira tentativa de definição de um sistema de tipos para JavaScript ~\cite{Thiemann-2005}.
+O sistema rastreia as possíveis características de um objeto e sinaliza conversões de tipos suspeitas.
+O trabalho realizado não cobre aspetos vitais do JavaScript, tal como herança baseada em protótipos, e não foi acompanhado por uma implementação na altura da sua publicação.\\
+  2008 & Maffeis et al.
+propõe uma primeira semântica operacional para a linguagem ECMAScript versão 3.
+(ES3)~\cite{operational_semantics_javascript-2008}.
+Esta semântica foi usada para analisar várias propriedades de segurança de aplicações web e mashups.\\
+  2010 & Guha et al.~\cite{essence_javascript-2010} definem a linguagem $\lambda$JS, um cálculo formal que  captura as características mais fundamentais do ES3.
+$\lambda$JS vem com um interpretador \emph{Racket}~\cite{Racket} e um tradutor de programas ES3 em expressões $\lambda$JS.\\
+  2012 & Gardner et al.~\cite{Gardner-2012} adapta ideias da lógica de separação para fornecer uma lógica de Hoare escalável para um subconjunto de JavaScript, com base em uma semântica operacional fiel ao ES3.\\
+  2012 & Politz et al.~\cite{Politz-2012} extende a linguagem $\lambda$JS do ES3 para o ES5 incluindo, pela primeira vez, uma semântica formal da JavaScript \texttt{property descriptors} e tratamento completo da declaração \texttt{eval}.\\
+  2014 & Bodin et al.
+apresenta o \emph{JSCert}~\cite{JSCert-2014}, uma formalização do standard ES5 escrito em \emph{Coq}~\cite{Coq}, e \emph{JSRef} um interpretador executável de referências extraido do \emph{Coq} para o \emph{OCaml}~\cite{OCaml}, provado correto em relação ao \emph{JSCert} e testado usando os testes da bateria Test262~\cite{Test262}.\\
+  2015 & Park et al.
+apresenta \emph{KJS}~\cite{KJS-2015}, a semântica formal do ES5 mais completa até à data.
+O \emph{KJS} foi desenvolvido usando a framework \emph{K}~\cite{Kframework}, um sistema consolidado de re-escrita de termos que suporta vários tipos de análise simbólica baseados em \emph{reachability logic} (All-Path Reachability Logic \cite{All-Path_Reachability_Logic}).
+\emph{KJS} foi testada cuidadosamente usando a bateria de testes, passando em todos os 2.782 testes do core da linguagem.
+O \emph{KJS} pode ser executado simbolicamente, podendo ser usado para análise formal e verificação de programas ES5.\\
+  2015 & Gardner et al.
+\cite{JSCert_1Year-2015}  estendem o projecto \emph{JSCert} para suportar \emph{arrays} ES5.
+Para tal, os autores ligam o \emph{JSRef} à implementação da biblioteca Array do motor V8 \cite{V8} da Google.
+Os autores adicionalmente avaliam o estado atual do projeto \emph{JSCert}, fornecendo uma análise detalhada da metodologia como um todo e uma análise detalhada dos testes que o projeto passa/falha.
+\end{tabular}
+\vspace{1em}
+\caption{Linha temporal que contém os trabalhos relacionados mais relevantes para os objetivos do projeto (1 of 2)}
+\label{tab:related-work_timeline-1}
+\end{table}
+\vspace{-2em}
+
+\begin{table}[h!]
+\renewcommand{\arraystretch}{1.8}
+\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
+
+  2017 & Fragoso Santos et al.
+cria a ferramenta \emph{JaVert}~\cite{Javert-2017}: uma ferramenta de verificação JavaScript que permite raciocínio semi-automático sobre as propriedades funcionais de correcção dos programas ES5.
+JaVert não assume qualquer simplificação da semântica da linguagem, analisando todos os \emph{corner-cases} da linguagem.
+A carga de anotações necessária é substancial.\\
+  2018 & Charguéraud et al.
+apresenta \emph{JSExplain}~\cite{JSExplain-2018}, um interpretador de referências para a linguagem ES5 que segue as especificações e que produz traços de execução inspecionáveis.
+Também inclui uma interface de inspeção de código que permite ao programador executar passo a passo o seu programa ES5 mas também o próprio código do interpretador ES5, enquanto executa os programas ES5.
+\\
+  2020 & Sampaio et al.
+~\cite{Javert-2020} estendem a ferramenta JaVerT para dar suporte ao raciocínio automático sobre eventos em JavaScript.
+Para tal, os autores desenvolvem diretamente em JavaScript uma implementação de referência de JavaScript promises.
+Os autores usam-na para testar a sua implementação de referência, passando 344 testes do total de 474 testes dedicados a JavaScript promises.
+Cerca de 106 testes foram excluídos devido a features ES6 que ainda não são suportadas e a testes que requerem o modo \emph{non-strict}.\\
+  2020 & Jihyeok Park et al.
+apresenta o JISET~\cite{JISET-2020} a primeira ferramenta que sintetiza automaticamente o analizador e que traduz AST-IR diretamente de uma especificação da linguagem dada.
+\end{tabular}
+\vspace{1em}
+\caption{Linha temporal que contém os trabalhos relacionados mais relevantes para os objetivos do projeto (2 of 2)}
+\label{tab:related-work_timeline-2}
+\end{table}
+
 \section{Design and Methodology}
@@
 Firstly, we want to improve \texttt{MetaData262 v0} in two ways: add support for missing metadata elements, and improve the infrastructure for the metadata computation, making the obtained metadata more reliable and the process faster and reproducible.
-Secondly, we want to create a visualisation system for easy selection of tests and display of test statistics.
+Secondly, we want to create a visualization system for easy selection of tests and display of test statistics.
 
@@
 \begin{itemize}
-\item \emph{Version Calculation} - The version calculation is flawed in both the static and dynamic approachs, meaning that the hybrid approach that uses the results of the previous two will also be compromised.
+\item \emph{Version Calculation} - The version calculation is flawed in both the static and dynamic approaches, meaning that the hybrid approach that uses the results of the previous two will also be compromised.
 The static approach is flawed since the \texttt{Esprima} parser only fully supports up to \texttt{ECMAScript} version 7.
@@
 %
-\item \emph{Built-in Calculation} - The built-in calculation is also flawed in both the static and dynamic approachs.
+\item \emph{Built-in Calculation} - The built-in calculation is also flawed in both the static and dynamic approaches.
 The static approach has the same flaw since it also uses \texttt{Esprima} as parser.
@@
 % Para 1 --> components a serem testados
-The evaluation of this thesis will focus in its two main elemtents: the metadata calculation infrastructure and the \texttt{Metadata262 Visualization System}.
+The evaluation of this thesis will focus in its two main elements: the metadata calculation infrastructure and the \texttt{Metadata262 Visualization System}.
 The metadata calculation infrastructure will be evaluated for its precision and performance.
 As for the \texttt{Metadata262 Visualization System}, it will be evaluated for its reliability.
-Each createria of the evatuation will be described in more detail bellow.
+Each criteria of the evaluation will be described in more detail bellow.
 
@@
 The version metadata calculated is being calculated for the first time.
-Thus, to evaluate the presition of the metadata calculation we will select a small random group of tests to analyze manually.
+Thus, to evaluate the precision of the metadata calculation we will select a small random group of tests to analyze manually.
 A second evaluation will be made to the tests that have the biggest difference between the version calculated and the version that was the latest at the time the test was added to \texttt{Test262}.
 Firstly, we will find the date the tests were added to \texttt{Test262} and associate them with the latest \texttt{ECMAScript} version at that time.
-Afterwards, we will select the tests with the biggest discrepancy between the version assotiated and the version calculated.
+Afterwards, we will select the tests with the biggest discrepancy between the version associated and the version calculated.
 Finally we will preform a manual evaluation to ensure that the right version was calculated.
@@
 The end-to-end tests will guarantee that the website interface works correctly.
-Finally, the load tests will allow us to estimate the amount of users that the systems can support, this way allowig us to estimate the resorces need.
+Finally, the load tests will allow us to estimate the amount of users that the systems can support, this way allowing us to estimate the resources need.
 
@@
 The most prone to error metadata calculation will be checked by hand in order to guarantee the correctness of the results presented.
-The development of the tests for the visualization system and the system itself will overlap durin the first two weeks of the evaluation.
+The development of the tests for the visualization system and the system itself will overlap during the first two weeks of the evaluation.
 
@@
 \label{sec:Conclusion}
+With this thesis we will build an infrastructure that enables convenient access to metadata about the \texttt{Test262} conformance suite.
+We will improve upon the metadata calculation of \texttt{MetaData262 v0} as well as making it accessible to users via a web interface.
+This will enable \texttt{JavaScript} engine developers to filter tests easily, in particular developers of partial implementations.
+This will allow developers to find and correct errors quicker and will also enable a standardization of the selection of applicable tests for partial implementations.
+
+This report starts with a description of the background to this thesis, in particular the \texttt{ECMAScript} standard, the \texttt{Test262} conformance test suite and \texttt{MetaData262 v0}.
+Then, we present the scientific work done in the scope of the \texttt{JavaScript} language, with studies from 2005 to 2020.
+Then we have the design and methodology with a detailed description of the contributions this thesis will make.
+Finally, we have the methods we will use to evaluate the quality of the work done.
 

### 2022-05-26 paper.tex
first draft related work

--- 2022-05-25 paper.tex
@@
 
+Scientific studies about analysis of programs and techniques of instrumentation of JavaScript encompasses a varied range of specifications such as: type systems (\cite{Type_Analysis_for_JavaScript}, \cite{Towards_Type_Inference_for_JavaScript}, \cite{Understanding_TypeScript},  \cite{Fast_and_precise_hybrid_type_inference_for_JavaScript}, \cite{Depedent_Types_for_JavaScript}, \cite{Safe_&_Efficient_Gradual_Typing_for_TypeScript},  \cite{Type_Inference_for_Static_Compilation_of_JavaScript}), abstract interpreters (\cite{SAFE:Formal_Specification_and_Implementation_of_a_Scalable_Analysis_Framework_for_ECMAScript}, \cite{JSAI:a_static_analysis_platform_for_JavaScript}, \cite{Scalable_and_Precise_Static_Analysis_of_JavaScript_Applications_via_Loop-Sensitivity},  \cite{Combining_string_abstract_domains_for_JavaScript_analysis:an_evaluation)}), \emph{points-to} analysis (\cite{Points-to_analysis_for_JavaScript}), Hoare logic (\cite{Javert-2017}, \cite{Javert-2019}), operational semantics (\cite{JSCert-2014}, \cite{KJS-2015}, \cite{JSExplain-2018}), intermediate representations and compilers (\cite{Dynamic_property_caches:a_step_towards_faster_JavaScript_proxy_objects}).
+Next, we have the related work  most relevant for this thesis, it is organized into two Tables \ref{tab:related-work_timeline-1} and \ref{tab:related-work_timeline-2}.
+This time line focuses on the most important investigation work, focused on the JavaScript language specification since its creation in 1995.
+
+% https://tex.stackexchange.com/a/196808
+\begin{table}[h!]
+\vspace{-2em}
+\renewcommand{\arraystretch}{1.8}
+
+\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
+  1995 & JavaScript is launched and implemented in Netscape Navigator 2.0 beta 3.\\
+  2005 & Thiemann presents the first try at defining a system of types possible for JavaScript ~\cite{Thiemann-2005}.
+The system detects the possible characteristics of an object and signals suspected type conversions.
+The work done does not cover important aspects of JavaScript, such as inheritance based on prototypes, and was not created along with an implementation at the time of the publication.\\
+  2008 & Maffeis et al.
+proposes a first operational semantics for the language ECMAScript version 3.
+(ES3)~\cite{operational_semantics_javascript-2008}.
+This first semantics was used to analyze various properties of security of web applications.\\
+  2010 & Guha et al.~\cite{essence_javascript-2010} defines a language $\lambda$JS, a formal calculation that encapsulates the fundamental characteristics of ES3.
+$\lambda$JS comes along an interpreter \emph{Racket}~\cite{Racket} and a translator from ES3 programs to expressions $\lambda$JS.\\
+  2012 & Gardner et al.~\cite{Gardner-2012} adapts the logic ideas of separation to build a scalable Hoare logic for a subset of JavaScript, based of semantics faithful to ES3.\\
+\end{tabular}
+\vspace{1em}
+\caption{Time line that contains the most relevant related work to the objectives of the project (1 of 2)}
+\label{tab:related-work_timeline-1}
+\end{table}
+\vspace{-2em}
+
+\begin{table}[ht!]
+\renewcommand{\arraystretch}{1.8}
+\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
+
+  2012 & Politz et al.~\cite{Politz-2012} extends the language $\lambda$JS of ES3 to ES5 included, fir the first time, a formal semantics of JavaScript \texttt{property descriptors} and complete handling of the \texttt{eval} declaration.\\
+  2014 & Bodin et al.
+presents \emph{JSCert}~\cite{JSCert-2014}, a formalization of the ES5 standard written in \emph{Coq}~\cite{Coq}, and \emph{JSRef} an executable interpreter of references extracted from \emph{Coq} to \emph{OCaml}~\cite{OCaml}, proving correct when related to \emph{JSCert} and tested using the test suite Test262~\cite{Test262}.\\
+  2015 & Park et al.
+presents \emph{KJS}~\cite{KJS-2015}, a formal semantics for ES5, the most complete to that moment.
+The \emph{KJS}  was developed using the framework \emph{K}~\cite{Kframework}, a consolidated system of rewriting of terminology that supports various types of symbolic analyses based on \emph{reachability logic} (All-Path Reachability Logic \cite{All-Path_Reachability_Logic}).
+\emph{KJS} was tested carefully using the test suite, passing all the 2.782 tests of the language core.
+The \emph{KJS} can be executed symbolically, being able to be used for formal analyses and verification of ES5 programs.\\
+  2015 & Gardner et al.
+\cite{JSCert_1Year-2015} extend the project \emph{JSCert} to support \emph{arrays} ES5.
+In order to do this, the authors connected the \emph{JSRef} to the implementation of the Array library of the V8 engine \cite{V8} from Google.
+Additionally, the authors evaluated the state of the project \emph{JSCert} at the time, giving a detailed analysis of the methodology as a whole and a detailed analysis of the tests that the project passed or failed.\\
+  2017 & Fragoso Santos et al.
+creates the tool \emph{JaVert}~\cite{Javert-2017}: a tool for verifying JavaScript that allows semiautomatic reasoning about the fundamental properties of correctness of ES5 programs.
+JaVert does not assume any simplification of the language semantics, analyzing all the \emph{corner-cases} of the language.
+The load of necessary notation is substantial.\\
+  2018 & Charguéraud et al.
+presents \emph{JSExplain}~\cite{JSExplain-2018}, an reference interpreter for the ES5 language that follows the specifications and produces an inspectable execution traces.
+It also includes an interface of code inspection that allows the programmer to execute his ES5 code step by step as well as the code of the ES5 interpreter, while executing ES5 programs.\\
+  2020 & Sampaio et al.
+~\cite{Javert-2020} extend the JaVerT tool to support automatic reasoning about JavaScript events.
+In order to do this, the authors developed a reference implementation of JavaScript promises directly in JavaScript.
+The authors used \colorbox{orange}{it} to test their reference implementation, passing 344 tests of the 474 total dedicated to JavaScript promises.
+Around 106 testes were excluded due to ES6 features that were not yet supported and to tests that require \emph{non-strict} mode.\\
+  2020 & Jihyeok Park et al. presents JISET~\cite{JISET-2020} the first tool that synthesizes automatically the analyzer and that translates AST-IR directly from a specification of a given language.
+\end{tabular}
+\vspace{1em}
+\caption{Time line that contains the most relevant related work to the objectives of the project (2 of 2)}
+\label{tab:related-work_timeline-2}
+\end{table}
+% QUESTION it e' Test262??
+
 \section{Design and Methodology}

### 2022-05-27 paper.tex
fix miguel trigo reference; fix abstract draft goals

--- 2022-05-25 paper.tex
@@
 \begin{abstract}
-% TODO
+The ECMAScript standard is the specification for JavaScript which is one of the most popular programming languages in the world.
+Implementing the standard is challenging in various ways, both due to its complexity and its scale.
+Test262 was created to alleviate the implementations difficulty by providing a conformance test suite that allows developers to test their implementations.
+Test262 has many tests and its hard to select the tests that belong to or need a feature in order to work, MetaData262 v0 was created to help with that.
+MetaData262 v0 calculates metadata of the tests in the Test262 test suite, making it available in MongoDB database.
+With this thesis, we will improve the MetaData262 v0 metadata calculations as well as build infrastructure to view the metadata through a web interface.
+The web interface will also allow to filter tests by some characteristics such as built-ins and version of the standard as well as display statistics of the metadata.
 
@@
 
+This thesis has two main goals , the first is to improve \texttt{MetaData262 v0} and the second is to create a web visualization system to make the metadata more accessible.
+This way building the \texttt{MetaData262 Visualization System} that will allow its users to easily search the \texttt{Test262} metadata  filtering it by its characteristics.
+
+\paragraph{MetaData262 v0 improvements}
+This goal aims to improve \texttt{MetaData262 v0} in the precision and performance while making it more complete and scalable and easier to replicate.
+
+\paragraph{MetaData262 Visualization System}
+This goals aims to create a web platform to easily visualize the metadata calculated by \texttt{MetaData262 v0}.
+The system will provide filters with which a user will be able to filter the test by its characteristics.
+Then, the system will provide the filtered tests in various formats as well as useful statistics about the tests selected.
+
 \section{Background}
@@
 % Para1 --> introduzir o trabalho do Miguel Trigo 
-This thesis continues the master thesis of Miguel Trigo\cite{Thesis_Miguel_Trigo}, in which he developed a preliminary version of \texttt{MetaData262}.
+This thesis continues the master thesis of Miguel Trigo\cite{MetaDataV0}, in which he developed a preliminary version of \texttt{MetaData262}.
 More specifically, he built a MongoDB database storing the metadata of all Test262 tests, representing the metadata of each test as a JSON object.
@@
 
+% Scientific studies about analysis of programs and techniques of instrumentation of JavaScript encompasses a varied range of specifications such as: type systems (\cite{Type_Analysis_for_JavaScript}, \cite{Towards_Type_Inference_for_JavaScript}, \cite{Understanding_TypeScript},  \cite{Fast_and_precise_hybrid_type_inference_for_JavaScript}, \cite{Depedent_Types_for_JavaScript}, \cite{Safe_&_Efficient_Gradual_Typing_for_TypeScript},  \cite{Type_Inference_for_Static_Compilation_of_JavaScript}), abstract interpreters (\cite{SAFE:Formal_Specification_and_Implementation_of_a_Scalable_Analysis_Framework_for_ECMAScript}, \cite{JSAI:a_static_analysis_platform_for_JavaScript}, \cite{Scalable_and_Precise_Static_Analysis_of_JavaScript_Applications_via_Loop-Sensitivity},  \cite{Combining_string_abstract_domains_for_JavaScript_analysis:an_evaluation)}), \emph{points-to} analysis (\cite{Points-to_analysis_for_JavaScript}), Hoare logic (\cite{Javert-2017}, \cite{Javert-2019}), operational semantics (\cite{JSCert-2014}, \cite{KJS-2015}, \cite{JSExplain-2018}), intermediate representations and compilers (\cite{Dynamic_property_caches:a_step_towards_faster_JavaScript_proxy_objects}).
+Next, we have the related work  most relevant for this thesis, it is organized into two Tables \ref{tab:related-work_timeline-1} and \ref{tab:related-work_timeline-2}.
+This time line focuses on the most important investigation work, focused on the JavaScript language specification since its creation in 1995.
+
+% % https://tex.stackexchange.com/a/196808
+% \begin{table}[h!]
+% \vspace{-2em}
+% \renewcommand{\arraystretch}{1.8}
+
+% \begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
+%   1995 & JavaScript is launched and implemented in Netscape Navigator 2.0 beta 3.\\
+%   2005 & Thiemann presents the first try at defining a system of types possible for JavaScript ~\cite{Thiemann-2005}.
+The system detects the possible characteristics of an object and signals suspected type conversions.
+The work done does not cover important aspects of JavaScript, such as inheritance based on prototypes, and was not created along with an implementation at the time of the publication.\\
+%   2008 & Maffeis et al.
+proposes a first operational semantics for the language ECMAScript version 3.
+(ES3)~\cite{operational_semantics_javascript-2008}.
+This first semantics was used to analyze various properties of security of web applications.\\
+%   2010 & Guha et al.~\cite{essence_javascript-2010} defines a language $\lambda$JS, a formal calculation that encapsulates the fundamental characteristics of ES3.
+$\lambda$JS comes along an interpreter \emph{Racket}~\cite{Racket} and a translator from ES3 programs to expressions $\lambda$JS.\\
+%   2012 & Gardner et al.~\cite{Gardner-2012} adapts the logic ideas of separation to build a scalable Hoare logic for a subset of JavaScript, based of semantics faithful to ES3.\\
+% \end{tabular}
+% \vspace{1em}
+% \caption{Time line that contains the most relevant related work to the objectives of the project (1 of 2)}
+% \label{tab:related-work_timeline-1}
+% \end{table}
+% \vspace{-2em}
+
+% \begin{table}[ht!]
+% \renewcommand{\arraystretch}{1.8}
+% \begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
+
+%   2012 & Politz et al.~\cite{Politz-2012} extends the language $\lambda$JS of ES3 to ES5 included, fir the first time, a formal semantics of JavaScript \texttt{property descriptors} and complete handling of the \texttt{eval} declaration.\\
+%   2014 & Bodin et al.
+presents \emph{JSCert}~\cite{JSCert-2014}, a formalization of the ES5 standard written in \emph{Coq}~\cite{Coq}, and \emph{JSRef} an executable interpreter of references extracted from \emph{Coq} to \emph{OCaml}~\cite{OCaml}, proving correct when related to \emph{JSCert} and tested using the test suite Test262~\cite{Test262}.\\
+%   2015 & Park et al.
+presents \emph{KJS}~\cite{KJS-2015}, a formal semantics for ES5, the most complete to that moment.
+The \emph{KJS}  was developed using the framework \emph{K}~\cite{Kframework}, a consolidated system of rewriting of terminology that supports various types of symbolic analyses based on \emph{reachability logic} (All-Path Reachability Logic \cite{All-Path_Reachability_Logic}).
+\emph{KJS} was tested carefully using the test suite, passing all the 2.782 tests of the language core.
+The \emph{KJS} can be executed symbolically, being able to be used for formal analyses and verification of ES5 programs.\\
+%   2015 & Gardner et al.
+\cite{JSCert_1Year-2015} extend the project \emph{JSCert} to support \emph{arrays} ES5.
+In order to do this, the authors connected the \emph{JSRef} to the implementation of the Array library of the V8 engine \cite{V8} from Google.
+Additionally, the authors evaluated the state of the project \emph{JSCert} at the time, giving a detailed analysis of the methodology as a whole and a detailed analysis of the tests that the project passed or failed.\\
+%   2017 & Fragoso Santos et al.
+creates the tool \emph{JaVert}~\cite{Javert-2017}: a tool for verifying JavaScript that allows semiautomatic reasoning about the fundamental properties of correctness of ES5 programs.
+JaVert does not assume any simplification of the language semantics, analyzing all the \emph{corner-cases} of the language.
+The load of necessary notation is substantial.\\
+%   2018 & Charguéraud et al.
+presents \emph{JSExplain}~\cite{JSExplain-2018}, an reference interpreter for the ES5 language that follows the specifications and produces an inspectable execution traces.
+It also includes an interface of code inspection that allows the programmer to execute his ES5 code step by step as well as the code of the ES5 interpreter, while executing ES5 programs.\\
+%   2020 & Sampaio et al.
+~\cite{Javert-2020} extend the JaVerT tool to support automatic reasoning about JavaScript events.
+In order to do this, the authors developed a reference implementation of JavaScript promises directly in JavaScript.
+The authors used \colorbox{orange}{it} to test their reference implementation, passing 344 tests of the 474 total dedicated to JavaScript promises.
+Around 106 testes were excluded due to ES6 features that were not yet supported and to tests that require \emph{non-strict} mode.\\
+%   2020 & Jihyeok Park et al. presents JISET~\cite{JISET-2020} the first tool that synthesizes automatically the analyzer and that translates AST-IR directly from a specification of a given language.
+% \end{tabular}
+% \vspace{1em}
+% \caption{Time line that contains the most relevant related work to the objectives of the project (2 of 2)}
+% \label{tab:related-work_timeline-2}
+% \end{table}
+% % QUESTION it e' Test262??
+
 \section{Design and Methodology}

### 2022-05-27 paper.tex
draft introduction before meeting with the teacher

--- 2022-05-26 paper.tex
@@
 Test262 was created to alleviate the implementations difficulty by providing a conformance test suite that allows developers to test their implementations.
-There is the project MetaData262 v0 that calculates the metadata of the Test262 test suite.
+Test262 has many tests and its hard to select the tests that belong to or need a feature in order to work, MetaData262 v0 was created to help with that.
+MetaData262 v0 calculates metadata of the tests in the Test262 test suite, making it available in MongoDB database.
 With this thesis, we will improve the MetaData262 v0 metadata calculations as well as build infrastructure to view the metadata through a web interface.
@@
 
+% paragrafo - porque e' que javascript e' relevante (uma das linguagens mais usadas no momento)
+JavaScript (JS) is a programming language mainly used in the development of client side web applications, also being one of the most popular programming languages.
+According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as second most active languages on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} as well as on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}
+
+% Para 2 --> implemetar JS e' dificil: complexidade tamanho
+% Para 2.5 --> Test262 ajuda a implementar permitindo testar implementacoes --> test262 tambem e' grande dificil selecionar os testes apropriados
+JavaScript is defined in the \texttt{ECMAScript} standard, a document written in English.
+The standard has be growing with each version, both in size and complexity.
+The growth of the standard is illustrated in Figure \ref{fig:paginas_standard_version}.
+This makes the JavaScript a language very hard to implement with many corner cases.
+\texttt{Test262} was created to aid the development of JavaScript engines, by providing a conformance test suite that developer can use to test their implementations.
+As the standard is extensive and complex and its conformance test suite needs to thoroughly test the standard implementation, then, the test suite it self becomes very extensive.
+The standard being complex and extensive also leads to many of its implementations being only partial.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.7\textwidth]{images/standard_pages_evolution.png}
+    \caption{Evolution of ECMAScript standard's number of pages}
+    \label{fig:paginas_standard_version}
+\end{figure}
+
+% Para 4 --> MetaData v0 calcula metadata --> disponibiliza mongoDB
+
+When implementing a JavaScript engine it is very important to know the tests that the engine should pass for the feature that is being implemented.
+The selection of applicable tests for partial implementations is particularly difficult since tests for one feature may or may not use other features that are not implemented.
+\texttt{MetaData262 v0} appears to solve this issue, it processes the \texttt{Test262} suite and calculates metadata about the tests.
+\texttt{MetaData262 v0} has some particularly interesting properties such as a version of the standard that the test belongs to, the syntactic constructs that are used in the test, and the built-ins that are used by the test.
+The metadata of \texttt{MetaData262 v0} is made available in the form of a MongoDB database.
+
+% --> eu vou melhorar metadata
+% Para 5 --> Visualization system permite interagir com a metadata via web com interface simples --> filtrar; verdados varias maneiras; estatisticas
+In this thesis we will improve \texttt{MetaData262 v0} in its calculations of exiting metadata as well as adding the new metadata related with the harness functions used.
+We will also be building web application that interacts with the MongoDB database, allowing users to filter tests by their fields displaying the metadata of the tests along side statistics of the metadata.
+The aim of the web application is to make the metadata easily accessible.
+
+% Para 6 --> resumo do resto do relatorio
+This document is structured in the following way.
+Section \ref{sec:Goals} has the objectives of this thesis.
+Section \ref{sec:Background} describes the field in which this thesis is inserted, describing the \texttt{ECMAScript} standard, followed by a description the \texttt{Test262} test suite, and a description of \texttt{MetaData262 v0}.
+Next is section \ref{sec:Related Work} that presents the scientific work done that relates to this thesis theme.
+In section \ref{sec:Design and Methodology} is the design and methodology that this thesis is implemented with.
+As for section \ref{sec:Evaluation and Planning}, it describes how we plan on evaluating the results of this thesis as well as its timeline.
+Finally, section \ref{sec:Conclusion} is where we present the main conclusions and a summary of this proposal.
+
 \section{Goals}
 \label{sec:Goals}
+
+This thesis has two main goals , the first is to improve \texttt{MetaData262 v0} and the second is to create a web visualization system to make the metadata more accessible.
+This way building the \texttt{MetaData262 Visualization System} that will allow its users to easily search the \texttt{Test262} metadata  filtering it by its characteristics.
+
+\paragraph{MetaData262 v0 improvements}
+This goal aims to improve \texttt{MetaData262 v0} in the precision and performance while making it more complete and scalable and easier to replicate.
+
+\paragraph{MetaData262 Visualization System}
+This goals aims to create a web platform to easily visualize the metadata calculated by \texttt{MetaData262 v0}.
+The system will provide filters with which a user will be able to filter the test by its characteristics.
+Then, the system will provide the filtered tests in various formats as well as useful statistics about the tests selected.
 
@@
 % overview da linguagem JS
-
-% paragrafo - porque e' que javascript e' relevante (uma das linguagens mais usadas no momento)
-JavaScript (JS) is a programming language mainly used in the development of client side web applications, also being one of the most popular programming languages.
-According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as second most active languages on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} as well as on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}
 
@@
   2012 & Gardner et al.~\cite{Gardner-2012} adapts the logic ideas of separation to build a scalable Hoare logic for a subset of JavaScript, based of semantics faithful to ES3.\\
-\end{tabular}
-\vspace{1em}
-\caption{Time line that contains the most relevant related work to the objectives of the project (1 of 2)}
-\label{tab:related-work_timeline-1}
-\end{table}
-\vspace{-2em}
-
-\begin{table}[ht!]
-\renewcommand{\arraystretch}{1.8}
-\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
-
   2012 & Politz et al.~\cite{Politz-2012} extends the language $\lambda$JS of ES3 to ES5 included, fir the first time, a formal semantics of JavaScript \texttt{property descriptors} and complete handling of the \texttt{eval} declaration.\\
@@
 In order to do this, the authors connected the \emph{JSRef} to the implementation of the Array library of the V8 engine \cite{V8} from Google.
-Additionally, the authors evaluated the state of the project \emph{JSCert} at the time, giving a detailed analysis of the methodology as a whole and a detailed analysis of the tests that the project passed or failed.\\
+Additionally, the authors evaluated the state of the project \emph{JSCert} at the time, giving a detailed analysis of the methodology as a whole and a detailed analysis of the tests that the project passed or failed.
+\end{tabular}
+\vspace{2em}
+\caption{Time line that contains the most relevant related work to the objectives of the project (1 of 2)}
+\label{tab:related-work_timeline-1}
+\end{table}
+
+\begin{table}[ht!]
+\renewcommand{\arraystretch}{1.8}
+\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
+
   2017 & Fragoso Santos et al.
@@
 \end{table}
-% QUESTION it e' Test262??
+% % QUESTION it e' Test262??
 

### 2022-05-27 paper.tex
Last version before submition

--- 2022-05-26 paper.tex
@@
 
-\title{Infra-estrutura de Testes para Implementações de Referência do Standard ECMAScript}
+\title{MetaData262: Computing and Visualizing Test262 MetaData}
 \subtitle{}
@@
 %
-\titlerunning{Live Metadata for Test262}
+\titlerunning{MetaData262: Computing and Visualizing Test262 MetaData}
 % If the paper title is too long for the running head, you can set
@@
 \begin{abstract}
-The ECMAScript standard is the specification for JavaScript which is one of the most popular programming languages in the world.
-Implementing the standard is challenging in various ways, both due to its complexity and its scale.
-Test262 was created to alleviate the implementations difficulty by providing a conformance test suite that allows developers to test their implementations.
-There is the project MetaData262 v0 that calculates the metadata of the Test262 test suite.
-With this thesis, we will improve the MetaData262 v0 metadata calculations as well as build infrastructure to view the metadata through a web interface.
-The web interface will also allow to filter tests by some characteristics such as built-ins and version of the standard as well as display statistics of the metadata.
+The ECMAScript standard is the specification of the Java-Script language which is one of the most popular programming languages in the world.
+Implementing the ECMAScript standard is challenging in various ways, both due to its complexity and its scale.
+Test262 was created to alleviate this difficulty by providing a comprehensive test suite that allows developers to test their implementations.
+%
+However, Test262 does not come to a structured description of its tests, making it impossible to filter tests by the JavaScript features they target.
+The MetaData262 v0 system was created to help with this issue, providing a database containing the metadata of Test262 tests and a collection of scripts to compute that metadata.
+%
+With this thesis, we will improve MetaData262 v0 in two separate ways.
+First, we will re-engineer the system for metadata computation, making it more precise and performant and enriching the computed metadata with additional information.
+Second, we will develop a web interface for the visualization of Test262, allowing developers to easily filter tests by feature.
 
@@
 
+% paragrafo - porque e' que javascript e' relevante (uma das linguagens mais usadas no momento)
+JavaScript (JS) is a programming language mainly used in the development of client-side web applications, also being one of the most popular programming languages.
+According to both GitHub and StakeOverflow statistics, JavaScript finished 2021 as the second most active language on GitHub\footnote{Second most utilized language based GitHub pull requests - https://madnight.github.io/githut/} as well as on StackOverflow.\footnote{Tendencies based on the Tags used - https://insights.stackoverflow.com/trends}
+
+% Para 2 --> implemetar JS e' dificil: complexidade tamanho
+% Para 2.5 --> Test262 ajuda a implementar permitindo testar implementacoes --> test262 tambem e' grande dificil selecionar os testes apropriados
+JavaScript is defined in the \texttt{ECMAScript} standard~\cite{ECMAScript}, a document written in English.
+The standard has been growing with each new version, both in size and complexity.
+The growth of the standard is illustrated in Figure \ref{fig:paginas_standard_version}.
+This continuous growth in complexity makes the JavaScript a language very hard to implement with many corner cases.
+\texttt{Test262}~\cite{Test262} was created to aid the development of JavaScript engines by providing a conformance test suite that developers can use to test their implementations.
+As the standard is extensive and complex, a conformance test suite for JavaScript must necessarily be very large given that it must cover all the corner cases of the standard.
+The standard being complex and extensive also leads to many of its implementations being only partial, particularly those made in academic contexts, for example JSRef~\cite{JSCert-2014} and JISET~\cite{JISET-2020}.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.7\textwidth]{images/standard_pages_evolution.png}
+    \caption{Evolution of ECMAScript standard's number of pages}
+    \label{fig:paginas_standard_version}
+\end{figure}
+
+% Para 4 --> MetaData v0 calcula metadata --> disponibiliza mongoDB
+
+When implementing a JavaScript engine or a JavaScript analysis, it is very important to know the tests that should pass for the specific feature being implemented.
+The selection of applicable tests for partial implementations is particularly difficult since tests for one feature may or may not use other features that are not implemented.
+For instance, a test targeting the Math library might use features of the array library.
+\texttt{MetaData262 v0} is a project developed at IST with the goal of facilitating this problem.
+In particular, MetaData262 v0 processes the \texttt{Test262} test suite and calculates metadata about the tests, storing that metadata in a NoSQL database that can be easily inspected.
+\texttt{MetaData262 v0} stores for each test various properties, such as the JavaScript version to which the test belongs, the syntactic constructs that are used in the test, and the built-ins that are used by the test.
+
+% --> eu vou melhorar metadata
+% Para 5 --> Visualization system permite interagir com a metadata via web com interface simples --> filtrar; verdados varias maneiras; estatisticas
+In this thesis we will improve \texttt{MetaData262 v0} in two ways.
+First, we will develop a \texttt{MetaData262 v1}, a new version of MetaData262 with an improved infrastructure for computing the metadata of Test262 tests, making it more precise and extending it with additional information.
+Second, we will build \texttt{MetaData262Viz}, a web application for the visualization of the metadata that will allow the developers to easily select tests per feature and obtain statistics regarding the selected testes.
+
+% Para 5.1 --> improve MetaData262 v0
+As described above, MetaData262 v1 will improve MetaData262 v0 in the following ways. Below, we give a brief overview of the most relevant improvements:
+
+\begin{itemize}
+\item \emph{Better Precision:} MetaData262 v0 computes for each test the version to which the test belongs and the built-in objects with which the test interacts.
+Unfortunately, this computation is far from precise as it relies on the specific implementations of JavaScript engines.
+For instance, MetaData262 v0 determines that a test is meant for version 5 if a JS engine implementing version 5 passes the test.
+Naturally, this approach is not error-free since it is never the case that a JS engine implements one and only one version of the language fully.
+To improve precision, we will build a dedicated infrastructure to streamline the addition of new JavaScript engines to our metadata computation pipeline and use this new infrastructure to compute a new version of the metadata.
+%
+\item \emph{Better Performance} As said before, the metadata computation pipeline of MetaData262 v0 requires running various JS engines on each test of Test262.
+This is extremely time consuming given the high number of tests contained in Test262 (around 40K).
+Currently this is done sequentially.
+To improve on this situation, we will associate each JavaScript engine with a docker container and run the docker containers in parallel.
+With docker containers running in parallel the scalability and performance of \texttt{MetaData262 v1} will improve.
+In addiction to that, the JavaScript engines being associated with docker containers will also make \texttt{MetaData262 v1} easier to replicate.
+%
+\item \emph{More information:} We will extend the computed metadata of each test with additional information.
+In particular, we plan to associate with each test: (1) the date in  which the test was introduced and (2) the harness files and functions that the test uses.
+Basically, the Test262 harness is a set of files that contain auxiliary functions used by Test262 tests, such as assert and assertEquals.
+\end{itemize}
+
+% Para 5.2 --> build MetaData262 visualization system
+MetaData262Viz will be a visualization system for the metadata computed with MetaData262 v1.
+In particular, MetaData262Viz will allow developers to filter Test262 tests by feature and to obtain statistics about the selected tests.
+The statistics will generate charts built in real time with the metadata of the tests selected.
+Some examples of those charts to generate are a bar charts with the amount of tests per version and a pie chart with the the amount of tests per built-in used.
+We have already developed a prototype of MetaData262Viz built with Node.js~\cite{Node.js} and React~\cite{React}.
+The prototype already allows filtering the tests by their version and the built-ins used.
+As for the output, it can already display the full metadata or just the path of the tests selected.
+
+% Para 6 --> resumo do resto do relatorio
+This document is structured in the following way.
+Section \ref{sec:Goals} has the objectives of this thesis.
+Section \ref{sec:Background} gives an overview of the field of the thesis, describing the \texttt{ECMAScript} standard, the \texttt{Test262} test suite, and the project \texttt{MetaData262 v0}.
+Next, Section \ref{sec:Related Work} presents the scientific work done that relates to this thesis topic.
+Section \ref{sec:Design and Methodology} presents the design and methodology that will be be followed in this thesis.
+As for Section \ref{sec:Evaluation and Planning}, it describes how we plan to evaluate the results of this thesis as well as its timeline.
+Finally, Section \ref{sec:Conclusion} presents the main conclusions and a summary of this proposal.
+
 \section{Goals}
 \label{sec:Goals}
+
+This thesis has two main goals. The first is to improve \texttt{MetaData262 v0} and the second is to create a web visualization system to make the metadata more accessible.
+%
+In the end, we hope that developers of JS engines and analyses will use the MetaData262Viz for filtering Test262 tests per feature, making the test selection process more efficient, effective, and reliable.
+
+\paragraph{MetaData262 v0 improvements}
+This goal is to improve the precision and performance of MetaData v0 making it more complete, scalable, and easier to replicate.
+
+\paragraph{MetaData262Viz}
+This goal aims to create a web platform to easily visualize the metadata calculated by \texttt{MetaData262 v0}.
+The system will provide filters with which the developer will be able to filter Test262 tests by their features.
+Then, the system will output the filtered tests in various formats as well as useful statistics about them.
 
@@
 % Falar sobre o comité -  quem controla a evolucao do ECMAScript
-ECMAScript standard\cite{ECMAScript} is the official document, written in the English language, in which the JavaScript language is defined.
+ECMAScript standard is the official document, written in the English language, in which the JavaScript language is defined.
 This document is in constant evolution, being updated by the ECMA Technical Committee 39 (TC39), which is responsible for maintaining the standard.
@@
 The standard specifies the \texttt{JavaScript} language, to ensure its multiple compilers and interpreters implementations are coherent.
-Some of the \texttt{JavaScript} compilers are the Hop~\cite{Hop} and the JSC~\cite{JSC} compilers, the most popular interpreters are Node.js~\cite{Node.js} and SpiderMonkey~\cite{SpiderMonkey}.
+Some of the \texttt{JavaScript} compilers are the Hop~\cite{Hop} and the JSC~\cite{JSC} compilers, the most popular interpreters are Node.js and SpiderMonkey~\cite{SpiderMonkey}.
 These are only four implementations among many others, which come along with the many use cases that \texttt{JavaScript} has.
@@
 % MetaParagrafo tres tipos
-\colorbox{orange}{The remaining} subsection provides a description of the three types of artifact described in the standard.
-% QUESTION ainda nao foi lido pelo prof
+The remaining subsection provides a description of the three types of artifact described in the standard.
 
@@
 Figure \ref{fig:Array_pop_example} shows a simplified version of an array performing the pop function, where \texttt{(a)} and \texttt{(b)} are the before and after respectively.
-Before preforming \texttt{pop} \texttt{(a)}, the array has three  properties \texttt{length}, \texttt{0}, and \texttt{1}.
+Before performing \texttt{pop} \texttt{(a)}, the array has three  properties \texttt{length}, \texttt{0}, and \texttt{1}.
 Property \texttt{length} represents the size of the array that has value \texttt{2}, while the properties \texttt{0} and \texttt{1} store the first (\texttt{banana}) and second (\texttt{kiwi}) elements of the array respectively.
-After \texttt{pop} is preformed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}) and the \texttt{length} property (highlighted in green) is decremented by one since the size of the array changes to one.
+After \texttt{pop} is performed \texttt{(b)}, the last element is of the array is removed (highlighted in red at \texttt{(a)}) and the \texttt{length} property (highlighted in green) is decremented by one since the size of the array changes to one.
 
-% TODO change length with quotation marks
 \begin{figure}[ht]
@@
 Implementing a \texttt{JavaScript} engine is particularly difficult since it involves dealing with the many corner cases that exist in the language.
-To test that corner cases are correctly dealt with there is \texttt{Test262}\cite{Test262}, the ECMAScript standard test battery.
-Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity, the total number of tests is around \colorbox{orange}{39837} divided into \colorbox{orange}{87} subfolders, each correspond to roughly one section of the standard.
+To test that corner cases are correctly dealt with there is \texttt{Test262}, the ECMAScript standard test battery.
+Although, \texttt{Test262} is vital to the \texttt{JavaScript} engines, it is very hard to maintain due it's complexity, the total number of tests is around 40000 divided into 87 subfolders, each correspond to roughly one section of the standard.
 \texttt{Test262} complexity grows with changes to the standard since in most cases backward compatible is maintained except for a few select cases.
@@
 %
-\item \texttt{flags} - contains a list of booleans for each test property, the properties being: (1) \texttt{onlyStrict}, the test is only executed in strict mode; (2) noStrict, the test will only be executed in mode \emph{sloppy}; (3) \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; (4) \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; (5) \texttt{async}, the test is contains asynchronous functions; (6) \texttt{generated}, the test generates the files specified by the property; (7) \texttt{CanBlockIsFalse} and (8) \texttt{CanBlockIsTrue}, the test will run if the property \texttt{CanBlock} of the \colorbox{orange}{\texttt{Agent Record}} executing it is false and true respectively; (9) \texttt{non-deterministic}, indicates that the semantics used in the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
+\item \texttt{flags} - contains a list of booleans for each test property, the properties being: (1) \texttt{onlyStrict}, the test is only executed in strict mode; (2) noStrict, the test will only be executed in mode \emph{sloppy}; (3) \texttt{module}, the test must be integrated as a \texttt{JavaScript} module; (4) \texttt{raw}, executes the test without any modification, which implies running as \texttt{noStrict}; (5) \texttt{async}, the test is contains asynchronous functions; (6) \texttt{generated}, the test generates the files specified by the property; (7) \texttt{CanBlockIsFalse} and (8) \texttt{CanBlockIsTrue}, the test will run if the property \texttt{CanBlock} of the \texttt{Agent Record} executing it is false and true respectively; (9) \texttt{non-deterministic}, indicates that the semantics used in the test are intentionally under-specified and therefore the test passing or failing should not be regarded as an indication of reliability or conformance;
 % QUESTION clarificar se e' necessario citar o que e' 'Agent Record' uma vez que nao vou explicar
@@
 Some tests also have the wrong metadata.
-As part of this thesis, we plan to process all the tests to check and correct their corresponding metadata as well as completing the metadata that is \colorbox{orange}{missing.}
+As part of this thesis, we plan to process all the tests to check and correct their corresponding metadata as well as completing the metadata that is missing.
 
@@
 %% ...
-\colorbox{orange}{The example} in Figure \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id.
+The example in Figure \ref{fig:Test262_example} has 2 keywords, description and the deprecated es5id.
 Besides the obvious upgrade from \texttt{es5id} to \texttt{esid} it would be useful to have \texttt{includes} with the \texttt{harness} files needed to execute the test.
@@
 
-Scientific studies about analysis of programs and techniques of instrumentation of JavaScript encompasses a varied range of specifications such as: type systems (\cite{Type_Analysis_for_JavaScript}, \cite{Towards_Type_Inference_for_JavaScript}, \cite{Understanding_TypeScript},  \cite{Fast_and_precise_hybrid_type_inference_for_JavaScript}, \cite{Depedent_Types_for_JavaScript}, \cite{Safe_&_Efficient_Gradual_Typing_for_TypeScript},  \cite{Type_Inference_for_Static_Compilation_of_JavaScript}), abstract interpreters (\cite{SAFE:Formal_Specification_and_Implementation_of_a_Scalable_Analysis_Framework_for_ECMAScript}, \cite{JSAI:a_static_analysis_platform_for_JavaScript}, \cite{Scalable_and_Precise_Static_Analysis_of_JavaScript_Applications_via_Loop-Sensitivity},  \cite{Combining_string_abstract_domains_for_JavaScript_analysis:an_evaluation)}), \emph{points-to} analysis (\cite{Points-to_analysis_for_JavaScript}), Hoare logic (\cite{Javert-2017}, \cite{Javert-2019}), operational semantics (\cite{JSCert-2014}, \cite{KJS-2015}, \cite{JSExplain-2018}), intermediate representations and compilers (\cite{Dynamic_property_caches:a_step_towards_faster_JavaScript_proxy_objects}).
-Next, we have the related work  most relevant for this thesis, it is organized into two Tables \ref{tab:related-work_timeline-1} and \ref{tab:related-work_timeline-2}.
-This time line focuses on the most important investigation work, focused on the JavaScript language specification since its creation in 1995.
+There is a wide range of scientific studies on program analysis for JavaScript, including: type systems (\cite{Type_Analysis_for_JavaScript}, \cite{Towards_Type_Inference_for_JavaScript}, \cite{Understanding_TypeScript},  \cite{Fast_and_precise_hybrid_type_inference_for_JavaScript}, \cite{Depedent_Types_for_JavaScript}, \cite{Safe_&_Efficient_Gradual_Typing_for_TypeScript},  \cite{Type_Inference_for_Static_Compilation_of_JavaScript}), abstract interpreters (\cite{SAFE:Formal_Specification_and_Implementation_of_a_Scalable_Analysis_Framework_for_ECMAScript}, \cite{JSAI:a_static_analysis_platform_for_JavaScript}, \cite{Scalable_and_Precise_Static_Analysis_of_JavaScript_Applications_via_Loop-Sensitivity},  \cite{Combining_string_abstract_domains_for_JavaScript_analysis:an_evaluation)}), \emph{points-to} analysis (\cite{Points-to_analysis_for_JavaScript}), Hoare logic (\cite{Javert-2017}, \cite{Javert-2019}), operational semantics (\cite{JSCert-2014}, \cite{KJS-2015}, \cite{JSExplain-2018}), intermediate representations and compilers (\cite{Dynamic_property_caches:a_step_towards_faster_JavaScript_proxy_objects}).
+Here, we focus only on the related work most relevant for this thesis, covering all the most relevant academic efforts in modelling the JavaScript semantics.
 
-% https://tex.stackexchange.com/a/196808
-\begin{table}[h!]
-\vspace{-2em}
-\renewcommand{\arraystretch}{1.8}
+One of the first formal semantics of the JS language was presented by Thiemann~\cite{Thiemann-2005} as part of his work on defining a type system for JavaScript.
+Thiemann's semantics targets the third version of the language and is specified in small-step style.
+Importantly, this semantics ignores a number of important features of the JavaScript language, such as implicit type coercion, the treatment of global variables, and the non-strict mode.
 
-\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
-  1995 & JavaScript is launched and implemented in Netscape Navigator 2.0 beta 3.\\
-  2005 & Thiemann presents the first try at defining a system of types possible for JavaScript ~\cite{Thiemann-2005}.
-The system detects the possible characteristics of an object and signals suspected type conversions.
-The work done does not cover important aspects of JavaScript, such as inheritance based on prototypes, and was not created along with an implementation at the time of the publication.\\
-  2008 & Maffeis et al.
-proposes a first operational semantics for the language ECMAScript version 3.
-(ES3)~\cite{operational_semantics_javascript-2008}.
-This first semantics was used to analyze various properties of security of web applications.\\
-  2010 & Guha et al.~\cite{essence_javascript-2010} defines a language $\lambda$JS, a formal calculation that encapsulates the fundamental characteristics of ES3.
-$\lambda$JS comes along an interpreter \emph{Racket}~\cite{Racket} and a translator from ES3 programs to expressions $\lambda$JS.\\
-  2012 & Gardner et al.~\cite{Gardner-2012} adapts the logic ideas of separation to build a scalable Hoare logic for a subset of JavaScript, based of semantics faithful to ES3.\\
-\end{tabular}
-\vspace{1em}
-\caption{Time line that contains the most relevant related work to the objectives of the project (1 of 2)}
-\label{tab:related-work_timeline-1}
-\end{table}
-\vspace{-2em}
+In 2008, Maffeis et al.
+proposed the first operational semantics for a significant subset of ECMAScript~\cite{operational_semantics_javascript-2008}.
+They target the 3rd version of the language.
+This semantics was formalized in big-step style and comprises a large collection of rules.
+Unfortunately, the semantics has not been automated, meaning that it cannot be tested.
+Nevertheless, the authors use the proposed semantics to reason about various security properties of JS web applications.
 
-\begin{table}[ht!]
-\renewcommand{\arraystretch}{1.8}
-\begin{tabular}{@{\,}r <{\hskip 2pt} !{\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}} >{\raggedright\arraybackslash}p{316pt}}
+In 2010, Guha et al define lambdaJS~\cite{essence_javascript-2010}, a lambda calculus to be used as a compilation target for JavaScript.
+The authors claim that lambdaJS capture the fundamental behavior of the JS language.
+The authors also define a desugaring function that compiles ES 3 programs to lambdaJS and implement a lambdaJS interpreter in Racket.
+However, the authors do not test their infrastructure against Test262.
+In order to test this work against Test262, one would have to compile all Test262 tests to lambdaJS and then run the compiled test suite using the lambdaJS interpreter.
+Importantly, the lambdaJS infrastructure does not support the entire ES3 standard.
+Hence, if one was to test the proposed infrastructure, one would first have to clearly define the target subset of ES3 and filter the appropriate tests from Test262.
+Later, the same authors propose S5~\cite{Politz-2012}, an extension of lambdaJS to the 5th version of the standard.
+In contrast to lambdaJS, which the authors did not test, the authors do test S5, running it against Test262.
+They pass a total of 8157 tests out of a total of 11606 tests (approx.
+70\%).
+However, the authors admit that they do not implement the whole language, stating that they only implement 60\% of the built-in objects.
+Nevertheless, they were not able to filter the tests appropriately to include only the tests with the features that they do implement.
+This means that they might have bugs in features that they implement that were disregarded as being caused by unimplemented features.
+Without an appropriate labeling of the tests, the only way of guaranteeing that they are not failing tests that they should be passing is to analyze all failing tests manually, which the authors do not do.
 
-  2012 & Politz et al.~\cite{Politz-2012} extends the language $\lambda$JS of ES3 to ES5 included, fir the first time, a formal semantics of JavaScript \texttt{property descriptors} and complete handling of the \texttt{eval} declaration.\\
-  2014 & Bodin et al.
-presents \emph{JSCert}~\cite{JSCert-2014}, a formalization of the ES5 standard written in \emph{Coq}~\cite{Coq}, and \emph{JSRef} an executable interpreter of references extracted from \emph{Coq} to \emph{OCaml}~\cite{OCaml}, proving correct when related to \emph{JSCert} and tested using the test suite Test262~\cite{Test262}.\\
-  2015 & Park et al.
-presents \emph{KJS}~\cite{KJS-2015}, a formal semantics for ES5, the most complete to that moment.
-The \emph{KJS}  was developed using the framework \emph{K}~\cite{Kframework}, a consolidated system of rewriting of terminology that supports various types of symbolic analyses based on \emph{reachability logic} (All-Path Reachability Logic \cite{All-Path_Reachability_Logic}).
-\emph{KJS} was tested carefully using the test suite, passing all the 2.782 tests of the language core.
-The \emph{KJS} can be executed symbolically, being able to be used for formal analyses and verification of ES5 programs.\\
-  2015 & Gardner et al.
-\cite{JSCert_1Year-2015} extend the project \emph{JSCert} to support \emph{arrays} ES5.
-In order to do this, the authors connected the \emph{JSRef} to the implementation of the Array library of the V8 engine \cite{V8} from Google.
-Additionally, the authors evaluated the state of the project \emph{JSCert} at the time, giving a detailed analysis of the methodology as a whole and a detailed analysis of the tests that the project passed or failed.\\
-  2017 & Fragoso Santos et al.
-creates the tool \emph{JaVert}~\cite{Javert-2017}: a tool for verifying JavaScript that allows semiautomatic reasoning about the fundamental properties of correctness of ES5 programs.
-JaVert does not assume any simplification of the language semantics, analyzing all the \emph{corner-cases} of the language.
-The load of necessary notation is substantial.\\
-  2018 & Charguéraud et al.
-presents \emph{JSExplain}~\cite{JSExplain-2018}, an reference interpreter for the ES5 language that follows the specifications and produces an inspectable execution traces.
-It also includes an interface of code inspection that allows the programmer to execute his ES5 code step by step as well as the code of the ES5 interpreter, while executing ES5 programs.\\
-  2020 & Sampaio et al.
-~\cite{Javert-2020} extend the JaVerT tool to support automatic reasoning about JavaScript events.
-In order to do this, the authors developed a reference implementation of JavaScript promises directly in JavaScript.
-The authors used \colorbox{orange}{it} to test their reference implementation, passing 344 tests of the 474 total dedicated to JavaScript promises.
-Around 106 testes were excluded due to ES6 features that were not yet supported and to tests that require \emph{non-strict} mode.\\
-  2020 & Jihyeok Park et al. presents JISET~\cite{JISET-2020} the first tool that synthesizes automatically the analyzer and that translates AST-IR directly from a specification of a given language.
-\end{tabular}
-\vspace{1em}
-\caption{Time line that contains the most relevant related work to the objectives of the project (2 of 2)}
-\label{tab:related-work_timeline-2}
-\end{table}
-% QUESTION it e' Test262??
+In 2014, Bodin et al.
+present JSCert~\cite{JSCert-2014} a formal semantics of JavaScript implemented in the Coq proof assistant~\cite{Coq}.
+JSCert targets the 5th version of the standard and is written in pretty-big-step style~\cite{Arthur-Chargueraud_ESOP_2013}.
+Besides the semantics, the authors also implement a JS interpreter in Coq~\cite{Coq} called JSRef.
+They prove that JSRef is correct with respect to JSCert and use the extraction mechanism of Coq to obtain an executable version of JSRef in OCaml~\cite{OCaml}.
+Using the OCaml version of JSRef, the authors were able to test JSCert/JSRef against Test262.
+JSRef passes 1796 tests out of a total 2782 tests for the core language.
+Note that the tests for the core language include all the test that do not explicitly target the built-in objects.
+Given that JSCert models the entire core language, the following question naturally arises: how come does JSCert fail around 35\% of the tests targeting the core language?
+The answer is simple: many tests targeting the core language interact with built-in objects, such as arrays, strings, the methods of the Object object, the methods of the Math object etc.
+Again, the authors were not able to automatically filter the Test262 tests for the tests that do not make use of built-in objects.
+%
+Later, some the authors of JSCert developed an extension of the JSCert project in which they connected JSRef with the V8 Array implementation.
+More concretely, instead of implementing the array methods directly in Coq, the authors used the JavaScript implementations of these methods that came with the V8 JS engine.
+The authors then tested the extended version of JSRef against Test262 passing 2440 tests out of 2782 tests.
+Note that even with the V8 array library some tests fail.
+This can be caused by the fact that they make use of other built-in object or because there actually bugs in the JSCert/JSRef implementation.
+In order to understand the precise reasons, one would have to manually inspect all the failing tests.
+
+In 2015, Park et al develop KJS~\cite{KJS-2015} a formalization of the 5th version of the ECMAScript standard in K re-writing framework.
+KJS is written in small-step style and allows for symbolic reasoning over JS programs through the symbolic facilities of the K framework.
+KJS was tested against Test262 passing  2782 tests out of a total of 2782 tests for the core language.
+Interestingly, the total number of tests presented for the core language in the KJS paper does not coincide with the number presented in the JSCert project.
+This discrepancies make the results of these papers difficult to assess.
+One of the goals of this thesis is precisely to improve this situation by establishing a ground truth.
+
+Recently, Chargueraud et al.
+developed JSExplain~\cite{JSExplain-2018}, a new JavaScript reference interpreter written in a subset of OCaml.
+JSExplain is written in big-step style with the goal of being as close as possible to text of the standard.
+JSExplain comes with a compiler from the used subset of OCaml to JavaScript, using which one can interpret JavaScript programs directly in JavaScript.
+With this compiler, the authors develop a new JavaScript code stepper with which the developer can code step not only their own code but also the code of the JavaScript interpreter.
+JSExplain is also tested  against Test262.The authors do not provide a detailed break down of the tests, reporting only that JSExplain passes more than 5000 tests.
+Again, a precise evaluation would require a clear definition of the targeted features of the language and the selection of the corresponding tests.
+With MetaData262, this can be done easily.
+
+Finally, JISET~\cite{JISET-2020} was the last JS reference interpreter developed by the research community.
+It targets the 10th version of the language and its code is semi-automatically generated from the HTML version of the specification.
+Unsurprisingly, the authors also test JISET against Test262, reporting that it passes  16355 tests out of a total of 35990 tests.
+The authors justify the high failing rate by stating that they do no model most of the JS built-in objects.
+However, the failing tests might or might be related to unmodelled features.
+Again, in order to ensure that the failing tests were caused by unmodelled features, one would have to analyze all of them manually, which is impractical given the high number of tests.
 
@@
 We recall that this thesis has two goals.
-Firstly, we want to improve \texttt{MetaData262 v0} in two ways: add support for missing metadata elements, and improve the infrastructure for the metadata computation, making the obtained metadata more reliable and the process faster and reproducible.
+Firstly, we want to improve MetaData262 v0 in two ways: add support for missing metadata elements, and improve the infrastructure for the metadata computation, making the obtained metadata more reliable and the process faster and reproducible.
 Secondly, we want to create a visualization system for easy selection of tests and display of test statistics.
@@
 If the properties of a built-in are changed directly then it will not be detected.
-For example if a test preforms \texttt{Array.prototype.pop = 3;} then the property \texttt{prototype} of the Array built-in will be changed even though no function call was made.
+For example if a test performs \texttt{Array.prototype.pop = 3;} then the property \texttt{prototype} of the Array built-in will be changed even though no function call was made.
 This can not be detected by the current dynamic approach.
@@
 This calculation can be improved by completing \texttt{includes} metadata and adding an additional \texttt{harness-functions} metadata.
-Both \texttt{includes} and \texttt{harness-functions} are mentioned in earlier sections, the first stores the files of the \texttt{harness} needed to run the test and the second will store the functions from those files that are actually used in the test.
+Both \texttt{includes} and harness-functions are mentioned in earlier sections, the first stores the files of the \texttt{harness} needed to run the test and the second will store the functions from those files that are actually used in the test.
 \end{itemize}
@@
 % built-ins --> em particular, apanhar property look-ups dinamicamente
-The improvements the \texttt{MetaData} calculation precision, in particular the version dynamic approach, will enhance the issue of performance.
-This issue will be improved with the use of \texttt{docker}\cite{docker} containers and this way we will also improve the ease to replicate the calculations of the \texttt{MetaData}.
-%
-The duration to calculate the version with the dynamic approach scales linearly with the amount of engines per standard version being used.
-In \texttt{MetaData262 v0} only two JS engines are being used and the dynamic approach already takes multiple hours to run.
-%
-With the use of \texttt{docker} containers we will be able to parallelize the work of the dynamic approaches by multiplying the docker images that run the JS engines.
-Containers running in parallel will not only improve the performance but also allow MetaData262 to be replicated in an easier way since it requires a smaller configuration effort to replicate the \texttt{MetaData} calculation.
-% TODO rescrever comentarios do prof
+An important contribution of this thesis is the set up of an automated infrastructure for the computation of the metadata.
+As discussed above, the dynamic computation of the ES version of each test, requires running various versions of multiple JS engines.
+So far, in order to do so, these engines must be installed in the local machine running the dynamic-version-computation script.
+This is impractical as we consider a larger number of engines.
+To ameliorate this issue, we plan to use docker containers, assigning a docker container to each version and making all the containers available as part of our metadata computation infrastructure.
+This architecture will enable us to parallelize the work of the dynamic approaches by multiplying the docker containers that run the JS engines.
+This parallelization will allow scalability because otherwise since the duration to calculate the version with the dynamic approach will scale linearly with the amount of engines per standard version being used.
+Docker containers will not only improve the performance but also allow MetaData262 to be replicated in an easier way since it requires a smaller configuration effort to replicate the \texttt{MetaData} calculation.
 
-\subsection{MetaData262 Visualization System}
+\subsection{MetaData262Viz}
 \label{sub:metadata262_visualization_system}
@@
 % Para 0 --> Motivar a criacao do system de visualizacao
-The \texttt{MetaData262 Visualization System} appears to ease the barrier to access the \texttt{MetaData262 v0} information and statistics.
+The \texttt{MetaData262Viz} appears to ease the barrier to access the \texttt{MetaData262 v0} information and statistics.
 Right now the metadata is hosted in a MongoDB database leading to the need to have knowledge of MongoDB to get the information.
 After getting the information there would most likely be a need to process that metadata and maybe even generating charts to provide statistics.
-This is visualization system will make this process much easier.
+This visualization system will make this process much easier.
 
@@
 % Para 1 --> O que o website vai mostrar e o que para isso e' necessario fazer
-The \texttt{MetaData262 Visualization System} is in its core a web application that allows users visualize and filter the \texttt{Test262} metadata easily.
+The \texttt{MetaData262Viz} is in its core a web application that allows users visualize and filter the \texttt{Test262} metadata easily.
 The system will also update itself whenever there are changes in \texttt{Test262}.
@@
 % Para 2 --> O que ja esta feito
-The development of the web application has already began, \colorbox{orange}{Figure} \ref{fig:website} shows the progress already made.
-The image has 2 main parts: the filters part, and the information part.
+The development of the web application has already began, Figure \ref{fig:website} shows the progress already made. The image has 2 main parts: the filters part, and the information part.
 %
@@
 
-% TODO rescrever tudo com comentarios do prof
 % Para 3 --> Trade off the fazer a filtragem local or no servidor
-In Figure \ref{fig:website} there are two search buttons because there at the time trade offs being made with either way of searching, the following results were run in the laptop serving the web application.
-The backend search starts getting slow whenever the amount of tests being returned surpasses 2500 in the local machine, taking on roughly half a second to get the results.
-The local search has the problem that it takes around 4 seconds to get all the metadata from the server at page load.
-The local search has the advantage of filtering the tests in a few milliseconds.
-Caching the metadata in the users browser would allow the metadata to be only loaded one, making the usability of the web application much better.
+Figure \ref{fig:website} two approaches to the search functionality: server-side search (\emph{Backend Search}) and client-side search (\emph{Local Search}).
+We provide both approaches because each has its trade offs.
+The following results were achieved using the laptop that was serving the web application.
+The server-side search starts slowing down whenever the amount of tests being returned surpasses 2500, taking roughly half a second to get the results.
+Whereas, the client-search takes around 4 seconds to load all the metadata from the server at page load but is able to filter any query within a few milliseconds.
+We believe the client-side search to be the most appropriate since is only needs the initial load time.
+In addition to that, the user can fill in the filters for the search while the data is loading.
+To further improve this issue, we will be caching the metadata in the users browser this way the metadata would be loaded once.
 
@@
 % Para 4 --> Os graficos que queremos gerar --> bibliotecas
-This web application will allow the user to view charts about the metadata of the tests queried.
-M.
-Trigo's thesis he generated the charts shown in Figure \ref{fig:grafico_bigodes} and Figure \ref{fig:grafico_stackBar} as well as some other pie charts and bar charts.
-In the web application those charts will be generated in real time with the help of external libraries.
-Especially the charts below, the box and whiskers chart and the stacked bar chart, are not supported by many of the most popular libraries.
-Two of the libraries that are capable of generating all the charts needed are the \texttt{KendoReact}\cite{KendoReact} an UI library and \texttt{Victory}\cite{Victory} a chart library.
-\texttt{KendoReact} is a full UI library, charts being only part of its 126 components.
-Whereas, the \texttt{Victory} library has only 34 components and all are directly related with generating charts.
-The Victory library also seams simpler to use, the community around it seams more active which would be important in case of difficulty while implementing the charts needed.
+The second use case of the web application is to allow the user to view statistical information about the tests and the queried tests in the form of a variety of charts, such as pie charts, bar charts and box and whiskers charts and stacked bar charts.
+In his thesis, M.
+Trigo generated some statistical data about the entire Test262 test suite by hand, introducing the relevant data in a spreadsheet software and then constructing the charts manually.
+Here, we intend to automate this process, generating those graphs in real time with the help of a JS graphical library.
+For instance, Figures 9 and 10 illustrate two graphs generated by M.
+Trigo:
+
+\begin{itemize}
+\item \emph{Figure \ref{fig:grafico_bigodes}} show a box and whiskers chart about the number of assert statements that are used in each category.
+\item \emph{Figure \ref{fig:grafico_stackBar}} show a stack bar chart with the amount of tests distributed by version in each category.
+\end{itemize}
+
+In order to generate such graphs in real time, we are considering two libraries:
+
+\begin{itemize}
+\item \emph{KendoReact\cite{KendoReact} -} This library a UI library that makes available to React 126 components.
+Of those 126 components one is \emph{Charts} with a total of 16 charts available.
+KendoReact is a very popular UI library for React.
+\item \emph{Victory\cite{Victory} -} This library focuses on chart generation.
+Victory has a total of 34 components directly related to chart generation.
+Victory is also open-source and one of the most popular chart libraries for React.
+\end{itemize}
+
+So far, we are more inclined to use Victory because because we only need the Chart components.
+Victory is also one of the most popular chart libraries with a big community which might be a big advantage if implementation difficulties arise.
 
@@
 % Para 1 --> components a serem testados
-The evaluation of this thesis will focus in its two main elements: the metadata calculation infrastructure and the \texttt{Metadata262 Visualization System}.
-The metadata calculation infrastructure will be evaluated for its precision and performance.
-As for the \texttt{Metadata262 Visualization System}, it will be evaluated for its reliability.
-Each criteria of the evaluation will be described in more detail bellow.
+The evaluation of this thesis will focus on its two main elements: the metadata calculation infrastructure and the \texttt{MetaData262Viz}.
+The metadata calculation infrastructure will be evaluated with respect to its precision and performance.
+As for the \texttt{MetaData262Viz}, it will be evaluated for its reliability.
+The evaluation criteria will be described in more detail bellow.
 
 % Para 2 --> Metadata calculation precision  
-\paragraph{Evaluation of metadata calculation precision}
-The precision of the metadata calculated is very hard to access since the is no ground truth available.
+\paragraph{Metadata calculation precision}
+The precision of the metadata calculated is very hard to access since there is no ground truth available.
 The version metadata calculated is being calculated for the first time.
-Thus, to evaluate the precision of the metadata calculation we will select a small random group of tests to analyze manually.
-A second evaluation will be made to the tests that have the biggest difference between the version calculated and the version that was the latest at the time the test was added to \texttt{Test262}.
-Firstly, we will find the date the tests were added to \texttt{Test262} and associate them with the latest \texttt{ECMAScript} version at that time.
-Afterwards, we will select the tests with the biggest discrepancy between the version associated and the version calculated.
-Finally we will preform a manual evaluation to ensure that the right version was calculated.
-Note that the version discrepancy is merely indicative, more the once tests were detected missing and subsequently added to complement \texttt{Test262}.
-These tests were added when they were detected missing and not when the version was released.
+To evaluate the precision of the metadata calculation we will analyze a small subset of the Test262 tests manually, comparing the version calculated with their real version.
+We will use two separate strategies for selecting tests for manual inspection:
+\begin{itemize}
+\item \emph{Random Strategy:} we will select a small subset of tests randomly, guaranteeing uniform coverage of the various test folders; 
+%
+\item \emph{Heuristic Selection:} we will select the tests that exhibit the largest different between the computed version and the version that was the latest at the time the test was added to \texttt{Test262}.
+To do this, we will extend the metadata of the tests with the date in which they were added to Test262 or updated.
+It is often the case that tests are added for features of previous versions; this happens for instance when Test262 maintainers detect missing coverage issues, in such cases the newly added tests target the version that was active when the corresponding uncovered feature was introduced instead of the current one.
+\end{itemize}
 
 % Para 3 --> Metadata calculation performance
-\paragraph{Evaluation of metadata calculation performance}
-To evaluate performance improvements made to \texttt{MetaData262 v0} we will compare the time needed to preform the metadata calculation between the old and new infrastructures.
-This must be ran in the same system and with the least concurrent processes running in order to make sure we have the least interference, this way achieving the most trustworthy results.
+\paragraph{Metadata calculation performance}
+To evaluate performance improvements made to \texttt{MetaData262 v0} we will compare the time needed to perform the metadata calculation between the old and new infrastructures.
+This must be run in the same system and with the least concurrent processes running in order to make sure we have the least interference, this way achieving the most trustworthy results.
 
 % Para 4 --> visualization system
-\paragraph{Evaluation of the visualization system reliability}
-The visualization system evaluation will have three types of tests: unit tests, end-to-end tests and load tests.
+\paragraph{MetaData262Viz reliability}
+The MetaData262Viz evaluation will have three types of tests: unit tests, end-to-end tests and load tests.
 The unit tests will ensure the correctness of the system's results.
 The end-to-end tests will guarantee that the website interface works correctly.
-Finally, the load tests will allow us to estimate the amount of users that the systems can support, this way allowing us to estimate the resources need.
+Finally, the load tests will allow us to estimate the amount of users that the system can support, this way allowing us to estimate the resources need.
 
@@
 
-The Plan for this thesis is based on the tasks that need to be preformed in order to accomplish the objectives defined in section \ref{sec:Goals}.
-Thus, this thesis has 4 main task to accomplish: improving the metadata calculation, concluding the visualization system, evaluating the results, and writing the thesis.
+The Plan for this thesis is based on the tasks that need to be performed in order to accomplish the objectives defined in Section \ref{sec:Goals}.
+Thus, this thesis has 4 main task to accomplish: improving the metadata calculation, concluding the MetaData262Viz, evaluating the results, and writing the thesis.
 These tasks are represented in the Gantt diagram in Figure \ref{fig:gantt} and each task is described in more detail afterwards.
@@
 
-% gantt tese
 \begin{figure}[ht]
     \centering
-    \includegraphics[width=1\textwidth]{images/gantt_thesis.png}
+    \includegraphics[width=1\textwidth]{images/gantt_thesis_month.png}
     \caption{Gantt diagram with the thesis plan}
@@
 
-% gantt tese tempo dividido por mes
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=1\textwidth]{images/gantt_thesis_month.png}
-    \caption{Gantt diagram with the thesis plan}
-    \label{fig:gantt_month}
-\end{figure}
-
 \paragraph{Improving metadata calculation}
-This task consists of improving \texttt{MetaData262 v0} by improving its precision and performance as discussed in section \ref{sec:Design and Methodology}.
-The precision will be improved with more JS engines and a JS parser that supports more recent versions of the standard.
+This task consists of improving \texttt{MetaData262 v0} by improving its precision and performance as discussed in Section \ref{sec:Design and Methodology}.
+The precision will be improved with the inclusion of more JS engines and a JS parser that supports more recent versions of the standard.
 As for the performance, we will parallelize the metadata calculation using \texttt{docker} containers.
 
-\paragraph{Concluding visualization system}
+\paragraph{Concluding MetaData262Viz}
 This task consists of concluding \texttt{MetaData262 v0} visualization system by adding the possibility to view statistics about the queried tests and improving the system's reliability.
@@
 This task consists of evaluating the results of the metadata generated.
-The most prone to error metadata calculation will be checked by hand in order to guarantee the correctness of the results presented.
-The development of the tests for the visualization system and the system itself will overlap during the first two weeks of the evaluation.
+We will manually check the metadata calculated for a small subset of Test262 tests, using the selection strategies discussed in this.
+The development of the tests for the MetaData262Viz and the system itself will overlap during the first two weeks of the evaluation.
 
@@
 \label{sec:Conclusion}
-With this thesis we will build an infrastructure that enables convenient access to metadata about the \texttt{Test262} conformance suite.
-We will improve upon the metadata calculation of \texttt{MetaData262 v0} as well as making it accessible to users via a web interface.
+With this thesis we will build an infrastructure that enables convenient access to the metadata of the \texttt{Test262} test suite.
+We will improve the metadata calculation of \texttt{MetaData262 v0} and make it accessible to users via a web interface.
 This will enable \texttt{JavaScript} engine developers to filter tests easily, in particular developers of partial implementations.
-This will allow developers to find and correct errors quicker and will also enable a standardization of the selection of applicable tests for partial implementations.
-
-This report starts with a description of the background to this thesis, in particular the \texttt{ECMAScript} standard, the \texttt{Test262} conformance test suite and \texttt{MetaData262 v0}.
-Then, we present the scientific work done in the scope of the \texttt{JavaScript} language, with studies from 2005 to 2020.
-Then we have the design and methodology with a detailed description of the contributions this thesis will make.
-Finally, we have the methods we will use to evaluate the quality of the work done.
+This will allow developers to find and correct errors more quickly and also enable a standardization of test selection process for partial implementations.
 

### 2022-05-30 paper.tex
small fix

@@
 % Para 3 --> Trade off the fazer a filtragem local or no servidor
-Figure \ref{fig:website} two approaches to the search functionality: server-side search (\emph{Backend Search}) and client-side search (\emph{Local Search}).
+Figure \ref{fig:website} shows two approaches to the search functionality: server-side search (\emph{Backend Search}) and client-side search (\emph{Local Search}).
 We provide both approaches because each has its trade offs.
@@
 In addition to that, the user can fill in the filters for the search while the data is loading.
-To further improve this issue, we will be caching the metadata in the users browser this way the metadata would be loaded once.
+To further improve this issue, we will be caching the metadata in the users browser this way the metadata would only be loaded once.
 
